### 2016-09-10 KKosmology.tex
Version by simondp

\documentclass[aps,prd,12pt,superscriptaddress,showpacs,showkeys,longbibliography,reprint]{revtex4-1}
\usepackage{calrsfs} 
\usepackage{amsmath,amsthm,latexsym,amssymb,amsfonts}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage[%
  colorlinks=true,
  urlcolor=blue,
  linkcolor=blue,
  citecolor=blue
]{hyperref}
\usepackage{etoolbox}
\usepackage{breqn}

%% \makeatletter
%% \let\cat@comma@active\@empty
%% \makeatother

%------------------
%--------- Definitions
%------------------
\input{Def-article.tex}

\hypersetup{%
  pdftitle={Dimensional reduction on $S^1$ of Lanczos--Lovelock and Lovelock--Cartan theories of gravity},
  pdfauthor={Oscar Castillo-Felisola,}{Cristobal Corral,}{Simon del Pino,}{Francisca Ramirez.},
  pdfkeywords={Torsion,} {Generalised Gravity.},
  pdflang={English}
}

%------------------
%--------- Document
%------------------
\begin{document}

\title{Five Dimensional Kaluza--Klein Cosmology in Lovelock--Cartan Theory}

\author{Oscar \surname{Castillo-Felisola}}
\email{o.castillo.felisola@gmail.com}
\affiliation{\UTFSM.}
\affiliation{\CCTVal.}

\author{Cristobal \surname{Corral}}
\affiliation{\UTFSM.}
\affiliation{\CCTVal.}

\author{Sim\'on \surname{del~Pino}}
\email{simon.delpino.m@mail.pucv.cl}
\affiliation{\PUCV.}

\author{Francisca \surname{Ram\'irez}.}
\affiliation{\UTFSM.}

%% --------- Abstract
\begin{abstract}
 We study the Kaluza-Klein dimensional reduction of an $N$ dimensional Riemann--Cartan manifold when the compact section is $S^1$. First, we focus on the details of dimensional reduction in the presence of torsion to then study the particular Einstein--Cartan theory with the Gauss--Bonnet term in five dimensions. We look for the general cosmological solution of the Friedman-Robertson-Walker familly in the four-manifold to interpret the remaining fields. The model, treated in five dimensional vacuum, presents torsion both in the reduced manifold as well as in the compact section, manifest as a non-vanishing four dimensional matter field that adds to the effective energy density and pressure. This allows the model to present expanding periods as well as contracting universes, and a radius of the compact section that it asymtotes zero in some solutions. 
\end{abstract}

%% \pacs{02.40.Ma,04.50.Kd,04.90.+e}
%% \keywords{Affine Gravity, Torsion, Generalised Gravity.}

\maketitle

\section{Introduction}
The geometric theory of gravity, General Relativity (GR), has been highly successful in describing the physics at large scales. Two major predictions were recently confirmed in one experiment that invoque two of the most fascinating conepts that this theory developed: the propagation of ripples of spacetime, gravitational waves, originated by the merge of two black holes where detected by the LIGO collaboration \cite{Abbott:2016blz} confirming the existence of both, astrophysical balck holes as well as the unprecedented nature of the signal detected. With this, we can rely in the geometric nature of this interaction. At the galactic scale, the gravitational lensing produced by the local distribution of energy and matter suggest the presence of a kind of matter that is not seen through telescopes. This abundance of matter in the galaxy is also compatible with the velocity profyle of stars at the outer regions of it and is needed to reach the current state of structures formation. This matter would interact mostly, if not only, through gravity \cite{Sofue:2000jx}. At the cosmic scale, observatinal evidence points out an accelerated current state of the Universe \cite{Riess:1998cb}. Such a behaviour needs the existence of a form of vacuum energy of exotic nature to give account of the accelerated expansion. This is evidence of unknown physics, and it has been the motivation for looking for new gravitational degrees of freedom beyond Einsten's GR. Among the familly of extended gravity theories, higher dimensional gravity can give us a clue to explore the manifestation of new degrees of freedom as material fields of certain class. Along this lines is that we consider here an extended model of gravity given by the so called Lovelock action~\cite{Lovelock:1971yv}. 
The Einstein--Hilbert action with cosmological constant is the unique action in four dimensions that yields second order field equations for the metric. In higher dimensions, however, this does not remains true. A particular combination of squared terms in the curvature can be added to the gravitational action to give rise a second order system for the metric with no ghosts. The particular quadratic term is called Gauss--Bonnet term, and in $N$ dimensions reads
\begin{equation}\label{GB}
  \mathcal{L}_{GB}=\text{d}^N \!x \,\sqrt{-g}\left(\tilde{R}^2-4\tilde{R}_{\mu\nu}\tilde{R}^{\mu\nu}+\tilde{R}_{\alpha\beta\mu\nu}
  \tilde{R}^{\alpha\beta\mu\nu}\right),
\end{equation}
where $\tilde{R}_{\alpha\beta\mu\nu}$ is the Riemannian curvature of a manifold with metric $g_{\mu\nu}$ and $g$ its determinant. $\tilde{R}_{\mu\nu}$ and $\tilde{R}$ are the Ricci tensor and Ricci scalar built with the curvature.

Written in four dimensions, Eq.~\eqref{GB} adds no dynamics to the metric, since it represents a topological invariant proportional to the Euler characteristic and therefore can always be written locally as a boundary term, while in lower dimensions vanishes identically. Is then natural to extend gravitational Lagrangians by dimensional continuation of Euler densities to build a general theory for a metric in arbitrary dimensions, while keeping the two important features of Einstein's General Relativity just mentioned: the absence of ghosts and second order field equations. These family of Lagrangians for gravity where introduced by Lovelock and bear his name. It can be built in arbitrary dimensions as a linear combination of dimensionally continued topological invariants. Equivalently, the previous Lagrangian can be written in exterior forms, that provide shorter expressions together with a natural framework for further generalizations, in $N$ dimensions reads
\begin{equation}\label{GB2}
\mathcal{L}_{GB} = \epsilon_{a_1...a_N} \tilde{R}^{a_1a_2}\wedge\tilde{R}^{a_3a_4}\wedge e^{a_5}\wedge...\wedge e^{a_N}.
\end{equation}
This Lagrangian was also identified as the low energy correction for a spin two field in string theory~\cite{Zwiebach:1985uq}. Thus, the consideration of further Lovelock terms in a higher dimensional manifold can be interpreted as semiclassical corrections of a gravity action.
The higher dimensional scenario seems to be a natural framework for unification. Early in the last century, it was realized by T. Kaluza and developed by O. Klein, that the existence of one extra direction in Einstein geometric theory of gravity would give rise a unified picture of gravity and electromagnetism together with a spectrum of new heavy particles~\cite{Kaluza:1921tu,*Klein:1926tv}. This opened the posibility of a geometrical understanding of interactions given by non-Abelian gauge groups as a concequence of the topology of a spatial compact section in a higher dimensional spacetime. Eventhough the original model, and further developments have not yet reached a realistic phenomenology, it is clear that this possibility of unification of internal and spacetimes symmetries, together with a material manifestation of geometric degrees of freedom provides a fruitful arena for looking for clues of unity among the interactions. The compactification of higher Lovelock terms has been considered \cite{MuellerHoissen:1985mm,MuellerHoissen:1989yv} and its consequences in cosmology in \cite{MuellerHoissen:1985ij,Deruelle:1986iv,Deruelle:1989fj}. In the Riemannian five dimensional theory, exact wormholes solutions have been found in vacuum \cite{Dotti:2006cp,Dotti:2007az}, and suported by non-exotic matter, making the sign of the Gauss-Bonnet coupling the source of exotic behaviour \cite{Mehdizadeh:2015jra}, higher order Lovelock models \cite{Mehdizadeh:2015dta}, as well as in higher order compactified Lovelock theories with torsion \cite{Canfora:2008ka} also have these solutions. Classically, wormholes must be supported by a kind of phantom energy, compatible with the cosmological hypothesis. We see their theoretical existence in these extended models as a good omen for the search of new degrees of freedom that could provide explanation for the exotic matter-energy abundance in the Universe. \\
Torsional degrees of freedom, on the other hand, as a further feature of spacetime had a \textit{renaissance} with the works of D. W. Sciama and T. Kibble \cite{Kibble:1961ba} in Poincar\'e gauge theory. In that framework, spacetime torsion arises naturally as the field strength of a gauge connection for the Poincar\'e group (see \cite{Hehl:1976kj,Blagojevic:2002du} for a complete review). The pose of gravity as a gauge theory has major benefits in the problem of a unifyied picture for interactions since it allows it to interpret it as arising by means of a localization procedure, where spacetime is a Riemann-Cartan manifold. In this context, it has been shown that torsional functions produce an accelerating profyle compatible with the current state of the Universe \cite{Shie:2008ms}, as well as non-minimal coupling of Euler forms in Riemann--Cartan spacetimes allow torsion to behave as matter of exotic characteristics \cite{Toloza:2013wi} customary for expansion.
It becomes essential to understand Kaluza--Klein's ideas in the presence of torsion. Work on the subject reports the phenomenology of extra dimensional torsion \cite{Kalinowski:1980da}, metric dependent torsional models of extra dimensions \cite{Shankar:2012vd} and its consequences in cosmology \cite{Chen:2009ep}, compatification of higher dimensional Brans--Dicke models with torsion \cite{German:1993bq}. Torsion free solutions where found in \cite{Aros:2007nn} for models in first order compactified gravity.\\
Eventhough at the macroscopic scale, spacetime seems to lack torsion, there is no logical constraint to assume it non-existing, as E. Cartan argued, moreover since vacuum predictions of Einstein--Cartan theory hold the experimental tests of General Relativity.\\
This work is organized as follows: in Sec.~\ref{KK} we discuss the introduction of torsional degrees of freedom compatible with the Kaluza--Klein geometric reasoning, as well as fix our notation and conventions. Sec.~\ref{5EGB} studies the dynamics of the general Lovelock--Cartan action in five-dimensional spacetime and its dimensional reduction. Then, in Sec.~\ref{cosmos} we look for cosmological solutions in the reduced four-dimensional manifold of the Friedmann--Robertson--Walker class, with torsion both in the reduced manifold and as new dynamical fields arising from the compact section taken to be $S^1$. Conslusions and remarks are given next in Sec.~\ref{conclusions}. We also found appropriate to incorporate a number of appendices to make this work a self contained article.
%_-----------------------------------------------------------------------------------------

\section{Kaluza--Klein Ansatz in Riemann--Cartan Spacetimes\label{KK}}
Let $M_N$ be a $N$ dimensional differential manifold. Every quantity defined on $M_N$ will wear hats $\hat{x}$. Capital Greek characters (coordinate indices) and capital Latin characters (Lorentz indices) run along the $N$ dimensions, while lower case run in the $(N-1)$ dimensional reduced manifold.\\
Let $\hat{\omega}^{AB}$ be the spin connection 1-form that describes the affine structure of the manifold and $\hat{e}^A=\hat{e}^{A}_{\ \Gamma}\,\text{d}\hat{x}^\Gamma$ be the vielbein 1-form that defines the metric structure of the same manifold through the relation $\hat{g}_{\Gamma\Delta}=\hat{e}^{A}_{\ \Gamma}\hat{e}^{B}_{\ \Delta}\hat{\eta}_{AB}$. The vielbein is the mapping that relates coordinate indices (Greek characters) with the Lorentz indices (Latin characters). In our convention the flat Lorentzian metric will be $\hat{\eta}_{AB}=\,$diag$\,(-,+,...,+)$, and the $N$ dimensional Levi-Civita symbol $\hat{\epsilon}_{A_1...A_{N-1}}$ is reduced by defining
\begin{align*}
\epsilon_{a_1...a_{N-1}}\equiv\hat{\epsilon}_{a_1...a_{N-1}N}.
\end{align*} 
Our convention fixes $\hat{\epsilon}_{01...N}=1$.\\
Riemannian fields (torsion free) will also be explicited by a tilde, as it was done in Eq.~\eqref{GB} and Eq.~\eqref{GB2} for the curvature.\\
Due to the topology of the manifold, we can expand the depenence of the fields on the extra coordinate, $z$, in Fourier series as  
\begin{align}\label{Fourier}
\varrho(\hat{x}^\Gamma)&=\sum_n\varrho_{(n)}(x^\gamma)e^{in z}.
\end{align}
For the rest of the paper, we will focus in the $n=0$ mode of the expasion, also refered to as the low energy sector.
\subsection{Metric and affine structure}
The Kaluza-Klein (KK) ansatz for the metric lies on the premise that the compact section of $M_N$ is orthogonal to the rest of the manifold at each point. This leads to the conclusion that the metric has the following structure
\begin{equation}
  \hat{g}_{\Gamma\Delta} =
  \begin{pmatrix}
    g_{\gamma\delta} +\frac{\hat{g}_{\gamma z}\hat{g}_{\delta z}}{\hat{g}_{zz}}&\hat{g}_{\gamma z}\\
    \hat{g}_{z\delta} & \hat{g}_{zz}
  \end{pmatrix}
  =
  \begin{pmatrix}
    g_{\gamma\delta} + \phi A_\gamma A_\delta&\phi A_\gamma\\
    \phi A_{\delta} & \phi
  \end{pmatrix},
\end{equation}
and its inverse
\begin{equation}
  \hat{g}^{\Gamma\Delta}=
  \begin{pmatrix}
    g^{\gamma\delta}&-A^\gamma\\
    -A^{\delta} & \phi^{-1}+A^2
  \end{pmatrix}.
\end{equation}
It introduces a scalar field $\phi$ and a vector field $A_\mu$ as new gravitational degrees of freedom materialized in the reduced manifold.
The vielbein that keeps this structure for the metric has the following form
\begin{equation}
  \label{Dvielbein}
  \hat{e}^A_{\ \ \Gamma} =
  \begin{pmatrix}
    \hat{e}^a_{\ \gamma}& 0\\
    \hat{e}^N_{\ \gamma} & \hat{e}^N_{\ \ z}
  \end{pmatrix}
  =
  \begin{pmatrix}
    e^a_{\ \gamma}& 0\\
    \sqrt{\phi}A_\gamma & \sqrt{\phi}
  \end{pmatrix},
\end{equation}
and its inverse, $\hat{E}_A^{\ \ \Gamma}$ defined such that $\hat{E}_A^{\ \ \Gamma}\hat{e}^A_{\ \ \Delta}=\delta^\Gamma_{\Delta}$, read
\begin{equation}
  \label{Dinversevielbein}
  \hat{E}_A^{\ \ \Gamma} =
  \begin{pmatrix}
    \hat{E}_a^{\ \gamma}& 0\\
    \hat{E}_a^{\ z} & \hat{E}_N^{\ \ z}
  \end{pmatrix}
  =
  \begin{pmatrix}
    E_a^{\ \gamma}& 0\\
    -A_a & \sqrt{\phi}^{-1}
  \end{pmatrix},
\end{equation}
and it shows that the compact section $S^1$ has its own tangent space $T_pS^1$ at each point $p\in M_N$, and is independent to $T_pM_{N-1}$ in the sense that they do not mix. It seems reasonable to think that if we want two manifolds to be orthogonal to one another, they will not mix vectors after the tangent mapping.\\
The basis of vielbeins is defined modulo Lorentz transformations. Any transformed basis 
$
\hat{e}^{A\prime}=\hat{\Lambda}^A_{\ B}\hat{e}^B,
$
where $\hat{\Lambda}$ is a Lorentz matrix, is as suitable as $\hat{e}^A$, and therefore share the structure in Eq.~\eqref{Dvielbein}. The Lorentz transformations are then reduced by this fact, imposing over the Loretnz matrix 
$\hat{\Lambda}^a_{\ N}=0.
$

%_-----------------------------------------------------------------------------------------

A Riemannian connection compatible with a $N$ dimensional vielbein, Eq.~\eqref{Dvielbein}, is built under the premise that $\mbox{d}\hat{e}^A + \hat{\tilde{\omega}}^{AB} \wedge \hat{e}^A = 0$. Thus, we find
\begin{align}
  \label{DRiemannconnection}
  \hat{\tilde{\omega}}^{ab}&=\tilde{\omega}^{ab}-\frac{1}{2}\sqrt{\phi}F^{ab}\hat{e}^N,\\
  \hat{\tilde{\omega}}^{Na}&=\frac{1}{2}\sqrt{\phi}F^a_{\ \ l}e^l+\frac{1}{2}\partial^a\ln\phi\hat{e}^N,
\end{align}
where $\tilde{\omega}^{ab}$ is the Riemannian spin connection of the reduced manifold. The construction of a theory for a spacetime with spin connection \eqref{DRiemannconnection} leads to the original KK theory. It gives rise the usual $U(1)$-invairant field strength, 
\begin{align*}
F=\text{d}A=\frac{1}{2}F_{ab}\, e^a\wedge e^b,
\end{align*} 
non-minimally coupled to the scalar field.\\
Nevertheless, $M_N$ is in this case, not entirely described by $\hat{\tilde{\omega}}^{AB}$, but by a more general spin connection, independent of the metric degrees of freedom (see Appendix \ref{Riemann-Cartan} for details). We can assume a general connection 1-form of the same type (regarding its $M_{N-1}\times S^1$ decomposition) as in eq.~\eqref{DRiemannconnection}. Thus
the most general spin connection on $M_N$ compatible with the KK decomposition is given by 
\begin{equation}\label{Nconnection}
  \hat{\omega}^{AB} \equiv
  \begin{pmatrix}
    \omega^{ab}+\alpha^{ab}\hat{e}^N&\beta^a+\gamma^a\hat{e}^N\\
    -\beta^b-\gamma^b\hat{e}^N&0
  \end{pmatrix}.
\end{equation}
It contemplates the torsional degrees of freedom appart of the usual metric ones. 
The decomposition \eqref{Nconnection} add new independent fields into the picture. Here $\alpha^{ab}$ is a $0$-form, $\beta^a$ a vector valued $1$-form and $\gamma^a$ a $0$-form that admit the Fourier expansion \eqref{Fourier} regarding their dependence on the extra dimension. Here, as we pointed out earlier, we will take the $0$-mode of such an expansion.

%_-----------------------------------------------------------------------------------------

\subsection{Curvature and torsion}
Using the definition of curvature
$\hat{R}^{AB}=\mbox{d}\hat{\omega}^{AB}+\hat{\omega}^A_{\ \ C}\wedge\hat{\omega}^{CB}$,
we find
\begin{align}\label{R1}
  \hat{R}^{ab}&=R^{ab}+\sqrt{\phi}\alpha^{ab}F-\beta^a\wedge\beta^b\notag\\
  &+\left(\mbox{D}\alpha^{ab}+\frac{1}{2}\alpha^{ab}\mbox{d}\ln\phi-2\beta^{[a}\gamma^{b]}\right)\wedge\hat{e}^N,\\
  \label{R2}
  \hat{R}^{aN}&=\left(\mbox{D}\beta^a+\sqrt{\phi}\gamma^a F\right)\notag\\
  &-\left(\alpha^a_{\ b}\beta^b-\mbox{D}\gamma^a-\frac{1}{2}\gamma^a\mbox{d}\ln\phi\right)\wedge\hat{e}^N.
\end{align}
where we have used the fact that
$
  \mbox{d}\hat{e}^N=\frac{1}{2}\mbox{d}\ln\phi\wedge\hat{e}^N+\sqrt{\phi}F.
$
Similarly from the definition of torsion
$\hat{T}^A = \hat{\mbox{D}}\hat{e}^A=\mbox{d}\hat{e}^A+\hat{\omega}^A_{\ \ B}\wedge\hat{e}^B$
we find its distinctive parts to be
\begin{align}\label{T1}
  \hat{T}^a &= T^a+\left(\beta^a-\alpha^a_{\ \ b}e^b\right)\wedge\hat{e}^N,\\
  \label{T2}
  \hat{T}^N &= \sqrt{\phi}F-\beta_b\wedge e^b+\left(\frac{1}{2}\mbox{d}\ln\phi+\gamma_be^b\right)\wedge\hat{e}^N.
\end{align}

\subsubsection{Bianchi identities}

Considering the Bianchi identities of the KK structure described above, we find relevant information about the new fields. Taking the exterior covariant derivative of the field strengths
$\hat{R}$ and $\hat{T}$, the Bianchi identities are
\begin{align*}
  \hat{\text{D}}\hat{R}^{AB} &= 0, & \hat{\text{D}}\hat{T}^A &= \hat{R}^A_{\ \ B}\wedge\hat{e}^B.
\end{align*}
A careful decomposition of the first one in its distinctive parts, gives the Bianchi identity for the curvature of the reduced spacetime together with the second derivative rules
\begin{align*}
  \text{D}R^{ab} &=0, & \text{D}\mbox{D}\alpha^{ab} &=R^a_{\ \ l}\alpha^{lb}+R^b_{\ \ l}\alpha^{al},\\
  \text{D}\text{D}\beta^a &= R^a_{\ \ b}\beta^{b}, & \text{D}\text{D}\gamma^a &= R^a_{\ \ b}\gamma^{b}.
\end{align*}
This is taken as proof of the tensorial condition of these new fields. While consideration of the second one gives nothing but its equivalent for the manifold $M_{N-1}$, this is
\begin{equation*}
  \mbox{D}T^a=R^a_{\ \ b}\wedge e^b.
\end{equation*}

%_-----------------------------------------------------------------------------------------

\section{Five Dimensional Einstein--Gauss--Bonnet Reduction\label{5EGB}}
We will work in this section, the dimensional reduction of a gravitational Lagrangian that includes quadratic terms in the curvature, namely the Gauss--Bonnet term. The action is given by
\begin{widetext}
  \begin{equation}\label{action5EGB}
    I=\int\limits_{M_5}\hat{\epsilon}_{ABCDE}\Big(\frac{\alpha_0}{5}\hat{e}^A\wedge\hat{e}^B\wedge\hat{e}^C\wedge
    \hat{e}^D\wedge\hat{e}^E
    +\frac{\alpha_1}{3}\hat{R}^{AB}\wedge\hat{e}^C\wedge\hat{e}^D\wedge\hat{e}^E
    +\alpha_2\hat{R}^{AB}\wedge\hat{R}^{CD}
    \wedge\hat{e}^E\Big),
  \end{equation}
where $\alpha_0$, $\alpha_1$ and $\alpha_2$ are coupling constants. Variations with respect to the vielbein and spin connection give the equations
  \begin{align}\label{delta e}
    \hat{\epsilon}_{ABCDE}\Big(\alpha_0\hat{e}^A\wedge\hat{e}^B\wedge\hat{e}^C\wedge\hat{e}^D
    +\alpha_1\hat{R}^{AB}\wedge\hat{e}^C\wedge\hat{e}^D
    +\alpha_2\hat{R}^{AB}\wedge\hat{R}^{CD}\Big)&=0,
    \\\label{delta w}
    \hat{\epsilon}_{ABCDE}\Big(\alpha_1\hat{e}^C\wedge\hat{e}^D+
    2\alpha_2\hat{R}^{CD}\Big)\wedge\hat{T}^E&=0,
  \end{align}
  respectively. 
\end{widetext}
In this work, we will focus on the general case when
\begin{equation}\label{delta}
\Delta\equiv\alpha_1^2-4\alpha_0\alpha_2\neq 0.
\end{equation}
Otherwise, the Lagrangian becomes a gauge invariant Lagrangian for a larger group, namely the $AdS_5$ group, and is recognized as its Chern--Simons form \cite{Zanelli:2005sa}, \cite{Troncoso:1999pk}. In general, Lovelock--Cartan (LC) theory allows the torsional term
\begin{align*}
\mathcal{L}_{T}\sim\hat{R}_{AB}\wedge\hat{T}^A\wedge\hat{e}^B,
\end{align*}
that in five dimensions is a boundary term proportional to the Nieh-Yan invariant, thus it does not participate in the dynamics. Action \eqref{action5EGB} is the general LC theory in five dimensions.\\
In terms of the KK ansatz presented in previous sections, the equations of motion can be decomposed in to its distinctive parts. Eq.~\eqref{delta e} leads to the following set
\begin{align}\label{equation}
\epsilon_{abcd}\Big[\alpha_0 e^a\wedge e^b\wedge e^c+\frac{1}{2}\alpha_1\left(M^{ab}-L^a\wedge e^b\right)\wedge e^c&\notag\\
-\alpha_2\left(L^a\wedge M^{bc}+K^a\wedge N^{bc}\right)\Big]&=0,\\
\epsilon_{abcd}K^a\wedge\left(\alpha_1e^b\wedge e^c+2\alpha_2M^{bc}\right)&=0,\\
\epsilon_{abcd}\big(\alpha_0 e^a\wedge e^b\wedge e^c\wedge e^d&\notag\\
+\alpha_1M^{ab}\wedge\wedge e^c\wedge e^d+\alpha_2M^{ab}\wedge M^{cd}\big)&=0,\\
\epsilon_{abcd}\left(\alpha_1N^{ab}\wedge e^c\wedge e^d+2\alpha_2N^{ab}M^{cd}\right)&=0,
\end{align}
while Eq.~\eqref{delta w} gives
\begin{align}
\epsilon_{abcd}\Big[\alpha_1\left(e^c\wedge e^d\wedge Z-2e^c\wedge T^d\right)
+2\alpha_2\Big(N^{cd}\wedge W&\notag\\
+M^{cd}\wedge Z+2L^c\wedge T^d
+2K^c\wedge V^d\Big)\Big]&=0,\\
\epsilon_{abcd}\left[\alpha_1e^c\wedge e^ d\wedge W+2\alpha_2\left(M^{cd}\wedge W+2K^c\wedge T^d\right)\right]&=0,\\
\epsilon_{abcd}\left(\alpha_1e^b\wedge e^c+2\alpha_2M^{bc}\right)\wedge T^d&=0,\\
\epsilon_{abcd}\left[\alpha_1e^b\wedge e^c V^d+2\alpha_2\Big(M^{bc}\wedge V^d+N^{bc}\wedge T^d\Big)\right]&=0.
\end{align}
The fields $M^{ab}$, $N^{ab}$, $L^a$, $K^a$, $W$, $V^a$ and $Z$ are defined through \eqref{R1}, \eqref{R2}, \eqref{T1} and \eqref{T2} such that
\begin{align*}
\hat{R}^{ab}&=M^{ab}+N^{ab}\wedge\hat{e}^5, & \hat{R}^{5a}&=K^a+L^a\wedge\hat{e}^5,\\
\hat{T}^a&=T^a+V^a\wedge\hat{e}^5, & \hat{T}^5&=W+Z\wedge\hat{e}^5.
\end{align*}

%_-----------------------------------------------------------------------------------------
\section{Dimensionally Reduced EGB Cosmology}\label{cosmos}
\subsection{Cosmological ansatz}
We will look for cosmological solutions of the equations of motion. For that, we demand the symmetries assumed by the cosmological principle, namely, isotropy and homogeneity of the fields involve. This is achieved by making the Lie derivative in the direction of the Killing vectors of the symmetries to vanish for each field. Appendix \ref{homotropic} is devoted to the details of how to reach the general ansatz that we present next. The vielbein compatible with the Friedman-Robertson-Walker metric is determined by one function called scale factor $a(t)$,
\begin{align}
  \label{vielbein cosmo}
  e^0&=\mbox{d}t, & e^1&=\frac{a(t)}{\sqrt{1-kr^2}}\mbox{d}r,\\
  e^2&=a(t)r\mbox{d}\theta, & e^3&=a(t)r\sin\varphi\mbox{d}\varphi, 
\end{align}
where $k=+1,0,-1$ determines if the spatial section is closed, flat or an open $3$-space respectivelly. The scalar field is a time-dependent function
\begin{equation}
  \phi=\phi(t).
\end{equation}
On the other hand, the spin connection adds two functions we called $\omega(t)$ and $f(t)$,
\begin{align}
  \omega^{0i}&=\omega(t) e^i,\\
  \omega^{12}&=-\frac{\sqrt{1-kr^2}}{a(t)r}e^2-f(t)e^3,\\
  \omega^{13}&=-\frac{\sqrt{1-kr^2}}{a(t)r}e^3+f(t)e^2,\\
  \omega^{23}&=-\frac{\cot\theta}{a(t)r}e^3-f(t)e^1.
\end{align}
The non-vanishing components of the spin connection induced by the compact section of the manifold are
\begin{align}
  \beta^0=-b(t)e^0,&\ \beta^i=\beta(t)e^i,\\
  \label{gamma cosmo}
  \gamma^0&=-\gamma(t).
\end{align}
For this ansatz, the $1$-form $A$ has the form $A_t(t)\text{d}t$, which can be shown to have vanishing field strength $F$. Thus without loss of generality, it can be set to zero by means of a $U(1)$ gauge transformation, or equivalently, a diffeomorphism transformation along the $z$-direction.\\
The equations of motion for the cosmological ansatz gives a system of differential equations for the time dependent functions defined above. Curvature and torsion for this ansatz can be seen in Appendix \ref{homotropic}. We found that the only non-Riemannian branch demands  $\beta^a=0$. Thus $b(t)=\beta(t)=0$. Otherwise, the system degenerates at the CS point. The final set of equations is the following:
\begin{align}
  \label{eqn1}
  h\left[\alpha_1+2\alpha_2\left(\omega^2+\frac{k}{a^2}-f^2\right)\right]+4\alpha_2f\left(\dot{f}+Hf\right)&=0,\\
  2\alpha_0+\alpha_1\left(\dot{\omega}+2\omega^2+\frac{k}{a^2}-f^2\right)\notag\\
  +2\alpha_2\left(\dot{\omega}+\omega^2\right)\left(\omega^2+\frac{k}{a^2}-f^2\right)&=0,\\
  2\alpha_2\left(\frac{1}{2}\dot{\Phi}+\gamma\right)f\left(\dot{f}+Hf\right)
  +h^2\left(\alpha_1-2\alpha_2\omega\gamma\right)&=0,\\
  \omega\left(\frac{1}{2}\dot{\Phi}+\gamma\right)+\dot{\gamma}+\frac{1}{2}\gamma\dot{\Phi}-\frac{\alpha_1}{2\alpha_2}&=0,\\
  \left(\omega^2+\frac{k}{a^2}-f^2\right)\left(\alpha_1-2\alpha_2\omega\gamma\right)-\alpha_1\omega\gamma+2\alpha_0&=0,\\
  \label{eqn2}
  \left(\dot{\omega}+\omega^2\right)\left(\alpha_1-2\alpha_2\omega\gamma\right)-\alpha_1
  \omega\gamma+3\alpha_0-\frac{\alpha^2_1}{4\alpha_2}&=0,
\end{align}
where dotted functions stands for its time-derivatives. We have defined for simplicity the functions $h(t)=\omega(t)-H(t)$, where $H=\dot{a}/a$ is the Hubble function, and $\Phi=\ln\phi$.

\subsection{Solutions}

The system is integrated developing two non-Riemannian branches, one for each value of the parameter $u_\pm$ defenied to be
\begin{equation}\label{u}
  u_\pm=\frac{2\alpha_1\pm\sqrt{6\Delta}}{4\alpha_2},
\end{equation}
where $\Delta$ was given in Eq.~\eqref{delta}. Eqs.~\eqref{eqn1} to~\eqref{eqn2} reduced to the Riemannian system when \mbox{$f=h=0$} and \mbox{$\gamma=-\dot{\Phi}/2$}. The details of such a model can be seen in Refs.~\cite{Deruelle:1986iv,Deruelle:2003ck,Henriques:1986jw,Ishihara:1986if,Kleidis:1997mu}.

Due to Eq.~\eqref{u}, the solutions will be valid in the region of parameter space where $\Delta>0$. The function $\omega(t)$ satisfies the equation $\dot{\omega}+\omega^2+u_{\pm}=0$, that for the three significantly different values of $u_\pm$, has the following solutions
\begin{align}
  u_\pm>0&:\omega(t)=-\sqrt{u_\pm}\tan\left[\sqrt{u_\pm}\left(t-t_0\right)\right],\\
  u_\pm=0&:\omega(t)=\left(t-t_0\right)^{-1},\\
  u_\pm<0&:\omega(t)=\sqrt{-u_\pm}\tanh\left[\sqrt{-u_\pm}\left(t-t_0\right)\right],
\end{align}
where $t_0$ is an integration constant to be fixed. We relegate the time dependence of the remaining fields to these expressions for $\omega$ and list their explicit time dependence in Appendix~\ref{solutions t}. The solutions read
\begin{align}
  \gamma(\omega)&=\frac{u_\pm}{\omega},\\
  \phi(\omega)&=\frac{\phi_0\,\omega^2}{\omega^2+u_\pm}\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2\left(\omega^2+u_\pm\right)}\right],\\
  a(\omega)&=\frac{a_0}{\sqrt{\omega^2+u_\pm}}\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2\left(\omega^2+u_\pm\right)}\right],\\
  f^2(\omega)&=\left(\omega^2+u_\pm\right)\frac{k}{a_0^2}\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2\left(\omega^2+u_\pm\right)}\right]\notag\\
  &+\omega^2-Y_\pm,
\end{align}
where we have defined the constant
\begin{equation*}
  Y_{\pm}=\mp2\frac{\alpha_1 u_\pm-2\alpha_0}{\sqrt{6\Delta}}=-\frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2},
\end{equation*}
and where $a_0$ and $\phi_0$ are integration constants.

\subsubsection{Bouncing universes}
For $u_\pm>0$ the solutions are periodic. In that case, the scale factor starts from a singular point and reach a future one after a time $t_{\text{crunch}}=\pi/\sqrt{u_\pm}$. Depending on the particular sector of parameter space, the solution undergoes an expanding and contracting age that allow it to reach a minimum value without collapsing, to expand again before a big crunch. Otherwise, it reaches a maximum value before collapse. It also shows an inflationary phase at early times when the scale factor grows exponentially.

\subsubsection{Expanding and contracting universes}
For $u_\pm\leq 0$, depending on the sign of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will infinitely expand while the scalar field reach an asymtotic vanishing value for late cosmic times, after a period of expansion. This is interpreted as the late cosmic time behaviour of the radius of the compact spacial section $S^1$, that for this familly of solutions provides a natural explanation for its scale. Otherwise, in contrast to the previous behaviour, the scale factor contracts to an asymtotic vanishing value while the scalar field has no upper bound. In the case of strictly negative $u_\pm$, for either contracting or expanding scale factor, its initial value remains finite, thus the solutions do not present an original singularity. It has been reported that the GB term in a gravitational action can prevent the universe to expand from an initial singularity \cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if}, which is also the case here.

\subsubsection{Effective energy density and pressure}
Eq.~\eqref{equation} take the familiar form
\begin{align}\label{einstein equation}
\frac{1}{2}\epsilon_{abcd}\left(\alpha_0e^a\wedge e^b\wedge e^c+\frac{1}{2}\alpha_1\tilde{R}^{ab}\wedge e^c\right)&=\tau^{eff}_d,
\end{align}
when we use \eqref{curvature decomp} and define an effective energy-momentum form as
\begin{align}
\tau_a^{eff}&=\frac{1}{2}\epsilon_{abcd}\bigg[\frac{1}{2}\alpha_1\left(\kappa^b_{\ l}\wedge\kappa^{lc}+\tilde{\text{D}}\kappa^{bc}\right)\wedge e^d\notag\\
&+\left(\text{D}\gamma^b+\frac{1}{2}\gamma^b\text{d}\ln\phi\right)\wedge\left(\frac{1}{2}\alpha_1e^c\wedge e^d+\alpha_2R^{cd}\right)\bigg].
\end{align}
The contributions of $F$, $\alpha^{ab}$ and $\beta^a$ have not been taken into account since for all practical purposes of this article, they vanish. The left hand side of \eqref{einstein equation} is Eintein's $3$-form, that in components gives Einstein's tensor. We identify the right hand side as an effective energy-momentum form that contemplates the torsional functions and the higher dimensional metric degrees of freedom as part of the gravitating matter of a four dimensional Universe in the form of a perfect fluid. For this energy-momentum form, we identify the energy density $\rho$ and pressure $p$ by
\begin{align*}
\tau_0^{eff}&=-\frac{1}{3!}\rho\epsilon_{0ijk}e^i\wedge e^j\wedge e^k,\\
\tau_i^{eff}&=-\frac{1}{2}p\epsilon_{0ijk}e^0\wedge e^j\wedge e^k.
\end{align*} 
Hence
\begin{align}\label{rho}
\rho&=-\frac{3}{2}\alpha_1\left(h^2-f^2+2Hh\right)\notag\\
&+3\omega\gamma\left[\frac{1}{2}\alpha_1+\alpha_2\left(\omega^2+\frac{k}{a^2}-f^2\right)\right],\\
\label{presion}
p&=\alpha_1\left[\dot{h}+Hh+\frac{1}{2}\left(h^2-f^2+2Hh\right)\right]\notag\\
&-2\omega\gamma\left[\frac{1}{2}\alpha_1+\alpha_2\left(\dot{\omega}+H\omega\right)\right]\notag\\
&+\left(\dot{\gamma}+\frac{1}{2}\gamma\dot{\Phi}\right)\left[\frac{1}{2}\alpha_1+\alpha_2\left(\omega^2+\frac{k}{a^2}-f^2\right)\right].
\end{align}
The identification satisfy $\dot{\rho}+3H\left(\rho+p\right)=0$. Using equations of motion, one can show that
\begin{align*}
\rho\sim H^2+\frac{k}{a^2}+\text{const.}
\end{align*}
For $u_\pm<0$, this function remains finite at the begining.
From expressions \eqref{rho} and \eqref{presion} one can see that the induced energy density and pressure are not positive definite quantities. In fact, due to the presence of torsion and extra dimensional fields, the universe solutions undergoes through accelerating expansive age with no need of extra fluids of exotic class.

%_-----------------------------------------------------------------------------------------

\section{Discussion and Conclusions}\label{conclusions}
It is known that KK theory, in its simplest version, give rise matter fields a manifestation of extra dimensional metric degrees of freedom. The kind of matter that arises in the reduced spacetime is subordinated to the topology taken for the extra compact manifold. In general, those fields will gravitate and modulate the behaviour of spacetime in its time evolution. We have constructed the simplest model in five dimensional spacetime that allow the presence of torsion in the KK decomposition. In the cosmological model, the induced matter present due to an extra compact section taken to be $S^1$ are the usual scalar field --a metric degree of freedom-- and a vector field that comes from the dimensional reduction of a general spin connection. Torsion exists both in the reduced manifold and in the higher dimensional and indeed, it participate in the effective energy-momentum form that gravitates as a perfect fluid. Since the sign of the energy density nor of the presures in the fluid are in no way guaranteed, it is interpreted as a class of matter of exotic characteristics. Exact solution for a five dimensional vacuum are studied. Accelerating, and bouncing and inflationary universes are possible by the equations of motion.\\
There are periods of time where $f^2(t)$ becomes negative. As we can see from Eq.~\eqref{homotorsion}, $f$ corresponds to the completely antisymmetric part of torsion and is unseen by classical particles following geodesics along spacetime. It does not couple to spin-0 or spin-1 bosons, but it does appear as an effective interaction term in the Dirac Lagrangian writen in a curved spacetime \cite{Carroll:1994dq}
\begin{align*}
\mathcal{L}_{\text{Interaction}}&=-\frac{i}{8}T_{\alpha\mu\nu}\bar{\Psi}\Gamma^{\alpha\mu\nu}\Psi\\
&=\frac{3}{2}f\bar{\Psi}\Gamma^0\Gamma^5\Psi.
\end{align*}
It couples to the chiral current. For imaginary values of $f(t)$, the corresponding violation of current conservation can be interpreted as particle creation. Similar results where found in \cite{Toloza:2013wi} for another cosmological model with torsion that share some of the couplings of a reduced action. 
\begin{acknowledgments}
The authors would like to thank O. Miskovic, A. Toloza and J. Zanelli for enlightening discussions in this work.
\end{acknowledgments}
%_-----------------------------------------------------------------------------------------

\appendix
\section{Riemann-Cartan Geometry}\label{Riemann-Cartan}
In Riemann--Cartan geometry, both the vielbein $e^A$ and the spin connection $\omega^{AB}$ are independent fields. We will drop the hats since we will not deal with the KK decomposition. Stationary variations of a gravity actions must be taken with respect to both fields, allowing in general, the presence of torsion. The spin connection, however, can be decomposed in a Riemannian part $\tilde{\omega}^{AB}$ that satisfies $\tilde{\mbox{D}}e^A=0$, and a contorsion $\kappa^{AB}=-\kappa^{BA}$, such that 
\begin{equation}\label{spin separation}
  \omega^{AB}=\tilde{\omega}^{AB}+\kappa^{AB}.
\end{equation} 
Therefore, the torsion 2-form defined as the covariant derivative of the vielbein with respect to the total spin connection, is
\begin{equation}
  T^A=\kappa^A_{\ B}\wedge e^B.
\end{equation}
On the other hand, the curvature 2-form also suffers corrections with respect to the Riemannian one due to the presence of torsional degrees of freedom. This can be seen explictly by taking the definition of curvature $R^{AB}=\mbox{d}\omega^{AB}+\omega^a_{\ C}\wedge\omega^{CB}$ and \eqref{spin separation} to find
\begin{equation}\label{curvature decomp}
  R^{AB}=\tilde{R}^{AB}+\kappa^A_{\ C}\wedge\kappa^{CB}+\tilde{\mbox{D}}\kappa^{AB},
\end{equation}
where $\tilde{R}^{AB}$ is the Riemannian curvature built with $\tilde{\omega}^{AB}$.

%_-----------------------------------------------------------------------------------------

\section{Isotropic--Homogeneus Ansatz}\label{homotropic}
The cosmological principle demands the spacial section of spacetime to be isotropic and homogeneus. This means that the fields involve in the model must be compatible with this assumption. A spacetime is an isotropic spacetime with respect to certain point $P$ if after a rotation with respect to an axis that pass through $P$, all the geometrical properties remain invariant, thus the spacetime looks the same in all directions. By homogeneity we understand that spacetime looks the same from every point $P$. These two assumptions are translated in the Killing equations for the fields, which are the vanishing of the Lie derivatives of the fields along the direction of the vector generators of symmetries $\{\zeta^\lambda_{i}\}$. In particular the Killing equations for the metric tensor and the torsion tensor are
\begin{align}
  \label{killing metric}
  \text{\textsterling}_i g_{\mu\nu}&=\zeta^\lambda_i\partial_\lambda g_{\mu\nu}+\partial_\mu\zeta^\lambda_i g_{\lambda\nu}+\partial_\nu\zeta^\lambda_i g_{\mu\lambda}=0,\\
  \text{\textsterling}_i T^\alpha_{\ \mu\nu}&=\zeta^\lambda_i\partial_\lambda T^\alpha_{\ \mu\nu}-\zeta^\alpha_i\partial_\lambda T^\lambda_{\ \mu\nu}
  +\zeta^\lambda_i\partial_\mu T^\alpha_{\ \lambda\nu}\notag\\
  &\quad +\zeta^\lambda_i\partial_\nu T^\alpha_{\ \mu\lambda}=0,
\end{align}
where $T^\alpha_{\ \mu\nu}$ are the components of the torsion 2-form defined by $T^a=\frac{1}{2}e^a_{\ \alpha}T^\alpha_{\ \mu\nu}\mbox{d}x^\mu\wedge\mbox{d}x^\nu$. The same must hold for the new fields
\begin{align}
  \label{killing vector}
  \text{\textsterling}_i A_\mu&=\zeta^\lambda_i\partial_\lambda A_\mu+\zeta^\lambda_i\partial_\mu A_\lambda=0,\\
  \text{\textsterling}_i \phi &=\zeta^\lambda_i\partial_\lambda\phi=0,
\end{align}
and also for the components (tensors) $\alpha_{\mu\nu}=-\alpha_{\nu\mu}$, $\beta_{\mu\nu}$ and $\gamma_\mu$, defined such that
\begin{align*}
  \alpha^{ab}&=E^{a\mu}E^{b\nu}\alpha_{\mu\nu},\\
  \beta^a&=E^{a\mu}\beta_{\mu\nu}\mbox{d}x^\nu,\\
  \gamma^a&=E^{a\mu}\gamma_\mu,
\end{align*} 
and whose Killing equations are analogous to the tensorial one, \eqref{killing metric} and vectorial one,\eqref{killing vector}.\\
The set of Killing vectors $\{\zeta^\lambda_i\}$ are, on the one hand, the generators of spatial rotations in three dimensions $SO(3)$, $\xi_i=\epsilon_{ijk}x_j\partial_k$, and the Killing vectors assosiated with spacial translations $\mathcal{P}_i=\sqrt{1-kr^2}\partial_i$, that satisfy the algebra
\begin{align}
  \left[\xi_i,\xi_j\right]&=\epsilon_{ijk}\xi_k,\\
  \left[\mathcal{P}_i,\mathcal{P}_j\right]&=-k\epsilon_{ijk}\xi_k,\\
  \left[\xi_i,\mathcal{P}_j\right]&=\epsilon_{ijk}\xi_k.
\end{align}
These requirements on the fields are translated to a set of first order differential equations whose most general solutions determines our ansatz structure \eqref{vielbein cosmo} to \eqref{gamma cosmo}. The fields strengths for isotropic--homogeneus vielbein and spin connection are written next. The curvature components are
\begin{align}
  R^{0i}&=\left(\dot{\omega}+H\omega\right)e^0\wedge e^i+f\omega\epsilon^i_{\ jk}e^j\wedge e^k,\\
  R^{ij}&=\left(\omega^2+\frac{k}{a^2}-f^2\right)e^i\wedge e^j
  -\left(\dot{f}+Hf\right)\epsilon^{ij}_{\ \ k}e^0\wedge e^k,
\end{align}
and the torsion components
\begin{equation}\label{homotorsion}
  T^0=0,\quad T^i=-he^0\wedge e^i+f\epsilon^i_{\ jk}e^j\wedge e^k,
\end{equation}
where $h(t)$ and $H(t)$ were defined in Section \ref{5EGB}.

%_-----------------------------------------------------------------------------------------
\section{Time Dependence of the Solutions}\label{solutions t}
We supply here the time dependent expressions of the cosmological solutions for the significantly differents values of $u_\pm$ in terms of the dimensionless parameter $\tau=\sqrt{|u_\pm|}(t-t_0)$.\\
\newline
$\bullet\ u_\pm>0$:
\begin{align}
  \gamma(t)&=-\sqrt{u_\pm}\cot\tau,\\
  \phi(t)&=\phi_0\sin^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2}\cos^2\tau\right],\\
  a(t)&=\frac{a_0}{\sqrt{u_\pm}}|\cos\tau|\exp\left[\frac{\pm\sqrt{6\Delta}}{24\alpha_2}\cos^2\tau\right],\\
  f^2(t)&=u_\pm\frac{k}{a_0^2}\sec^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cos^2\tau\right]\notag\\
  &+u_\pm\tan^2\tau-Y_\pm.
\end{align}
\\
\medskip
$\bullet\ u_\pm=0$:
\begin{align}
  \gamma(t)&=0,\\
  \phi(t)&=\phi_0\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2}(t-t_0)^2\right],\\
  a(t)&=a_0|t-t_0|\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2}(t-t_0)^2\right],\\
  f^2(t)&=\frac{k}{a_0^2(t-t_0)^2}\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2}(t-t_0)^2\right]\notag\\
  &+(t-t_0)^{-2}-Y_\pm.
\end{align}
\\
\medskip
$\bullet\ u_\pm<0$:
\begin{align}
  \gamma(t)&=-\sqrt{-u_\pm}\coth\tau,\\
  \phi(t)&=\phi_0\sinh^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2u_\pm}\cosh^2\tau\right],\\
  a(t)&=\frac{a_0}{\sqrt{-u_\pm}}\cosh\tau\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2u_\pm}\cosh^2\tau\right],\\
  f^2(t)&=-u_\pm\frac{k}{a_0^2}\cosh^{-2}\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cosh^2\tau\right]\notag\\
  &-u_\pm\tanh^2\tau-Y_\pm.
\end{align}
%%%%%%%%% BIBLIOGRAPHY %%%%%%%%%
%%\bibliographystyle{apsrev4-1}
\bibliography{References.bib}

\end{document}

### 2016-09-10 KKosmology.tex
Version by Criz

@@
-\documentclass[aps,prd,12pt,superscriptaddress,showpacs,showkeys,longbibliography,reprint]{revtex4-1}
+\documentclass[aps,prd,12pt,superscriptaddress,showpacs,showkeys,longbibliography,reprint,nofootinbib]{revtex4-1}
 \usepackage{calrsfs} 
@@
 
-\title{Five Dimensional Kaluza--Klein Cosmology in Lovelock--Cartan Theory}
+\title{Five-Dimensional Kaluza--Klein Cosmology in Lovelock--Cartan Theory}
 
@@
 \author{Cristobal \surname{Corral}}
+\email{cristobal.corral@usm.cl}
 \affiliation{\UTFSM.}
@@
 \begin{abstract}
- We study the Kaluza-Klein dimensional reduction of an $N$ dimensional Riemann--Cartan manifold when the compact section is $S^1$.
-First, we focus on the details of dimensional reduction in the presence of torsion to then study the particular Einstein--Cartan theory with the Gauss--Bonnet term in five dimensions.
-We look for the general cosmological solution of the Friedman-Robertson-Walker familly in the four-manifold to interpret the remaining fields.
-The model, treated in five dimensional vacuum, presents torsion both in the reduced manifold as well as in the compact section, manifest as a non-vanishing four dimensional matter field that adds to the effective energy density and pressure.
-This allows the model to present expanding periods as well as contracting universes, and a radius of the compact section that it asymtotes zero in some solutions.
+ We study the Kaluza--Klein dimensional reduction of the Lovelock--Cartan theory on a $N$-dimensional Riemann--Cartan manifold, where the compact manifold has $S^1$ topology.
+First, we focus on the dimensional reduction in presence of torsion, to proceed with the Lovelock--Cartan theory in five dimensions.
+We find cosmological solutions of the Friedman--Robertson--Walker class for the dimensionally reduced theory.
+In the five-dimensional vacuum, this theory allows a nontrivial torsion in contrast to the Einstein--Cartan theory.
+After the dimensional reduction is performed, the torsional degrees of freedom induce a nonvanishing energy-momentum tensor in four dimensions, which behaves as a perfect fluid.
+Cosmological solutions describing accelerated expansion phases as well as contracting solutions are found.
+The model shows a dynamical compactification of the extra dimension in some regions of the parameter space.
 \end{abstract}
@@
 \section{Introduction}
-The geometric theory of gravity, General Relativity (GR), has been highly successful in describing the physics at large scales.
-Two major predictions were recently confirmed in one experiment that invoque two of the most fascinating conepts that this theory developed: the propagation of ripples of spacetime, gravitational waves, originated by the merge of two black holes where detected by the LIGO collaboration \cite{Abbott:2016blz} confirming the existence of both, astrophysical balck holes as well as the unprecedented nature of the signal detected.
-With this, we can rely in the geometric nature of this interaction.
-At the galactic scale, the gravitational lensing produced by the local distribution of energy and matter suggest the presence of a kind of matter that is not seen through telescopes.
-This abundance of matter in the galaxy is also compatible with the velocity profyle of stars at the outer regions of it and is needed to reach the current state of structures formation.
-This matter would interact mostly, if not only, through gravity \cite{Sofue:2000jx}.
-At the cosmic scale, observatinal evidence points out an accelerated current state of the Universe \cite{Riess:1998cb}.
-Such a behaviour needs the existence of a form of vacuum energy of exotic nature to give account of the accelerated expansion.
-This is evidence of unknown physics, and it has been the motivation for looking for new gravitational degrees of freedom beyond Einsten's GR.
-Among the familly of extended gravity theories, higher dimensional gravity can give us a clue to explore the manifestation of new degrees of freedom as material fields of certain class.
-Along this lines is that we consider here an extended model of gravity given by the so called Lovelock action~\cite{Lovelock:1971yv}.
-The Einstein--Hilbert action with cosmological constant is the unique action in four dimensions that yields second order field equations for the metric.
-In higher dimensions, however, this does not remains true.
-A particular combination of squared terms in the curvature can be added to the gravitational action to give rise a second order system for the metric with no ghosts.
-The particular quadratic term is called Gauss--Bonnet term, and in $N$ dimensions reads
+
+%The theory of General Relativity (GR) has been proven to be one of the most successful description of gravity nowadays.
+In this framework, the gravitational interaction is understood as the geometrical manifestation of the spacetime.
+The experimental status of General Relativity (GR), regarding the solar system tests~\cite{Will:2014kxa} and the detection of gravitational waves by the LIGO collaboration~\cite{Abbott:2016blz}, have settled GR as one of the most accepted theories of gravity so far.
+However, despite these goals, the difficulties on finding out a quantum theory of gravitation, has driven the community to think that GR is not the ultimate gravitational theory.
+
+At the galactic scale, on the one hand, the gravitational lensing produced by the local distribution of energy and matter, suggests the presence of an exotic form of matter, which have not been seen by the current light-based telescopes~\cite{Sofue:2000jx}.
+Such an abundance in the galaxy, is also compatible with the velocity profile of the stars at the outter regions of it and is needed to reach the current state of structures formation.
+This matter would interact mostly (if not only) through gravity~\cite{Sofue:2000jx}.
+At the cosmic scale, on the other hand, experimental data obtained from the type Ia supernovae observations, indicates that our Universe is passing through a phase of accelerated expansion~\cite{Riess:1998cb}.
+The latter experimental fact suggests the existence of an exotic form of energy in order to generate such an observed behaviour.
+
+The shortcomings of GR on describing these manifestations of new physics, have been taken as the main motivation to look for new gravitational degrees of freedom beyond GR.
+Among the families of possible extensions, higher dimensional gravity could shed some lights on the nature of these new degrees of freedom, appearing as material fields within the dimensionally reduced theory.
+For instance, as it was shown in the early works of T.
+Kaluza and O.
+Klein, the existence of one extra dimension within GR would give rise to an unified picture of gravity and electromagnetism, along with a spectrum of new heavy particles~\cite{Kaluza:1921tu,*Klein:1926tv}.
+This fact opened the posibility of a geometrical understanding of interactions, where the gauge group arise as a consequence of the topology of the spatial compact manifold in a higher dimensional spacetime, while the gauge fields arise through the dimensional reduction.
+Even though the original model have not reached a realistic phenomenology so far, the idea of higher dimensions has been taken into account by string theory~\cite{Green:1987sp,*Green:1987mn} and by ADD~\cite{ArkaniHamed:1998rs,*Antoniadis:1998ig} and RS~\cite{Randall:1999ee,*Randall:1999vf} models, as an attempt of unification and as a solution to the hierarchy problem, respectively.
+
+It is well-known that, in four dimensions, the Einstein--Hilbert action with cosmological constant is the unique theory which leads to second order field equations for the metric.
+In higher dimensions, however, this is no longer true.
+A particular combination of squared terms in the curvature can be added to the gravitational action, giving rise to second order field equations for the metric.
+The most general theory in arbitrary dimensions, which preserves the same features of the four dimensional Einstein--Hilbert action with cosmological constant, is described by the Lanczos--Lovelock action~\cite{Lanczos:1938sf,Lovelock:1971yv}.
+Such a theory, possesses no ghosts~\cite{Zumino:1985dp} and it has the same degrees of freedom as the Einstein--Hilbert Lagrangian in arbitrary dimensions~\cite{Henneaux:1990au}.
+It is endowed with the so-called Gauss--Bonnet term which, in $N$-dimensions, reads as
 \begin{equation}\label{GB}
@@
 where $\tilde{R}_{\alpha\beta\mu\nu}$ is the Riemannian curvature of a manifold with metric $g_{\mu\nu}$ and $g$ its determinant.
-$\tilde{R}_{\mu\nu}$ and $\tilde{R}$ are the Ricci tensor and Ricci scalar built with the curvature.
+$\tilde{R}_{\mu\nu}$ and $\tilde{R}$ are the Ricci tensor and Ricci scalar, respectively, built up by means of the Riemannian curvature.
 
-Written in four dimensions, Eq.~\eqref{GB} adds no dynamics to the metric, since it represents a topological invariant proportional to the Euler characteristic and therefore can always be written locally as a boundary term, while in lower dimensions vanishes identically.
-Is then natural to extend gravitational Lagrangians by dimensional continuation of Euler densities to build a general theory for a metric in arbitrary dimensions, while keeping the two important features of Einstein's General Relativity just mentioned: the absence of ghosts and second order field equations.
-These family of Lagrangians for gravity where introduced by Lovelock and bear his name.
-It can be built in arbitrary dimensions as a linear combination of dimensionally continued topological invariants.
-Equivalently, the previous Lagrangian can be written in exterior forms, that provide shorter expressions together with a natural framework for further generalizations, in $N$ dimensions reads
+In four dimensions, the Gauss--Bonnet term~\eqref{GB} adds no dynamics to the metric, since it represents a topological invariant proportional to the Euler characteristic class, which can be written locally as a boundary term.
+In dimensions higher than four, however, the Gauss--Bonnet term does contribute to the field equations, while preserving the same features of GR.
+Equivalently, the previous $N$-dimensional Lagrangian can be written in exterior forms as
 \begin{equation}\label{GB2}
@@
 This Lagrangian was also identified as the low energy correction for a spin two field in string theory~\cite{Zwiebach:1985uq}.
-Thus, the consideration of further Lovelock terms in a higher dimensional manifold can be interpreted as semiclassical corrections of a gravity action.
-The higher dimensional scenario seems to be a natural framework for unification.
-Early in the last century, it was realized by T.
-Kaluza and developed by O.
-Klein, that the existence of one extra direction in Einstein geometric theory of gravity would give rise a unified picture of gravity and electromagnetism together with a spectrum of new heavy particles~\cite{Kaluza:1921tu,*Klein:1926tv}.
-This opened the posibility of a geometrical understanding of interactions given by non-Abelian gauge groups as a concequence of the topology of a spatial compact section in a higher dimensional spacetime.
-Eventhough the original model, and further developments have not yet reached a realistic phenomenology, it is clear that this possibility of unification of internal and spacetimes symmetries, together with a material manifestation of geometric degrees of freedom provides a fruitful arena for looking for clues of unity among the interactions.
-The compactification of higher Lovelock terms has been considered \cite{MuellerHoissen:1985mm,MuellerHoissen:1989yv} and its consequences in cosmology in \cite{MuellerHoissen:1985ij,Deruelle:1986iv,Deruelle:1989fj}.
-In the Riemannian five dimensional theory, exact wormholes solutions have been found in vacuum \cite{Dotti:2006cp,Dotti:2007az}, and suported by non-exotic matter, making the sign of the Gauss-Bonnet coupling the source of exotic behaviour \cite{Mehdizadeh:2015jra}, higher order Lovelock models \cite{Mehdizadeh:2015dta}, as well as in higher order compactified Lovelock theories with torsion \cite{Canfora:2008ka} also have these solutions.
+Thus, the consideration of further Lanczos--Lovelock terms in a higher dimensional manifold can be interpreted as the semiclassical corrections to the gravity action.
+
+The compactification of higher Lanczos--Lovelock terms has been considered in~\cite{MuellerHoissen:1985mm,MuellerHoissen:1989yv} and its consequences in cosmology in Refs.~\cite{MuellerHoissen:1985ij,Deruelle:1986iv,Deruelle:1989fj}.
+In the Riemannian five-dimensional theory, exact wormholes solutions have been found in vacuum in Refs.~\cite{Dotti:2006cp,Dotti:2007az}.
+The same kind of solutions were found without exotic matter, making the sign of the Gauss-Bonnet coupling the source of such an exotic behaviour~\cite{Mehdizadeh:2015jra}.
+These latter solutions were also obtained in higher order Lanczos--Lovelock models~\cite{Mehdizadeh:2015dta}, as well as in the higher order compactified Lanczos--Lovelock theories with torsion~\cite{Canfora:2008ka}.
 Classically, wormholes must be supported by a kind of phantom energy, compatible with the cosmological hypothesis.
-We see their theoretical existence in these extended models as a good omen for the search of new degrees of freedom that could provide explanation for the exotic matter-energy abundance in the Universe.
-\\
-Torsional degrees of freedom, on the other hand, as a further feature of spacetime had a \textit{renaissance} with the works of D.
-W.
-Sciama and T.
-Kibble \cite{Kibble:1961ba} in Poincar\'e gauge theory.
-In that framework, spacetime torsion arises naturally as the field strength of a gauge connection for the Poincar\'e group (see \cite{Hehl:1976kj,Blagojevic:2002du} for a complete review).
-The pose of gravity as a gauge theory has major benefits in the problem of a unifyied picture for interactions since it allows it to interpret it as arising by means of a localization procedure, where spacetime is a Riemann-Cartan manifold.
-In this context, it has been shown that torsional functions produce an accelerating profyle compatible with the current state of the Universe \cite{Shie:2008ms}, as well as non-minimal coupling of Euler forms in Riemann--Cartan spacetimes allow torsion to behave as matter of exotic characteristics \cite{Toloza:2013wi} customary for expansion.
-It becomes essential to understand Kaluza--Klein's ideas in the presence of torsion.
-Work on the subject reports the phenomenology of extra dimensional torsion \cite{Kalinowski:1980da}, metric dependent torsional models of extra dimensions \cite{Shankar:2012vd} and its consequences in cosmology \cite{Chen:2009ep}, compatification of higher dimensional Brans--Dicke models with torsion \cite{German:1993bq}.
-Torsion free solutions where found in \cite{Aros:2007nn} for models in first order compactified gravity.\\
-Eventhough at the macroscopic scale, spacetime seems to lack torsion, there is no logical constraint to assume it non-existing, as E.
-Cartan argued, moreover since vacuum predictions of Einstein--Cartan theory hold the experimental tests of General Relativity.\\
-This work is organized as follows: in Sec.~\ref{KK} we discuss the introduction of torsional degrees of freedom compatible with the Kaluza--Klein geometric reasoning, as well as fix our notation and conventions.
-Sec.~\ref{5EGB} studies the dynamics of the general Lovelock--Cartan action in five-dimensional spacetime and its dimensional reduction.
-Then, in Sec.~\ref{cosmos} we look for cosmological solutions in the reduced four-dimensional manifold of the Friedmann--Robertson--Walker class, with torsion both in the reduced manifold and as new dynamical fields arising from the compact section taken to be $S^1$.
-Conslusions and remarks are given next in Sec.~\ref{conclusions}.
-We also found appropriate to incorporate a number of appendices to make this work a self contained article.
-%_-----------------------------------------------------------------------------------------
+Their existence in these extended models, implies the presence of degrees of freedom that could provide an explanation for the exotic matter-energy abundance in the Universe.
+
+On the other hand, the renaissance of the Riemann--Cartan geometries came through the advent of the pioneer works of Sciama~\cite{Sciama:1962} and Kibble~\cite{Kibble:1961ba} on Poincar\'e gauge theories (see also Refs.~\cite{Hehl:1976kj,Blagojevic:2002du}).
+In such a framework, the torsion appears as the field strength of translations.
+%In such a framework, the vielbein is interpreted as the gauge connection of translations and the torsion as its field strength.
+The formulation of gravity as a gauge theory has major benefits in the problem of a unified picture of interactions, since it allows to interpret it as arising through a localization procedure, where the spacetime is a Riemann--Cartan manifold.
+%Due to this, several authors have relaxed the torsion-free condition since, as E. Cartan argued, there is no a fundamental reason to assume a torsion-free condition {\it a priori}. 
+Moreover, the vacuum predictions of its simplest formulation --the Einstein--Cartan theory-- holds the experimental tests of GR.
+The cosmological consequences of non-Riemannian geometries has been widely studied in the literature (for a review, see Ref.~\cite{Puetzfeld:2004yg}).
+%For instance, the pioneer works of Kopczy\'nski showed how the introduction of a spinning dust, acting as a source of the curvature and torsion, avoids the appearance of cosmological singularities~\cite{Kopczynski:1972,Kopczynski:1973}.
+In the context of the early evolution of the Universe, the spin density can drive the inflationary phase within the Einstein--Cartan theory, along with the avoidance of singularities~\cite{Gasperini:1986mv}.
+It has been shown in Ref.~\cite{Shie:2008ms} that torsional degrees of freedom might produce the current phase of accelerated expansion of the Universe.
+Additionally, the nonminimal coupling of the Euler characteristic form with a scalar field in Riemann--Cartan spacetimes, allows the torsional degrees of freedom to behave as an exotic induced-matter content, customary for the phase of accelerated expansion~\cite{Toloza:2013wi}.
+
+In the framework of the Kaluza--Klein theories in higher dimensional Riemann--Cartan geometries, phenomenology of the extra dimensional torsion was found in Ref.~\cite{Kalinowski:1980da}, metric dependent torsional models of extra dimensions was studied in~\cite{Shankar:2012vd}, along with its consequences in cosmology~\cite{Chen:2009ep}.
+Compatification of higher dimensional Brans--Dicke models with torsion was considered in Ref.~\cite{German:1993bq}.
+Torsion-free black-hole solutions were found in~\cite{Aros:2007nn}, for first order compactified gravity.
+
+In this work, we present a new class of cosmological solutions for the Lovelock--Cartan theory~\cite{Mardones:1990qc} in five dimensions, outside the Chern--Simons point.
+The geometry is described by the five-dimensional Riemann--Cartan manifold, where the compact manifold is $S^1$ and the four dimensional manifold is described by means of the Friedmann--Robertson--Walker metric.
+Such a theory implies a nonvanishing torsion in vacuum due to the presence of the Gauss--Bonnet term, in contrast to the five-dimensional Einstein--Cartan theory.
+After the dimensional reduction is performed, the torsional degrees of freedom adds to the energy density and pressure of the four dimensional matter fields.
+In some cases, the solutions avoid the appearance of initial singularities and drive the accelerated expansion of the Universe, while the radius of the compact manifold goes to zero.
+
+This work is organized as follows: In Sec.~\ref{KK} we introduce the torsional degrees of freedom compatible with the Kaluza--Klein geometry, as well as fixing our notation and conventions.
+In Sec.~\ref{5EGB}, we study the dynamics of the general Lovelock--Cartan action in five-dimensional spacetime and its dimensional reduction.
+In Sec.~\ref{cosmos}, we look for cosmological solutions in the reduced four-dimensional manifold of the Friedmann--Robertson--Walker class with nontrivial torsion in both: the reduced manifold as well as new dynamical fields arising from the compact manifold taken to be $S^1$.
+Conclusions and remarks are given in Sec.~\ref{conclusions}.
+We also incorporate a number of appendices to make this work a self-contained article.
 
 \section{Kaluza--Klein Ansatz in Riemann--Cartan Spacetimes\label{KK}}
-Let $M_N$ be a $N$ dimensional differential manifold.
-Every quantity defined on $M_N$ will wear hats $\hat{x}$.
-Capital Greek characters (coordinate indices) and capital Latin characters (Lorentz indices) run along the $N$ dimensions, while lower case run in the $(N-1)$ dimensional reduced manifold.\\
-Let $\hat{\omega}^{AB}$ be the spin connection 1-form that describes the affine structure of the manifold and $\hat{e}^A=\hat{e}^{A}_{\ \Gamma}\,\text{d}\hat{x}^\Gamma$ be the vielbein 1-form that defines the metric structure of the same manifold through the relation $\hat{g}_{\Gamma\Delta}=\hat{e}^{A}_{\ \Gamma}\hat{e}^{B}_{\ \Delta}\hat{\eta}_{AB}$.
+
+Let $M_N$ be a $N$-dimensional differential manifold.
+Every quantity defined on $M_N$ will be denoted by hats $\hat{x}$.
+Capital Greek characters (coordinate indices) and capital Latin characters (Lorentz indices) run over the $N$ dimensions, while lower case run in the $(N-1)$-dimensional reduced manifold.
+
+Let $\hat{\omega}^{AB}$ be the spin connection 1-form describing the affine structure of the $N$-dimensional manifold and $\hat{e}^A=\hat{e}^{A}_{\ \Gamma}\,\text{d}\hat{x}^\Gamma$ be the vielbein 1-form defining the metric structure of the same manifold through the relation $\hat{g}_{\Gamma\Delta}=\hat{e}^{A}_{\ \Gamma}\hat{e}^{B}_{\ \Delta}\hat{\eta}_{AB}$.
 The vielbein is the mapping that relates coordinate indices (Greek characters) with the Lorentz indices (Latin characters).
-In our convention the flat Lorentzian metric will be $\hat{\eta}_{AB}=\,$diag$\,(-,+,...,+)$, and the $N$ dimensional Levi-Civita symbol $\hat{\epsilon}_{A_1...A_{N-1}}$ is reduced by defining
+In our convention, the flat Lorentzian metric will be $\hat{\eta}_{AB}=\,$diag$\,(-,+,...,+)$, and the $N$-dimensional Levi-Civita symbol $\hat{\epsilon}_{A_1...A_N}$ is reduced by defining
 \begin{align*}
-\epsilon_{a_1...a_{N-1}}\equiv\hat{\epsilon}_{a_1...a_{N-1}N}.
+\hat{\epsilon}_{a_1...a_{N-1}N} \equiv \epsilon_{a_1...a_{N-1}},
 \end{align*} 
-Our convention fixes $\hat{\epsilon}_{01...N}=1$.\\
-Riemannian fields (torsion free) will also be explicited by a tilde, as it was done in Eq.~\eqref{GB} and Eq.~\eqref{GB2} for the curvature.\\
+since the coordinates of $M_N$ run from $0$ to $N-1$.
+Our convention fixes $\hat{\epsilon}_{01...N}=1$.
+Hereon, Riemannian fields (torsion-free) will be explicited by a tilde, as it was done in Eqs.~\eqref{GB} and~\eqref{GB2} for the curvature.
+
 Due to the topology of the manifold, we can expand the depenence of the fields on the extra coordinate, $z$, in Fourier series as  
 \begin{align}\label{Fourier}
-\varrho(\hat{x}^\Gamma)&=\sum_n\varrho_{(n)}(x^\gamma)e^{in z}.
+\hat{\varrho}(\hat{x}^\Gamma)&=\sum_n\varrho_{(n)}(x)e^{in z},
 \end{align}
-For the rest of the paper, we will focus in the $n=0$ mode of the expasion, also refered to as the low energy sector.
+where $x$ denotes the coordinates of the $(N-1)$-manifold, collectively. Henceforth, we will focus in the $n=0$ mode of the expasion, also refered to as the low energy sector.
+
 \subsection{Metric and affine structure}
-The Kaluza-Klein (KK) ansatz for the metric lies on the premise that the compact section of $M_N$ is orthogonal to the rest of the manifold at each point.
-This leads to the conclusion that the metric has the following structure
+
+The Kaluza--Klein (KK) ansatz for the metric lies on the premise that the compact manifold of $M_N$ is orthogonal to the rest of the manifold at each point.
+This leads to the following metric structure
 \begin{equation}
@@
 \end{equation}
-It introduces a scalar field $\phi$ and a vector field $A_\mu$ as new gravitational degrees of freedom materialized in the reduced manifold.
-The vielbein that keeps this structure for the metric has the following form
+It introduces a scalar field $\phi$ and a vector field $A_\mu$ as new gravitational degrees of freedom, materialized in the reduced manifold.
+The vielbein which holds this structure for the metric has the following form
 \begin{equation}
@@
 \end{equation}
-and its inverse, $\hat{E}_A^{\ \ \Gamma}$ defined such that $\hat{E}_A^{\ \ \Gamma}\hat{e}^A_{\ \ \Delta}=\delta^\Gamma_{\Delta}$, read
+and its inverse, $\hat{E}_A^{\ \ \Gamma}$, defined such that $\hat{E}_A^{\ \ \Gamma}\hat{e}^A_{\ \ \Delta}=\delta^\Gamma_{\Delta}$, reads
 \begin{equation}
@@
 \end{equation}
-and it shows that the compact section $S^1$ has its own tangent space $T_pS^1$ at each point $p\in M_N$, and is independent to $T_pM_{N-1}$ in the sense that they do not mix.
-It seems reasonable to think that if we want two manifolds to be orthogonal to one another, they will not mix vectors after the tangent mapping.\\
-The basis of vielbeins is defined modulo Lorentz transformations. Any transformed basis 
-$
-\hat{e}^{A\prime}=\hat{\Lambda}^A_{\ B}\hat{e}^B,
-$
+and it shows that the compact manifold $S^1$ has its own tangent space, $T_pS^1$, at each point $p\in M_N$, and is independent to $T_pM_{N-1}$ in the sense that they do not mix.
+It seems reasonable to think that if we want two manifolds to be orthogonal to one another, they will not mix vectors after the tangent mapping.
+
+The vielbeins are defined modulo Lorentz transformations. Any transformed basis $\hat{e}^{\prime A}=\hat{\Lambda}^A_{\ B}\,\hat{e}^B$,
 where $\hat{\Lambda}$ is a Lorentz matrix, is as suitable as $\hat{e}^A$, and therefore share the structure in Eq.~\eqref{Dvielbein}.
-The Lorentz transformations are then reduced by this fact, imposing over the Loretnz matrix
-$\hat{\Lambda}^a_{\ N}=0.
-$
+The Lorentz transformations are then reduced by this fact, imposing the condition $\hat{\Lambda}^a_{\ N}=0$.
 
-%_-----------------------------------------------------------------------------------------
-
-A Riemannian connection compatible with a $N$ dimensional vielbein, Eq.~\eqref{Dvielbein}, is built under the premise that $\mbox{d}\hat{e}^A + \hat{\tilde{\omega}}^{AB} \wedge \hat{e}^A = 0$.
+A Riemannian connection compatible with the $N$-dimensional vielbein~\eqref{Dvielbein}, is built under the premise that $\mbox{d}\hat{e}^A + \hat{\tilde{\omega}}^{AB} \wedge \hat{e}^A = 0$.
 Thus, we find
@@
 where $\tilde{\omega}^{ab}$ is the Riemannian spin connection of the reduced manifold.
-The construction of a theory for a spacetime with spin connection \eqref{DRiemannconnection} leads to the original KK theory.
-It gives rise the usual $U(1)$-invairant field strength,
+The construction of the five-dimensional Einstein--Hilbert action by means of the spin connection~\eqref{DRiemannconnection}, leads to the original KK theory.
+After the dimensional reduction is performed, the usual $U(1)$-invariant gauge field strength,
 \begin{align*}
@@
 \end{align*} 
-non-minimally coupled to the scalar field.\\
-Nevertheless, $M_N$ is in this case, not entirely described by $\hat{\tilde{\omega}}^{AB}$, but by a more general spin connection, independent of the metric degrees of freedom (see Appendix \ref{Riemann-Cartan} for details).
-We can assume a general connection 1-form of the same type (regarding its $M_{N-1}\times S^1$ decomposition) as in eq.~\eqref{DRiemannconnection}.
+appears nonminimally coupled to the scalar field.
+
+Nevertheless, in a Riemann--Cartan geometry $M_N$ is not entirely described by $\hat{\tilde{\omega}}^{AB}$, but by a more general spin connection, independent of the metric degrees of freedom (see Appendix \ref{Riemann-Cartan} for details).
+We can assume a general connection 1-form of the same type (regarding its $M_{N-1}\times S^1$ decomposition) as in Eq.~\eqref{DRiemannconnection}.
 Thus
@@
     -\beta^b-\gamma^b\hat{e}^N&0
-  \end{pmatrix}.
+  \end{pmatrix},
 \end{equation}
-It contemplates the torsional degrees of freedom appart of the usual metric ones. 
-The decomposition \eqref{Nconnection} add new independent fields into the picture.
-Here $\alpha^{ab}$ is a $0$-form, $\beta^a$ a vector valued $1$-form and $\gamma^a$ a $0$-form that admit the Fourier expansion \eqref{Fourier} regarding their dependence on the extra dimension.
-Here, as we pointed out earlier, we will take the $0$-mode of such an expansion.
+which contemplates the torsional degrees of freedom appart from the usual metric ones.
 
-%_-----------------------------------------------------------------------------------------
+The decomposition \eqref{Nconnection} adds new metric-independent fields into the reduced theory, after the dimensional reduction is performed.
+For instance, $\alpha^{ab}$ is a rank-2 Lorentz-valued antisymmetric $0$-form, which adds a new spin-$1$ field into the matter content of the reduced theory.
+The $\beta^a$ piece of the extra dimensional spin connection, is a Lorentz-valued vector $1$-form which adds a spin-$2$, spin-$1$ and a spin-$0$ fields into the reduced theory.
+The last piece, $\gamma^a$, is a Lorentz-valued vector $0$-form, which adds a new spin-$1$ field into the picture.
+All these new fields admit the Fourier expansion in Eq.~\eqref{Fourier}, regarding their dependence on the extra dimension.
+Here, as was pointed out earlier, we will consider only the zero mode of such an expansion.
 
 \subsection{Curvature and torsion}
-Using the definition of curvature
-$\hat{R}^{AB}=\mbox{d}\hat{\omega}^{AB}+\hat{\omega}^A_{\ \ C}\wedge\hat{\omega}^{CB}$,
-we find
+
+The $N$-dimensional Lorentz curvature and torsion are given by the Cartan structure equations
+\begin{align}
+\label{curvadef}
+\hat{R}^{AB} &= \mbox{d}\hat{\omega}^{AB}+\hat{\omega}^A_{\ \ C}\wedge\hat{\omega}^{CB}, \\
+\label{tordef}
+\hat{T}^A &= \hat{\mbox{D}}\hat{e}^A=\mbox{d}\hat{e}^A+\hat{\omega}^A_{\ \ B}\wedge\hat{e}^B,
+\end{align}
+respectively. Using the definition of curvature in Eq.~\eqref{curvadef}, along with the KK ansatz for the spin connection~\eqref{Nconnection}, we find
 \begin{align}\label{R1}
@@
   \label{R2}
-  \hat{R}^{aN}&=\left(\mbox{D}\beta^a+\sqrt{\phi}\gamma^a F\right)\notag\\
-  &-\left(\alpha^a_{\ b}\beta^b-\mbox{D}\gamma^a-\frac{1}{2}\gamma^a\mbox{d}\ln\phi\right)\wedge\hat{e}^N.
+  \hat{R}^{Na}&=-\left(\mbox{D}\beta^a+\sqrt{\phi}\gamma^a F\right)\notag\\
+  &+\left(\alpha^a_{\ b}\beta^b-\mbox{D}\gamma^a-\frac{1}{2}\gamma^a\mbox{d}\ln\phi\right)\wedge\hat{e}^N.
 \end{align}
-where we have used the fact that
-$
-  \mbox{d}\hat{e}^N=\frac{1}{2}\mbox{d}\ln\phi\wedge\hat{e}^N+\sqrt{\phi}F.
-$
-Similarly from the definition of torsion
-$\hat{T}^A = \hat{\mbox{D}}\hat{e}^A=\mbox{d}\hat{e}^A+\hat{\omega}^A_{\ \ B}\wedge\hat{e}^B$
-we find its distinctive parts to be
+where we have used the fact that $\mbox{d}\hat{e}^N=\frac{1}{2}\mbox{d}\ln\phi\wedge\hat{e}^N+\sqrt{\phi}F$.
+Similarly from the definition of torsion in Eq.~\eqref{tordef}, we find its distinctive parts to be
 \begin{align}\label{T1}
@@
 
-Considering the Bianchi identities of the KK structure described above, we find relevant information about the new fields. Taking the exterior covariant derivative of the field strengths
-$\hat{R}$ and $\hat{T}$, the Bianchi identities are
+Considering the Bianchi identities for the KK structure described above, we find relevant information about the new fields.
+Taking the exterior covariant derivative over the $N$-dimensional curvature and torsion, the Bianchi identities are
 \begin{align*}
-  \hat{\text{D}}\hat{R}^{AB} &= 0, & \hat{\text{D}}\hat{T}^A &= \hat{R}^A_{\ \ B}\wedge\hat{e}^B.
+  \hat{\text{D}}\hat{R}^{AB} &= 0 &\mbox{and}& &\hat{\text{D}}\hat{T}^A &= \hat{R}^A_{\ \ B}\wedge\hat{e}^B.
 \end{align*}
-A careful decomposition of the first one in its distinctive parts, gives the Bianchi identity for the curvature of the reduced spacetime together with the second derivative rules
+A careful decomposition of the first one into its distinctive parts, gives the Bianchi identity for the curvature of the reduced spacetime, together with the second derivative rules
 \begin{align*}
@@
 \end{align*}
-This is taken as proof of the tensorial condition of these new fields. While consideration of the second one gives nothing but its equivalent for the manifold $M_{N-1}$, this is
+Considering the second Bianchi identity, gives nothing but its equivalent for the manifold $M_{N-1}$, this is
 \begin{equation*}
@@
 
-%_-----------------------------------------------------------------------------------------
+This is taken as a proof of the tensorial nature of these new fields under the $4$-dimensional Lorentz transformations, which have been already discussed in Sec.~\ref{KK}.
 
-\section{Five Dimensional Einstein--Gauss--Bonnet Reduction\label{5EGB}}
-We will work in this section, the dimensional reduction of a gravitational Lagrangian that includes quadratic terms in the curvature, namely the Gauss--Bonnet term. The action is given by
+\section{Five-Dimensional Lovelock--Cartan Reduction\label{5EGB}}
+
+The most general theory requiring the Lagrangian to be (i)~an invariant $N$-form under local Lorentz transformations; (ii)~a local polynomial of the vielbein, the Lorentz connection and their exterior derivatives; (iii)~constructed without the Hodge dual, is the so-called Lovelock--Cartan theory of gravity~\cite{Mardones:1990qc}.
+This theory is the natural generalization of the Lanczos--Lovelock action, when torsional degrees of freedom are present.
+
+In five dimensions, the action principle of such a theory is given by  
 \begin{widetext}
@@
   \end{equation}
-where $\alpha_0$, $\alpha_1$ and $\alpha_2$ are coupling constants. Variations with respect to the vielbein and spin connection give the equations
+where $\alpha_0$, $\alpha_1$ and $\alpha_2$ are dimensionful coupling constants. Variation with respect to the $5$-dimensional vielbein and spin connection give the equations
   \begin{align}\label{delta e}
@@
 \end{widetext}
-In this work, we will focus on the general case when
+
+In general, the Lovelock--Cartan theory in five dimensions allows the torsional term~\cite{Mardones:1990qc},
+\begin{align}
+\label{boundary}
+\mathcal{L} \propto \hat{T}_A\wedge \hat{R}^A_{\ B}\wedge\hat{e}^B,
+\end{align}
+which can be written as a boundary term proportional to the Nieh-Yan invariant, adding no dynamics to the field equations.
+The action principle~\eqref{action5EGB}, along with the torsional term~\eqref{boundary}, describes the general Lovelock--Cartan theory in five dimensions.
+
+In this work, we will focus on the region in the parameter space where
 \begin{equation}\label{delta}
@@
 \end{equation}
-Otherwise, the Lagrangian becomes a gauge invariant Lagrangian for a larger group, namely the $AdS_5$ group, and is recognized as its Chern--Simons form \cite{Zanelli:2005sa}, \cite{Troncoso:1999pk}.
-In general, Lovelock--Cartan (LC) theory allows the torsional term
-\begin{align*}
-\mathcal{L}_{T}\sim\hat{R}_{AB}\wedge\hat{T}^A\wedge\hat{e}^B,
-\end{align*}
-that in five dimensions is a boundary term proportional to the Nieh-Yan invariant, thus it does not participate in the dynamics.
-Action \eqref{action5EGB} is the general LC theory in five dimensions.\\
-In terms of the KK ansatz presented in previous sections, the equations of motion can be decomposed in to its distinctive parts. Eq.~\eqref{delta e} leads to the following set
+This condition place us outside the Chern--Simons point since, otherwise, the field equations~\eqref{delta e} and~\eqref{delta w} remain invariant under $AdS_5$ gauge transformations, while the action principle~\eqref{action5EGB} can be recognized as its Chern--Simons form~\cite{Zanelli:2005sa,Troncoso:1999pk}.
+
+In terms of the KK ansatz presented in previous sections, the field equations can be decomposed into its distinctive parts.
+The equation of motion for the five-dimensional vielbein~\eqref{delta e}, leads to the following set of reduced field equations
 \begin{align}\label{equation}
-\epsilon_{abcd}\Big[\alpha_0 e^a\wedge e^b\wedge e^c+\frac{1}{2}\alpha_1\left(M^{ab}-L^a\wedge e^b\right)\wedge e^c&\notag\\
--\alpha_2\left(L^a\wedge M^{bc}+K^a\wedge N^{bc}\right)\Big]&=0,\\
-\epsilon_{abcd}K^a\wedge\left(\alpha_1e^b\wedge e^c+2\alpha_2M^{bc}\right)&=0,\\
+\epsilon_{abcd}\Big[\alpha_0 e^b\wedge e^c\wedge e^d + \frac{1}{2}\alpha_1\left(M^{bc}-L^b\wedge e^c\right)\wedge e^d&\notag\\
+-\alpha_2\left(L^b\wedge M^{cd} + K^b\wedge N^{cd}\right)\Big]&=0,\\
+\epsilon_{abcd}K^b\wedge\left(\alpha_1e^c\wedge e^d + 2\alpha_2M^{cd}\right)&=0,\\
 \epsilon_{abcd}\big(\alpha_0 e^a\wedge e^b\wedge e^c\wedge e^d&\notag\\
-+\alpha_1M^{ab}\wedge\wedge e^c\wedge e^d+\alpha_2M^{ab}\wedge M^{cd}\big)&=0,\\
++\alpha_1M^{ab}\wedge e^c\wedge e^d+\alpha_2M^{ab}\wedge M^{cd}\big)&=0,\\
 \epsilon_{abcd}\left(\alpha_1N^{ab}\wedge e^c\wedge e^d+2\alpha_2N^{ab}M^{cd}\right)&=0,
 \end{align}
-while Eq.~\eqref{delta w} gives
+while the equation of motion for the spin connection~\eqref{delta w} gives
 \begin{align}
@@
 \epsilon_{abcd}\left(\alpha_1e^b\wedge e^c+2\alpha_2M^{bc}\right)\wedge T^d&=0,\\
-\epsilon_{abcd}\left[\alpha_1e^b\wedge e^c V^d+2\alpha_2\Big(M^{bc}\wedge V^d+N^{bc}\wedge T^d\Big)\right]&=0.
+\epsilon_{abcd}\left[\alpha_1e^b\wedge e^c\wedge V^d+2\alpha_2\Big(M^{bc}\wedge V^d+N^{bc}\wedge T^d\Big)\right]&=0.
 \end{align}
-The fields $M^{ab}$, $N^{ab}$, $L^a$, $K^a$, $W$, $V^a$ and $Z$ are defined through \eqref{R1}, \eqref{R2}, \eqref{T1} and \eqref{T2} such that
+The fields $M^{ab}$, $N^{ab}$, $L^a$, $K^a$, $W$, $V^a$ and $Z$ are defined from the Eq.~\eqref{R1} to~\eqref{T2}, such that
 \begin{align*}
@@
 
-%_-----------------------------------------------------------------------------------------
-\section{Dimensionally Reduced EGB Cosmology}\label{cosmos}
+\section{Dimensionally Reduced Lovelock--Cartan Cosmology\label{cosmos}}
+
 \subsection{Cosmological ansatz}
-We will look for cosmological solutions of the equations of motion.
-For that, we demand the symmetries assumed by the cosmological principle, namely, isotropy and homogeneity of the fields involve.
-This is achieved by making the Lie derivative in the direction of the Killing vectors of the symmetries to vanish for each field.
-Appendix \ref{homotropic} is devoted to the details of how to reach the general ansatz that we present next.
-The vielbein compatible with the Friedman-Robertson-Walker metric is determined by one function called scale factor $a(t)$,
+
+In order to look for cosmological solutions, we demand the symmetries assumed by the cosmological principle, i.e.
+isotropy and homogeneity of the involved fields.
+This can be achieved by considering the Lie derivative of each field, along the Killing vectors which generates the symmetries, to vanish.
+The Appendix~\ref{homotropic} is devoted to the details of how to reach the general ansatz we are presenting next.
+The $4$-dimensional vielbein compatible with the Friedman--Robertson--Walker metric is determined only by one function, the so-called scale factor $a(t)$, and their components read
 \begin{align}
@@
 \end{align}
-where $k=+1,0,-1$ determines if the spatial section is closed, flat or an open $3$-space respectivelly. The scalar field is a time-dependent function
+where $k=+1,0,-1$ determines the spatial section of the four-dimensional manifold to be closed, flat or open, respectively.
+
+The scalar field is determined by one time-dependent function, denoted by
 \begin{equation}
@@
 \end{equation}
-On the other hand, the spin connection adds two functions we called $\omega(t)$ and $f(t)$,
+
+On the other hand, the spin connection adds two functions that we called $\omega(t)$ and $f(t)$,
 \begin{align}
@@
 \end{align}
-The non-vanishing components of the spin connection induced by the compact section of the manifold are
+
+The nonvanishing components of the spin connection induced by the compact manifold are
 \begin{align}
@@
 \end{align}
-For this ansatz, the $1$-form $A$ has the form $A_t(t)\text{d}t$, which can be shown to have vanishing field strength $F$.
-Thus without loss of generality, it can be set to zero by means of a $U(1)$ gauge transformation, or equivalently, a diffeomorphism transformation along the $z$-direction.\\
-The equations of motion for the cosmological ansatz gives a system of differential equations for the time dependent functions defined above.
-Curvature and torsion for this ansatz can be seen in Appendix \ref{homotropic}.
-We found that the only non-Riemannian branch demands  $\beta^a=0$.
+
+For this ansatz, however, the $1$-form $A$ is $A_t(t)\text{d}t$, which can be shown to have vanishing field strength $F$.
+Thus, without loss of generality, it can be fixed to zero by means of an $U(1)$ gauge transformation or, equivalently, a diffeomorphism transformation along the $z$-direction.
+
+The equations of motion for the cosmological ansatz give a system of differential equations for the time-dependent functions defined above.
+Curvature and torsion for this ansatz can be seen in Appendix~\ref{homotropic}.
+We find that the only non-Riemannian branch demands  $\beta^a=0$.
 Thus $b(t)=\beta(t)=0$.
-Otherwise, the system degenerates at the CS point.
+Otherwise, the system degenerates at the Chern--Simons point.
+
 The final set of equations is the following:
@@
   +2\alpha_2\left(\dot{\omega}+\omega^2\right)\left(\omega^2+\frac{k}{a^2}-f^2\right)&=0,\\
-  2\alpha_2\left(\frac{1}{2}\dot{\Phi}+\gamma\right)f\left(\dot{f}+Hf\right)
+  2\alpha_2\left(\frac{\dot{\phi}}{2\phi}+\gamma\right)f\left(\dot{f}+Hf\right)
   +h^2\left(\alpha_1-2\alpha_2\omega\gamma\right)&=0,\\
-  \omega\left(\frac{1}{2}\dot{\Phi}+\gamma\right)+\dot{\gamma}+\frac{1}{2}\gamma\dot{\Phi}-\frac{\alpha_1}{2\alpha_2}&=0,\\
+  \omega\left(\frac{\dot{\phi}}{2\phi}+\gamma\right)+\dot{\gamma}+\frac{\dot{\phi}}{2\phi}\gamma-\frac{\alpha_1}{2\alpha_2}&=0,\\
   \left(\omega^2+\frac{k}{a^2}-f^2\right)\left(\alpha_1-2\alpha_2\omega\gamma\right)-\alpha_1\omega\gamma+2\alpha_0&=0,\\
@@
 \end{align}
-where dotted functions stands for its time-derivatives. We have defined for simplicity the functions $h(t)=\omega(t)-H(t)$, where $H=\dot{a}/a$ is the Hubble function, and $\Phi=\ln\phi$.
+where dotted functions stand for its time-derivatives. For the sake of simplicity, we have defined $h(t)=\omega(t)-H(t)$, where $H=\dot{a}/a$ is the Hubble function.
 
@@
 
-The system is integrated developing two non-Riemannian branches, one for each value of the parameter $u_\pm$ defenied to be
+The system is integrated out developing two non-Riemannian branches, one for each value of the parameter $u_\pm$ defined to be
 \begin{equation}\label{u}
@@
 where $\Delta$ was given in Eq.~\eqref{delta}.
-Eqs.~\eqref{eqn1} to~\eqref{eqn2} reduced to the Riemannian system when \mbox{$f=h=0$} and \mbox{$\gamma=-\dot{\Phi}/2$}.
+The Eqs.~\eqref{eqn1} to~\eqref{eqn2} are reduced to the Riemannian system when \mbox{$f=h=0$} and \mbox{$\gamma=-\dot{\phi}/2\phi$}.
 The details of such a model can be seen in Refs.~\cite{Deruelle:1986iv,Deruelle:2003ck,Henriques:1986jw,Ishihara:1986if,Kleidis:1997mu}.
 
-Due to Eq.~\eqref{u}, the solutions will be valid in the region of parameter space where $\Delta>0$.
-The function $\omega(t)$ satisfies the equation $\dot{\omega}+\omega^2+u_{\pm}=0$, that for the three significantly different values of $u_\pm$, has the following solutions
+Due to Eq.~\eqref{u}, the solutions will be valid in the region of parameter space where $\Delta>0$. The function $\omega(t)$ satisfies the equation
+\begin{align}
+\dot{\omega}+\omega^2+u_{\pm} &= 0,
+\end{align}
+which, for the three significantly different values of $u_\pm$, has the following solutions
 \begin{align}
@@
 where $t_0$ is an integration constant to be fixed.
-We relegate the time dependence of the remaining fields to these expressions for $\omega$ and list their explicit time dependence in Appendix~\ref{solutions t}.
+We relegate the time-dependence of the remaining fields to these expressions for $\omega$ and list their explicit time dependence in Appendix~\ref{solutions t}.
 The solutions read
@@
 \begin{equation*}
-  Y_{\pm}=\mp2\frac{\alpha_1 u_\pm-2\alpha_0}{\sqrt{6\Delta}}=-\frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2},
+  Y_{\pm} = -\frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2},
 \end{equation*}
-and where $a_0$ and $\phi_0$ are integration constants.
+and $a_0$ and $\phi_0$ are integration constants.
 
-\subsubsection{Bouncing universes}
+\subsubsection{Bouncing solutions}
+
 For $u_\pm>0$ the solutions are periodic.
-In that case, the scale factor starts from a singular point and reach a future one after a time $t_{\text{crunch}}=\pi/\sqrt{u_\pm}$.
-Depending on the particular sector of parameter space, the solution undergoes an expanding and contracting age that allow it to reach a minimum value without collapsing, to expand again before a big crunch.
-Otherwise, it reaches a maximum value before collapse.
+In that case, the scale factor starts from a singular point and reaches a future one after a time $t_{\rm bounce}=\pi/\sqrt{u_\pm}$.
+Depending on the particular region on the parameter space, the scale factor undergoes an expanding and contracting age, allowing to reach a minimum value without collapsing, to expand again before a big crunch.
+Otherwise, it reaches a maximum value before its collapse.
 It also shows an inflationary phase at early times when the scale factor grows exponentially.
 
-\subsubsection{Expanding and contracting universes}
-For $u_\pm\leq 0$, depending on the sign of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will infinitely expand while the scalar field reach an asymtotic vanishing value for late cosmic times, after a period of expansion.
-This is interpreted as the late cosmic time behaviour of the radius of the compact spacial section $S^1$, that for this familly of solutions provides a natural explanation for its scale.
+\subsubsection{Expanding and contracting solutions}
+
+For $u_\pm\leq 0$, depending on the sign of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will infinitely expand while the scalar field reaches an asymtotic vanishing value for late cosmic times, after a period of expansion.
+This is interpreted as the late time behaviour of the radius of the compact manifold $S^1$, which for this family of solutions provides a natural explanation for its scale.
 Otherwise, in contrast to the previous behaviour, the scale factor contracts to an asymtotic vanishing value while the scalar field has no upper bound.
-In the case of strictly negative $u_\pm$, for either contracting or expanding scale factor, its initial value remains finite, thus the solutions do not present an original singularity.
-It has been reported that the GB term in a gravitational action can prevent the universe to expand from an initial singularity \cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if}, which is also the case here.
+In the case of strictly negative $u_\pm$, for either contracting or expanding scale factors, their initial values remain finite and the solutions do not present an initial singularity.
+It has been reported that the GB term in a gravitational action can prevent the universe to expand from an initial singularity \cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if}, which is also the present case.
 
 \subsubsection{Effective energy density and pressure}
-Eq.~\eqref{equation} take the familiar form
+
+The Eq.~\eqref{equation} can be written, by means of the identity~\eqref{curvature decomp}, in the familiar form
 \begin{align}\label{einstein equation}
-\frac{1}{2}\epsilon_{abcd}\left(\alpha_0e^a\wedge e^b\wedge e^c+\frac{1}{2}\alpha_1\tilde{R}^{ab}\wedge e^c\right)&=\tau^{eff}_d,
+-\frac{1}{2}\epsilon_{abcd}\left(\tilde{R}^{bc} - \frac{\Lambda}{3}\,e^b\wedge e^c\right)e^d &= \kappa_{G}\,\tau^{eff}_a,
 \end{align}
-when we use \eqref{curvature decomp} and define an effective energy-momentum form as
+where $\Lambda \equiv -6\alpha_0/\alpha_1$, $\kappa_{G} \equiv 2/\alpha_1$ and we have defined the effective energy-momentum $3$-form as
 \begin{align}
-\tau_a^{eff}&=\frac{1}{2}\epsilon_{abcd}\bigg[\frac{1}{2}\alpha_1\left(\kappa^b_{\ l}\wedge\kappa^{lc}+\tilde{\text{D}}\kappa^{bc}\right)\wedge e^d\notag\\
+\tau_a^{eff}&= \frac{1}{2}\epsilon_{abcd}\bigg[\frac{1}{2}\alpha_1\left(\kappa^b_{\ l}\wedge\kappa^{lc}+\tilde{\text{D}}\kappa^{bc}\right)\wedge e^d\notag\\
 &+\left(\text{D}\gamma^b+\frac{1}{2}\gamma^b\text{d}\ln\phi\right)\wedge\left(\frac{1}{2}\alpha_1e^c\wedge e^d+\alpha_2R^{cd}\right)\bigg].
@@
 The contributions of $F$, $\alpha^{ab}$ and $\beta^a$ have not been taken into account since for all practical purposes of this article, they vanish.
-The left hand side of \eqref{einstein equation} is Eintein's $3$-form, that in components gives Einstein's tensor.
-We identify the right hand side as an effective energy-momentum form that contemplates the torsional functions and the higher dimensional metric degrees of freedom as part of the gravitating matter of a four dimensional Universe in the form of a perfect fluid.
-For this energy-momentum form, we identify the energy density $\rho$ and pressure $p$ by
-\begin{align*}
-\tau_0^{eff}&=-\frac{1}{3!}\rho\epsilon_{0ijk}e^i\wedge e^j\wedge e^k,\\
-\tau_i^{eff}&=-\frac{1}{2}p\epsilon_{0ijk}e^0\wedge e^j\wedge e^k.
-\end{align*} 
-Hence
+The left hand side of the Eq.~\eqref{einstein equation} is the Eintein $3$-form, which is dual to the Einstein tensor with cosmological constant in four dimensions.
+We identify its right hand side as an effective energy-momentum $3$-form, which behaves as a perfect fluid.
+Such an energy-momentum contains the torsional and the higher dimensional degrees of freedom as part of the gravitating fields in four dimensions.
+From this, we identify the energy density $\rho$ and the pressure $p$ through
+\begin{align}
+\tau_0^{eff}&=-\frac{1}{3!}\,\rho\,\epsilon_{0ijk}e^i\wedge e^j\wedge e^k,\\
+\tau_i^{eff}&=-\frac{1}{2}\,p\,\epsilon_{0ijk}e^0\wedge e^j\wedge e^k,
+\end{align} 
+where
 \begin{align}\label{rho}
@@
 \end{align}
-The identification satisfy $\dot{\rho}+3H\left(\rho+p\right)=0$. Using equations of motion, one can show that
-\begin{align*}
-\rho\sim H^2+\frac{k}{a^2}+\text{const.}
-\end{align*}
-For $u_\pm<0$, this function remains finite at the begining.
+Using the solutions obtained in the previous section , it can be shown that
+\begin{align}
+\rho &= \frac{3}{2}\alpha_1\left(H^2 + \frac{k}{a^2}\right) + 3\alpha_0, \\
+p &= -\alpha_1\left(\dot{H} + \frac{3}{2}\,H^2 + \frac{k}{2a^2}\right) - 3\alpha_0,
+\end{align}
+satisfying the continuity equation $\dot{\rho}+3H\left(\rho+p\right)=0$.
+
+For $u_\pm<0$, the energy density remains finite at the begining.
 From expressions \eqref{rho} and \eqref{presion} one can see that the induced energy density and pressure are not positive definite quantities.
-In fact, due to the presence of torsion and extra dimensional fields, the universe solutions undergoes through accelerating expansive age with no need of extra fluids of exotic class.
+In fact, the Universe undergoes through an accelerating expansion phase due to the presence of torsion and the extra dimensional fields.
 
@@
 
-\section{Discussion and Conclusions}\label{conclusions}
-It is known that KK theory, in its simplest version, give rise matter fields a manifestation of extra dimensional metric degrees of freedom.
-The kind of matter that arises in the reduced spacetime is subordinated to the topology taken for the extra compact manifold.
+\section{Discussion and Conclusions\label{conclusions}}
+
+It is known that KK theory, in its simplest version, gives rise to matter fields as a manifestation of the extra dimensional metric degrees of freedom.
+The kind of matter arising in the reduced spacetime is subordinated to the topology taken for the extra compact manifold.
 In general, those fields will gravitate and modulate the behaviour of spacetime in its time evolution.
-We have constructed the simplest model in five dimensional spacetime that allow the presence of torsion in the KK decomposition.
-In the cosmological model, the induced matter present due to an extra compact section taken to be $S^1$ are the usual scalar field --a metric degree of freedom-- and a vector field that comes from the dimensional reduction of a general spin connection.
-Torsion exists both in the reduced manifold and in the higher dimensional and indeed, it participate in the effective energy-momentum form that gravitates as a perfect fluid.
-Since the sign of the energy density nor of the presures in the fluid are in no way guaranteed, it is interpreted as a class of matter of exotic characteristics.
-Exact solution for a five dimensional vacuum are studied.
-Accelerating, and bouncing and inflationary universes are possible by the equations of motion.\\
+
+We find cosmological solutions of the Friedmann--Robertson--Walker class for the dimensionally reduced Lovelock--Cartan theory in five dimensions~\cite{Mardones:1990qc}.
+This theory allows torsion to propagate in vacuum, in contrast to the Einstein--Cartan theory.
+The induced matter fields in the reduced cosmological scenario are the usual scalar field --a metric degree of freedom-- and a vector field, coming from the dimensional reduction of a general spin connection.
+Those fields participate in the effective energy-momentum form.
+Neither the energy density nor the presure are strictly positive definite, therefore, can be interpreted as a class of matter of exotic features.
+Five-dimensional vacuum solutions are studied, where accelerating, bouncing and inflationary solutions were found.
+
 There are periods of time where $f^2(t)$ becomes negative.
-As we can see from Eq.~\eqref{homotorsion}, $f$ corresponds to the completely antisymmetric part of torsion and is unseen by classical particles following geodesics along spacetime.
-It does not couple to spin-0 or spin-1 bosons, but it does appear as an effective interaction term in the Dirac Lagrangian writen in a curved spacetime \cite{Carroll:1994dq}
+As we can see from Eq.~\eqref{homotorsion}, $f$ corresponds to the completely antisymmetric part of torsion and is unseen by classical particles following geodesics.
+It does not couple minimally to spin-0 or spin-1 bosons, but it does appear as an effective interaction term in the Dirac Lagrangian written in a curved spacetime~\cite{Hehl:1976kj}
 \begin{align*}
-\mathcal{L}_{\text{Interaction}}&=-\frac{i}{8}T_{\alpha\mu\nu}\bar{\Psi}\Gamma^{\alpha\mu\nu}\Psi\\
-&=\frac{3}{2}f\bar{\Psi}\Gamma^0\Gamma^5\Psi.
+\mathcal{L}_{\text{Interaction}}&=-\frac{i}{8}T_{\alpha\mu\nu}\bar{\Psi}\Gamma^{\alpha\mu\nu}\Psi =\frac{3}{2}f\bar{\Psi}\Gamma^0\Gamma^5\Psi
 \end{align*}
@@
 For imaginary values of $f(t)$, the corresponding violation of current conservation can be interpreted as particle creation.
-Similar results where found in \cite{Toloza:2013wi} for another cosmological model with torsion that share some of the couplings of a reduced action.
+Similar results where found in~\cite{Toloza:2013wi} for another cosmological model with torsion which shares some of the couplings of the reduced action.
 \begin{acknowledgments}
@@
 \appendix
-\section{Riemann-Cartan Geometry}\label{Riemann-Cartan}
+
+\section{Riemann--Cartan Geometry\label{Riemann-Cartan}}
+
 In Riemann--Cartan geometry, both the vielbein $e^A$ and the spin connection $\omega^{AB}$ are independent fields.
-We will drop the hats since we will not deal with the KK decomposition.
-Stationary variations of a gravity actions must be taken with respect to both fields, allowing in general, the presence of torsion.
-The spin connection, however, can be decomposed in a Riemannian part $\tilde{\omega}^{AB}$ that satisfies $\tilde{\mbox{D}}e^A=0$, and a contorsion $\kappa^{AB}=-\kappa^{BA}$, such that
+We will drop the hats since we are not dealing with the KK decomposition.
+Stationary variations of a gravitational action principle must be taken with respect to both fields, allowing in general the presence of torsion.
+The spin connection, however, can be decomposed in a Riemannian part $\tilde{\omega}^{AB}$ satisfying $\tilde{\mbox{D}}e^A=0$, and a contorsion piece $\kappa^{AB}=-\kappa^{BA}$, such that
 \begin{equation}\label{spin separation}
@@
 \end{equation}
-On the other hand, the curvature 2-form also suffers corrections with respect to the Riemannian one due to the presence of torsional degrees of freedom.
-This can be seen explictly by taking the definition of curvature $R^{AB}=\mbox{d}\omega^{AB}+\omega^a_{\ C}\wedge\omega^{CB}$ and \eqref{spin separation} to find
+On the other hand, the curvature $2$-form also suffers corrections with respect to the Riemannian one due to the presence of torsional degrees of freedom.
+This can be seen explictly by taking the definition of curvature in Eq.~\eqref{curvadef} and~\eqref{spin separation} to find
 \begin{equation}\label{curvature decomp}
-  R^{AB}=\tilde{R}^{AB}+\kappa^A_{\ C}\wedge\kappa^{CB}+\tilde{\mbox{D}}\kappa^{AB},
+  R^{AB} = \tilde{R}^{AB} + \tilde{\mbox{D}}\kappa^{AB} + \kappa^A_{\ C}\wedge\kappa^{CB},
 \end{equation}
-where $\tilde{R}^{AB}$ is the Riemannian curvature built with $\tilde{\omega}^{AB}$.
+where $\tilde{R}^{AB}$ is the Riemannian curvature built up from $\tilde{\omega}^{AB}$.
 
-%_-----------------------------------------------------------------------------------------
+\section{Isotropic--Homogeneus Ansatz\label{homotropic}}
 
-\section{Isotropic--Homogeneus Ansatz}\label{homotropic}
 The cosmological principle demands the spacial section of spacetime to be isotropic and homogeneus.
-This means that the fields involve in the model must be compatible with this assumption.
-A spacetime is an isotropic spacetime with respect to certain point $P$ if after a rotation with respect to an axis that pass through $P$, all the geometrical properties remain invariant, thus the spacetime looks the same in all directions.
-By homogeneity we understand that spacetime looks the same from every point $P$.
-These two assumptions are translated in the Killing equations for the fields, which are the vanishing of the Lie derivatives of the fields along the direction of the vector generators of symmetries $\{\zeta^\lambda_{i}\}$.
+This means that the fields involved in the model must be compatible with this assumption.
+A spacetime is an isotropic spacetime with respect to certain point $P$ if after a rotation with respect to an axis passing through $P$, all the geometrical properties remain invariant, thus the spacetime looks the same in all directions.
+Homogeneity is understood such that spacetime looks the same from every point $P$.
+These two assumptions are translated in the Killing equations, which are the vanishing of the Lie derivatives of the fields along the vectors which generate the symmetries $\{\zeta^\lambda_{i}\}$.
 In particular the Killing equations for the metric tensor and the torsion tensor are
@@
 \end{align}
-and also for the components (tensors) $\alpha_{\mu\nu}=-\alpha_{\nu\mu}$, $\beta_{\mu\nu}$ and $\gamma_\mu$, defined such that
+and also for the components of $\alpha_{\mu\nu}=-\alpha_{\nu\mu}$, $\beta_{\mu\nu}$ and $\gamma_\mu$, defined such that
 \begin{align*}
@@
 \end{align*} 
-and whose Killing equations are analogous to the tensorial one, \eqref{killing metric} and vectorial one,\eqref{killing vector}.\\
-The set of Killing vectors $\{\zeta^\lambda_i\}$ are, on the one hand, the generators of spatial rotations in three dimensions $SO(3)$, $\xi_i=\epsilon_{ijk}x_j\partial_k$, and the Killing vectors assosiated with spacial translations $\mathcal{P}_i=\sqrt{1-kr^2}\partial_i$, that satisfy the algebra
+and whose Killing equations are analogous to the tensorial one,~\eqref{killing metric} and vectorial one,~\eqref{killing vector}.
+
+The set of Killing vectors $\{\zeta^\lambda_i\}$ are the generators of $SO(3)$, which generate the spatial rotations in three dimensions, $\xi_i=\epsilon_{ijk}x_j\partial_k$, and the Killing vectors assosiated with spacial translations $\mathcal{P}_i=\sqrt{1-kr^2}\partial_i$, satisfy the algebra
 \begin{align}
@@
 \end{align}
-These requirements on the fields are translated to a set of first order differential equations whose most general solutions determines our ansatz structure \eqref{vielbein cosmo} to \eqref{gamma cosmo}.
+These requirements on the fields are translated to a set of first order differential equations whose most general solutions determines our ansatz structure from Eq.~\eqref{vielbein cosmo} to \eqref{gamma cosmo}.
 The fields strengths for isotropic--homogeneus vielbein and spin connection are written next.
@@
 
-%_-----------------------------------------------------------------------------------------
-\section{Time Dependence of the Solutions}\label{solutions t}
-We supply here the time dependent expressions of the cosmological solutions for the significantly differents values of $u_\pm$ in terms of the dimensionless parameter $\tau=\sqrt{|u_\pm|}(t-t_0)$.\\
-\newline
+\section{Time dependence of the solutions\label{solutions t}}
+
+We supply here the time-dependent expressions for the cosmological solutions, given the differents values of $u_\pm$, in terms of the dimensionless parameter~$\tau=\sqrt{|u_\pm|}(t-t_0)$.
+\medskip
+
 $\bullet\ u_\pm>0$:
@@
   \gamma(t)&=-\sqrt{u_\pm}\cot\tau,\\
-  \phi(t)&=\phi_0\sin^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2}\cos^2\tau\right],\\
-  a(t)&=\frac{a_0}{\sqrt{u_\pm}}|\cos\tau|\exp\left[\frac{\pm\sqrt{6\Delta}}{24\alpha_2}\cos^2\tau\right],\\
+  \phi(t)&=\phi_0\sin^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2 u_\pm}\cos^2\tau\right],\\
+  a(t)&=\frac{a_0}{\sqrt{u_\pm}}|\cos\tau|\exp\left[\frac{\pm\sqrt{6\Delta}}{24\alpha_2 u_\pm}\cos^2\tau\right],\\
   f^2(t)&=u_\pm\frac{k}{a_0^2}\sec^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cos^2\tau\right]\notag\\
@@
 \end{align}
-\\
-\medskip
+
 $\bullet\ u_\pm=0$:
@@
 \end{align}
-\\
-\medskip
+
 $\bullet\ u_\pm<0$:

### 2016-09-12 KKosmology.tex
New abstract by Oscar

@@
 \hypersetup{%
-  pdftitle={Dimensional reduction on $S^1$ of Lanczos--Lovelock and Lovelock--Cartan theories of gravity},
+  pdftitle={Cosmology from Kaluza--Klein reduction of five-dimensional Lovelock--Cartan theory},
   pdfauthor={Oscar Castillo-Felisola,}{Cristobal Corral,}{Simon del Pino,}{Francisca Ramirez.},
-  pdfkeywords={Torsion,} {Generalised Gravity.},
+  pdfkeywords={Torsion,} {Cosmology,}{Kaluza--Klein reduction,}{Generalised Gravity.},
   pdflang={English}
@@
 
-\title{Five-Dimensional Kaluza--Klein Cosmology in Lovelock--Cartan Theory}
+\title{Cosmology from  Kaluza--Klein reduction of five-dimensional Lovelock--Cartan Theory}
 
@@
 \begin{abstract}
- We study the Kaluza--Klein dimensional reduction of the Lovelock--Cartan theory on a $N$-dimensional Riemann--Cartan manifold, where the compact manifold has $S^1$ topology.
-First, we focus on the dimensional reduction in presence of torsion, to proceed with the Lovelock--Cartan theory in five dimensions.
-We find cosmological solutions of the Friedman--Robertson--Walker class for the dimensionally reduced theory.
-In the five-dimensional vacuum, this theory allows a nontrivial torsion in contrast to the Einstein--Cartan theory.
-After the dimensional reduction is performed, the torsional degrees of freedom induce a nonvanishing energy-momentum tensor in four dimensions, which behaves as a perfect fluid.
-Cosmological solutions describing accelerated expansion phases as well as contracting solutions are found.
+  We study the Kaluza--Klein dimensional reduction of the Einstein--Gauss--Bonnet theory on a $N$-dimensional Riemann--Cartan manifold, where the compact dimension has $S^1$ topology.
+We proceed with a detailed reduction of the Lovelock--Cartan theory in five dimensions.
+We find cosmological solutions of the Friedman--Robertson--Walker class for the effective theory.
+  %%In the five-dimensional vacuum, this theory allows a nontrivial torsion in contrast to the Einstein--Cartan theory.
+  The torsional degrees of freedom induce a nonvanishing energy-momentum tensor in four dimensions, which behaves as a perfect fluid.
+We find cosmological solutions describing accelerated expansion phases as well as contracting universes.
 The model shows a dynamical compactification of the extra dimension in some regions of the parameter space.

### 2016-09-12 KKosmology.tex
Abstract and introduction, discussed!

@@
 \hypersetup{%
-  pdftitle={Cosmology from Kaluza--Klein reduction of five-dimensional Lovelock--Cartan theory},
+  pdftitle={Cosmology from Kaluza--Klein reduction on five-dimensional Lovelock--Cartan theory},
   pdfauthor={Oscar Castillo-Felisola,}{Cristobal Corral,}{Simon del Pino,}{Francisca Ramirez.},
@@
 
-\title{Cosmology from  Kaluza--Klein reduction of five-dimensional Lovelock--Cartan Theory}
+\title{Cosmology from  Kaluza--Klein reduction on five-dimensional Lovelock--Cartan Theory}
 
@@
 \begin{abstract}
-  We study the Kaluza--Klein dimensional reduction of the Einstein--Gauss--Bonnet theory on a $N$-dimensional Riemann--Cartan manifold, where the compact dimension has $S^1$ topology.
-We proceed with a detailed reduction of the Lovelock--Cartan theory in five dimensions.
+  We study the Kaluza--Klein dimensional reduction of the Lovelock--Cartan theory on a five-dimensional manifold, where the compact dimension has $S^1$ topology.
 We find cosmological solutions of the Friedman--Robertson--Walker class for the effective theory.
   %%In the five-dimensional vacuum, this theory allows a nontrivial torsion in contrast to the Einstein--Cartan theory.
-  The torsional degrees of freedom induce a nonvanishing energy-momentum tensor in four dimensions, which behaves as a perfect fluid.
-We find cosmological solutions describing accelerated expansion phases as well as contracting universes.
+  The torsional degrees of freedom induce a nonvanishing energy-momentum tensor in four dimensions.
+We find solutions describing accelerated expansion phases as well as contracting universes.
 The model shows a dynamical compactification of the extra dimension in some regions of the parameter space.
@@
 
-%% \pacs{02.40.Ma,04.50.Kd,04.90.+e}
-%% \keywords{Affine Gravity, Torsion, Generalised Gravity.}
+\pacs{02.40.Ma,04.50.Kd,04.90.+e}
+\keywords{Affine Gravity, Torsion, Generalised Gravity.}
 
@@
 In this framework, the gravitational interaction is understood as the geometrical manifestation of the spacetime.
-The experimental status of General Relativity (GR), regarding the solar system tests~\cite{Will:2014kxa} and the detection of gravitational waves by the LIGO collaboration~\cite{Abbott:2016blz}, have settled GR as one of the most accepted theories of gravity so far.
-However, despite these goals, the difficulties on finding out a quantum theory of gravitation, has driven the community to think that GR is not the ultimate gravitational theory.
+The experimental status of General Relativity (GR), regarding the solar system tests~\cite{Will:2014kxa} and the detection of signals consistent with the merge of two black holes by the LIGO collaboration~\cite{Abbott:2016blz}, have settled it as the most successful theory of gravity so far.
+However, the difficulties on finding out explanations for the so called \emph{dark sector} of the Universe,
+% and a \emph{consistent} quantum theory of gravitation,
+have driven the community to think that GR is not the ultimate gravitational theory. 
+The \emph{dark sector} of the Universe is composed by two different forms of ``matter'', whose effects are detected at different scales.
+On the one hand, at galactic scales, the gravitational lensing produced by the local distribution of energy, suggests the presence of an exotic form of matter, unseen by the current light-based telescopes~\cite{Sofue:2000jx}, for this it is named dark matter.
+Such an abundance in the galaxy is also compatible with the velocity profile of stars at its outter regions and it is needed to achieve the current structures formation.
+This matter would interact mostly (if not only) through gravity.
+On the other hand, at cosmic scales, the experimental data obtained from Type Ia supernovae observations, indicates that our Universe is passing through a phase of accelerated expansion~\cite{Riess:1998cb}.
+This introduces the dark energy hypothesis to generate such a behavior.
 
-At the galactic scale, on the one hand, the gravitational lensing produced by the local distribution of energy and matter, suggests the presence of an exotic form of matter, which have not been seen by the current light-based telescopes~\cite{Sofue:2000jx}.
-Such an abundance in the galaxy, is also compatible with the velocity profile of the stars at the outter regions of it and is needed to reach the current state of structures formation.
-This matter would interact mostly (if not only) through gravity~\cite{Sofue:2000jx}.
-At the cosmic scale, on the other hand, experimental data obtained from the type Ia supernovae observations, indicates that our Universe is passing through a phase of accelerated expansion~\cite{Riess:1998cb}.
-The latter experimental fact suggests the existence of an exotic form of energy in order to generate such an observed behaviour.
+The shortcomings of GR on describing these phenomena, are the main motivation to look for new gravitational degrees of freedom.
+Among the possible extensions, higher dimensional models could shed some light on the nature of these new degrees of freedom.
+For instance, as it was shown in the early works of T.~Kaluza and O.~Klein, the existence of an extra dimension within the GR framework would give rise to a unified picture of gravity and electromagnetism, along with a spectrum of new heavy particles~\cite{Kaluza:1921tu,*Klein:1926tv}.
+This idea opened the possibility of a novel geometrical understanding of interactions, where the gauge group arise as a consequence of the topology of the spatial compact manifold in a higher dimensional spacetime.
+The idea of higher dimensions comes naturally in diverse physical models, as for example: supersymmetry and supergravity, string theory~\cite{Green:1987sp,*Green:1987mn}, and novel proposals by Arkani-Hammed, Dimopoulos and Dvali~\cite{ArkaniHamed:1998rs,*Antoniadis:1998ig}, and Randall and Sundrum~\cite{Randall:1999ee,*Randall:1999vf} models, as attempts of solving the hierarchy problem.
 
-The shortcomings of GR on describing these manifestations of new physics, have been taken as the main motivation to look for new gravitational degrees of freedom beyond GR.
-Among the families of possible extensions, higher dimensional gravity could shed some lights on the nature of these new degrees of freedom, appearing as material fields within the dimensionally reduced theory.
-For instance, as it was shown in the early works of T.
-Kaluza and O.
-Klein, the existence of one extra dimension within GR would give rise to an unified picture of gravity and electromagnetism, along with a spectrum of new heavy particles~\cite{Kaluza:1921tu,*Klein:1926tv}.
-This fact opened the posibility of a geometrical understanding of interactions, where the gauge group arise as a consequence of the topology of the spatial compact manifold in a higher dimensional spacetime, while the gauge fields arise through the dimensional reduction.
-Even though the original model have not reached a realistic phenomenology so far, the idea of higher dimensions has been taken into account by string theory~\cite{Green:1987sp,*Green:1987mn} and by ADD~\cite{ArkaniHamed:1998rs,*Antoniadis:1998ig} and RS~\cite{Randall:1999ee,*Randall:1999vf} models, as an attempt of unification and as a solution to the hierarchy problem, respectively.
+In four dimensions, the Einstein--Hilbert action with cosmological constant is the most general theory which leads to second order field equations for the metric.
+In higher dimensions, particular combinations of higher order terms in the curvature can be added to the gravitational action, giving rise to second order field equations for the metric.
+The most general theory in arbitrary dimensions, which preserves the features of the four-dimensional Einstein--Hilbert, is described by the Lanczos--Lovelock action~\cite{Lanczos:1938sf,Lovelock:1971yv}.
+Such a theory, has no ghosts~\cite{Zumino:1985dp} and it has the same degrees of freedom as the Einstein--Hilbert Lagrangian in arbitrary dimensions~\cite{Henneaux:1990au}.
 
-It is well-known that, in four dimensions, the Einstein--Hilbert action with cosmological constant is the unique theory which leads to second order field equations for the metric.
-In higher dimensions, however, this is no longer true.
-A particular combination of squared terms in the curvature can be added to the gravitational action, giving rise to second order field equations for the metric.
-The most general theory in arbitrary dimensions, which preserves the same features of the four dimensional Einstein--Hilbert action with cosmological constant, is described by the Lanczos--Lovelock action~\cite{Lanczos:1938sf,Lovelock:1971yv}.
-Such a theory, possesses no ghosts~\cite{Zumino:1985dp} and it has the same degrees of freedom as the Einstein--Hilbert Lagrangian in arbitrary dimensions~\cite{Henneaux:1990au}.
-It is endowed with the so-called Gauss--Bonnet term which, in $N$-dimensions, reads as
+The simplest possible extra term in the Lanczos--Lovelock action  is a quadratic construction of curvatures, called Gauss--Bonnet term, which reads
 \begin{equation}\label{GB}
-  \mathcal{L}_{GB}=\text{d}^N \!x \,\sqrt{-g}\left(\tilde{R}^2-4\tilde{R}_{\mu\nu}\tilde{R}^{\mu\nu}+\tilde{R}_{\alpha\beta\mu\nu}
+  \mathcal{L}_{\mathrm{GB}} = \dn{N}{x}\sqrt{-g}\left(\tilde{R}^2-4\tilde{R}_{\mu\nu}\tilde{R}^{\mu\nu}+\tilde{R}_{\alpha\beta\mu\nu}
   \tilde{R}^{\alpha\beta\mu\nu}\right),
@@
 where $\tilde{R}_{\alpha\beta\mu\nu}$ is the Riemannian curvature of a manifold with metric $g_{\mu\nu}$ and $g$ its determinant.
-$\tilde{R}_{\mu\nu}$ and $\tilde{R}$ are the Ricci tensor and Ricci scalar, respectively, built up by means of the Riemannian curvature.
+$\tilde{R}_{\mu\nu}$ and $\tilde{R}$ are the Ricci tensor and Ricci scalar, respectively.
 
-In four dimensions, the Gauss--Bonnet term~\eqref{GB} adds no dynamics to the metric, since it represents a topological invariant proportional to the Euler characteristic class, which can be written locally as a boundary term.
-In dimensions higher than four, however, the Gauss--Bonnet term does contribute to the field equations, while preserving the same features of GR.
-Equivalently, the previous $N$-dimensional Lagrangian can be written in exterior forms as
-\begin{equation}\label{GB2}
-\mathcal{L}_{GB} = \epsilon_{a_1...a_N} \tilde{R}^{a_1a_2}\wedge\tilde{R}^{a_3a_4}\wedge e^{a_5}\wedge...\wedge e^{a_N}.
-\end{equation}
-This Lagrangian was also identified as the low energy correction for a spin two field in string theory~\cite{Zwiebach:1985uq}.
-Thus, the consideration of further Lanczos--Lovelock terms in a higher dimensional manifold can be interpreted as the semiclassical corrections to the gravity action.
+The Gauss--Bonnet term, in four dimensions, adds no dynamics to the metric, since it represents a topological invariant proportional to the Euler characteristic class, which can be written locally as a boundary term.
+In dimensions higher than four it contributes to the field equations.
+%% Equivalently, the previous $N$-dimensional Lagrangian can be written in exterior forms as
+%% \begin{equation}\label{GB2}
+%% \mathcal{L}_{GB} = \epsilon_{a_1...a_N} \tilde{R}^{a_1a_2}\wedge\tilde{R}^{a_3a_4}\wedge e^{a_5}\wedge...\wedge e^{a_N}.
+%% \end{equation}
+This Lagrangian was also identified as the low-energy correction for a spin two field in string theory~\cite{Zwiebach:1985uq}.
+%%Thus, the consideration of further Lanczos--Lovelock terms in a higher dimensional manifold can be interpreted as the semiclassical corrections to the gravity action.
 
-The compactification of higher Lanczos--Lovelock terms has been considered in~\cite{MuellerHoissen:1985mm,MuellerHoissen:1989yv} and its consequences in cosmology in Refs.~\cite{MuellerHoissen:1985ij,Deruelle:1986iv,Deruelle:1989fj}.
-In the Riemannian five-dimensional theory, exact wormholes solutions have been found in vacuum in Refs.~\cite{Dotti:2006cp,Dotti:2007az}.
-The same kind of solutions were found without exotic matter, making the sign of the Gauss-Bonnet coupling the source of such an exotic behaviour~\cite{Mehdizadeh:2015jra}.
-These latter solutions were also obtained in higher order Lanczos--Lovelock models~\cite{Mehdizadeh:2015dta}, as well as in the higher order compactified Lanczos--Lovelock theories with torsion~\cite{Canfora:2008ka}.
-Classically, wormholes must be supported by a kind of phantom energy, compatible with the cosmological hypothesis.
-Their existence in these extended models, implies the presence of degrees of freedom that could provide an explanation for the exotic matter-energy abundance in the Universe.
+On the one hand, classically, wormholes must be supported by a kind energy compatible with the cosmological hypothesis.
+Their existence in these extended models, implies the presence of degrees of freedom that could provide an explanation for the exotic matter/energy abundance in the Universe.
+In the Riemannian five-dimensional Lanczos--Lovelock theory, exact wormholes solutions have been found in vacuum~\cite{Dotti:2006cp,Dotti:2007az}.
+The same kind of solutions were found with matter~\cite{Mehdizadeh:2015jra}, in higher order Lanczos--Lovelock models~\cite{Mehdizadeh:2015dta}, and in higher order compactified Lanczos--Lovelock theories with torsion~\cite{Canfora:2008ka}.
+Additionally, the compactification of higher Lanczos--Lovelock terms has been considered in Refs.~\cite{MuellerHoissen:1985mm,MuellerHoissen:1989yv} and their cosmology in Refs.~\cite{MuellerHoissen:1985ij,Deruelle:1986iv,Deruelle:1989fj}.
 
-On the other hand, the renaissance of the Riemann--Cartan geometries came through the advent of the pioneer works of Sciama~\cite{Sciama:1962} and Kibble~\cite{Kibble:1961ba} on Poincar\'e gauge theories (see also Refs.~\cite{Hehl:1976kj,Blagojevic:2002du}).
-In such a framework, the torsion appears as the field strength of translations.
+On the other hand, the renaissance of the Riemann--Cartan geometries came through the pioneer works of Sciama~\cite{Sciama:1962} and Kibble~\cite{Kibble:1961ba} on Poincar\'e gauge theories (see also Refs.~\cite{Hehl:1976kj,Blagojevic:2002du}).
+In such a framework, the torsion appears as the field strength of translations, sourced by the spin-current.
 %In such a framework, the vielbein is interpreted as the gauge connection of translations and the torsion as its field strength.
@@
 %Due to this, several authors have relaxed the torsion-free condition since, as E. Cartan argued, there is no a fundamental reason to assume a torsion-free condition {\it a priori}. 
-Moreover, the vacuum predictions of its simplest formulation --the Einstein--Cartan theory-- holds the experimental tests of GR.
+Moreover, the vacuum predictions of its simplest formulation---the Einstein--Cartan theory---holds the experimental tests of GR.
 The cosmological consequences of non-Riemannian geometries has been widely studied in the literature (for a review, see Ref.~\cite{Puetzfeld:2004yg}).
@@
 
-In this work, we present a new class of cosmological solutions for the Lovelock--Cartan theory~\cite{Mardones:1990qc} in five dimensions, outside the Chern--Simons point.
-The geometry is described by the five-dimensional Riemann--Cartan manifold, where the compact manifold is $S^1$ and the four dimensional manifold is described by means of the Friedmann--Robertson--Walker metric.
-Such a theory implies a nonvanishing torsion in vacuum due to the presence of the Gauss--Bonnet term, in contrast to the five-dimensional Einstein--Cartan theory.
-After the dimensional reduction is performed, the torsional degrees of freedom adds to the energy density and pressure of the four dimensional matter fields.
+The extension of the Lanczos--Lovelock theories with nonvanishing torsion is known as Lovelock--Cartan~\cite{Mardones:1990qc}.
+In that framework, we present a new class of cosmological solutions in five dimensions, where the compact dimension is $S^1$.
+The theory admits a nonvanishing torsion in vacuum due to the presence of the Gauss--Bonnet term, in contrast to the five-dimensional Einstein--Cartan theory.
+After the dimensional reduction, the extra degrees of freedom generate an energy density and pressure as four-dimensional matter fields.
 In some cases, the solutions avoid the appearance of initial singularities and drive the accelerated expansion of the Universe, while the radius of the compact manifold goes to zero.
 
-This work is organized as follows: In Sec.~\ref{KK} we introduce the torsional degrees of freedom compatible with the Kaluza--Klein geometry, as well as fixing our notation and conventions.
+This work is organized as follows: In Sec.~\ref{KK} we introduce the torsional degrees of freedom compatible with the Kaluza--Klein geometry, and fix our notation and conventions.
 In Sec.~\ref{5EGB}, we study the dynamics of the general Lovelock--Cartan action in five-dimensional spacetime and its dimensional reduction.
-In Sec.~\ref{cosmos}, we look for cosmological solutions in the reduced four-dimensional manifold of the Friedmann--Robertson--Walker class with nontrivial torsion in both: the reduced manifold as well as new dynamical fields arising from the compact manifold taken to be $S^1$.
+In Sec.~\ref{cosmos}, we look for cosmological solutions of the Friedmann--Robertson--Walker class with nontrivial torsion.
 Conclusions and remarks are given in Sec.~\ref{conclusions}.

### 2016-09-13 KKosmology.tex
Section of Kaluza--Klein. Discussed!

@@
 In this framework, the gravitational interaction is understood as the geometrical manifestation of the spacetime.
-The experimental status of General Relativity (GR), regarding the solar system tests~\cite{Will:2014kxa} and the detection of signals consistent with the merge of two black holes by the LIGO collaboration~\cite{Abbott:2016blz}, have settled it as the most successful theory of gravity so far.
+The experimental status of General Relativity (GR), regarding the solar system tests~\cite{Will:2014kxa} and the detection of signals consistent with the merge of two black holes by the LIGO collaboration~\cite{Abbott:2016blz,Abbott:2016nmj}, have settled it as the most successful theory of gravity so far.
 However, the difficulties on finding out explanations for the so called \emph{dark sector} of the Universe,
@@
 
-Let $M_N$ be a $N$-dimensional differential manifold.
+In the following we will consider $M_N$ to be a $N$-dimensional differential manifold.
 Every quantity defined on $M_N$ will be denoted by hats $\hat{x}$.
@@
 
-Let $\hat{\omega}^{AB}$ be the spin connection 1-form describing the affine structure of the $N$-dimensional manifold and $\hat{e}^A=\hat{e}^{A}_{\ \Gamma}\,\text{d}\hat{x}^\Gamma$ be the vielbein 1-form defining the metric structure of the same manifold through the relation $\hat{g}_{\Gamma\Delta}=\hat{e}^{A}_{\ \Gamma}\hat{e}^{B}_{\ \Delta}\hat{\eta}_{AB}$.
-The vielbein is the mapping that relates coordinate indices (Greek characters) with the Lorentz indices (Latin characters).
-In our convention, the flat Lorentzian metric will be $\hat{\eta}_{AB}=\,$diag$\,(-,+,...,+)$, and the $N$-dimensional Levi-Civita symbol $\hat{\epsilon}_{A_1...A_N}$ is reduced by defining
-\begin{align*}
-\hat{\epsilon}_{a_1...a_{N-1}N} \equiv \epsilon_{a_1...a_{N-1}},
-\end{align*} 
-since the coordinates of $M_N$ run from $0$ to $N-1$.
+The differential structure on $M_N$ is determined by two fields.
+The spin connection 1-form, $\hat{\omega}^{AB} = \hat{\omega}^{AB}{}_{\Gamma}\,\text{d}\hat{x}^\Gamma$, describes the affine structure of the $N$-dimensional manifold, and the vielbein 1-form, $\hat{e}^A=\hat{e}^{A}{}_{\Gamma}\,\text{d}\hat{x}^\Gamma$, defines the metric structure of the same manifold through the relation $\hat{g}_{\Gamma\Delta} = \hat{e}^{A}{}_{\Gamma}\hat{e}^{B}{}_{\Delta}\hat{\eta}_{AB}$.
+The vielbein is the mapping that relates coordinate indices with the Lorentz indices.
+In our convention, both Lorentzian metrics \(\hat{\eta}\) and \(\eta\) have positive signature, and the Levi-Civita symbols are such that
+\[\hat{\epsilon}_{a_1 \cdots a_{N-1} N} \equiv \epsilon_{a_1 \cdots a_{N-1}},\]
+%% \[\hat{\epsilon}_{0 1 \cdots N-1} = {\epsilon}_{0 1 \cdots N-1} = 1,\]
+which allows us to define the reduced symbol by fixing the last index.
+Additionally, all Riemannian fields (torsion-free) will be explicited by a tilde, as it was done in Eq.~\eqref{GB} for the curvature.
+%% In our convention, the flat Lorentzian metric will be $\hat{\eta}_{AB}=\diag(-,+,\cdots,+)$, and the $N$-dimensional Levi-Civita symbol $\hat{\epsilon}_{A_1 \cdots A_N}$ is reduced by defining
+%% \begin{align*}
+%% \hat{\epsilon}_{a_1 \cdots a_{N-1} N} \equiv \epsilon_{a_1 \cdots a_{N-1}},
+%% \end{align*} 
+%% since the coordinates of $M_N$ run from $0$ to $N-1$.
 Our convention fixes $\hat{\epsilon}_{01...N}=1$.
-Hereon, Riemannian fields (torsion-free) will be explicited by a tilde, as it was done in Eqs.~\eqref{GB} and~\eqref{GB2} for the curvature.
+Hereon, Riemannian fields (torsion-free) will be explicited by a tilde, as it was done in Eqs.~\eqref{GB} for the curvature.
 
@@
 \begin{align}\label{Fourier}
-\hat{\varrho}(\hat{x}^\Gamma)&=\sum_n\varrho_{(n)}(x)e^{in z},
+  \hat{\varrho}(\hat{x}^\Gamma)&=\sum_n\varrho_{(n)}(x)e^{i n z},
 \end{align}
-where $x$ denotes the coordinates of the $(N-1)$-manifold, collectively. Henceforth, we will focus in the $n=0$ mode of the expasion, also refered to as the low energy sector.
+where $x$ denotes the coordinates of the $(N-1)$-manifold, collectively. Henceforth, we will focus in the $n=0$ mode of the expasion, also refered to as the low-energy sector.
 
@@
 
-The Kaluza--Klein (KK) ansatz for the metric lies on the premise that the compact manifold of $M_N$ is orthogonal to the rest of the manifold at each point.
+The Kaluza--Klein (KK) ansatz for the metric relies on the premise that the compact dimension of $M_N$ is orthogonal to the rest of the manifold at each point (see for example Ref.~\cite{Blagojevic:2002du}).
 This leads to the following metric structure
@@
 \end{equation}
-It introduces a scalar field $\phi$ and a vector field $A_\mu$ as new gravitational degrees of freedom, materialized in the reduced manifold.
-The vielbein which holds this structure for the metric has the following form
+It introduces a scalar field $\phi$ and a vector field $A_\mu$ as new gravitational degrees of freedom. The vielbein which holds this structure for the metric has the following form
 \begin{equation}
   \label{Dvielbein}
-  \hat{e}^A_{\ \ \Gamma} =
+  \hat{e}^A{}_{\Gamma} =
   \begin{pmatrix}
-    \hat{e}^a_{\ \gamma}& 0\\
-    \hat{e}^N_{\ \gamma} & \hat{e}^N_{\ \ z}
+    \hat{e}^a{}_{\gamma}& 0\\
+    \hat{e}^N{}_{\gamma} & \hat{e}^N{}_{z}
   \end{pmatrix}
@@
   \begin{pmatrix}
-    e^a_{\ \gamma}& 0\\
+    e^a{}_{\gamma}& 0\\
     \sqrt{\phi}A_\gamma & \sqrt{\phi}
@@
 \end{equation}
-and its inverse, $\hat{E}_A^{\ \ \Gamma}$, defined such that $\hat{E}_A^{\ \ \Gamma}\hat{e}^A_{\ \ \Delta}=\delta^\Gamma_{\Delta}$, reads
+and its inverse, $\hat{E}_A{}^{\Gamma}$, defined such that $\hat{E}_A{}^{\Gamma}\hat{e}^A{}_{\Delta}=\delta^\Gamma_{\Delta}$, reads
 \begin{equation}
   \label{Dinversevielbein}
-  \hat{E}_A^{\ \ \Gamma} =
+  \hat{E}_A{}^{\Gamma} =
   \begin{pmatrix}
-    \hat{E}_a^{\ \gamma}& 0\\
-    \hat{E}_a^{\ z} & \hat{E}_N^{\ \ z}
+    \hat{E}_a{}^{\gamma}& 0\\
+    \hat{E}_a{}^{z} & \hat{E}_N{}^{z}
   \end{pmatrix}
@@
   \begin{pmatrix}
-    E_a^{\ \gamma}& 0\\
+    E_a{}^{\gamma}& 0\\
     -A_a & \sqrt{\phi}^{-1}
@@
 \end{equation}
-and it shows that the compact manifold $S^1$ has its own tangent space, $T_pS^1$, at each point $p\in M_N$, and is independent to $T_pM_{N-1}$ in the sense that they do not mix.
-It seems reasonable to think that if we want two manifolds to be orthogonal to one another, they will not mix vectors after the tangent mapping.
+and it shows that the compact manifold $S^1$ has its own tangent space, $T_pS^1$, at each point $p\in M_N$, and is independent of $T_pM_{N-1}$ in the sense that they do not mix.
+%It seems reasonable to think that if we want two manifolds to be orthogonal to one another, they will not mix vectors after the tangent mapping.
 
-The vielbeins are defined modulo Lorentz transformations. Any transformed basis $\hat{e}^{\prime A}=\hat{\Lambda}^A_{\ B}\,\hat{e}^B$,
-where $\hat{\Lambda}$ is a Lorentz matrix, is as suitable as $\hat{e}^A$, and therefore share the structure in Eq.~\eqref{Dvielbein}.
-The Lorentz transformations are then reduced by this fact, imposing the condition $\hat{\Lambda}^a_{\ N}=0$.
+The vielbeins are defined modulo Lorentz transformations. Any transformed basis $\hat{e}^{\prime A} = \hat{\Lambda}^A{}_{B}\,\hat{e}^B$,
+where $\hat{\Lambda}$ is a Lorentz matrix, is as suitable as $\hat{e}^A$, and therefore share the structure of Eq.~\eqref{Dvielbein}.
+The Lorentz transformations are then constrained by $\hat{\Lambda}^a{}_{N}=0$.
 
-A Riemannian connection compatible with the $N$-dimensional vielbein~\eqref{Dvielbein}, is built under the premise that $\mbox{d}\hat{e}^A + \hat{\tilde{\omega}}^{AB} \wedge \hat{e}^A = 0$.
+A Riemannian connection compatible with the \mbox{$N$-dimensional} vielbein~\eqref{Dvielbein}, is built under the premise that $\mbox{d}\hat{e}^A + \hat{\tilde{\omega}}^{A}{}_B \wedge \hat{e}^B = 0$.
 Thus, we find
-\begin{align}
+\begin{equation}
   \label{DRiemannconnection}
+  \begin{split}
     \hat{\tilde{\omega}}^{ab}&=\tilde{\omega}^{ab}-\frac{1}{2}\sqrt{\phi}F^{ab}\hat{e}^N,\\
     \hat{\tilde{\omega}}^{Na}&=\frac{1}{2}\sqrt{\phi}F^a_{\ \ l}e^l+\frac{1}{2}\partial^a\ln\phi\hat{e}^N,
-\end{align}
-where $\tilde{\omega}^{ab}$ is the Riemannian spin connection of the reduced manifold.
+  \end{split}
+\end{equation}
+where $\tilde{\omega}^{ab}$ is the Riemannian spin connection of the reduced manifold and $F_{ab}$ are the components of the field strength of $A_b$ defined as
+\begin{equation}
+  F=\text{d}A=\frac{1}{2}F_{ab}\, e^a\wedge e^b.
+\end{equation}
 The construction of the five-dimensional Einstein--Hilbert action by means of the spin connection~\eqref{DRiemannconnection}, leads to the original KK theory.
-After the dimensional reduction is performed, the usual $U(1)$-invariant gauge field strength,
-\begin{align*}
-F=\text{d}A=\frac{1}{2}F_{ab}\, e^a\wedge e^b,
-\end{align*} 
-appears nonminimally coupled to the scalar field.
 
-Nevertheless, in a Riemann--Cartan geometry $M_N$ is not entirely described by $\hat{\tilde{\omega}}^{AB}$, but by a more general spin connection, independent of the metric degrees of freedom (see Appendix \ref{Riemann-Cartan} for details).
+Nevertheless, in a Riemann--Cartan geometry $M_N$ is not entirely described by $\hat{\tilde{\omega}}^{AB}$, but by a more general spin connection, independent of the metric degrees of freedom (see Appendix~\ref{Riemann-Cartan} for details).
 We can assume a general connection 1-form of the same type (regarding its $M_{N-1}\times S^1$ decomposition) as in Eq.~\eqref{DRiemannconnection}.
-Thus
-the most general spin connection on $M_N$ compatible with the KK decomposition is given by 
+Thus, the most general spin connection on $M_N$ compatible with the KK decomposition is given by
 \begin{equation}\label{Nconnection}
@@
   \begin{pmatrix}
-    \omega^{ab}+\alpha^{ab}\hat{e}^N&\beta^a+\gamma^a\hat{e}^N\\
-    -\beta^b-\gamma^b\hat{e}^N&0
-  \end{pmatrix},
+    \omega^{ab}+\alpha^{ab}\hat{e}^N & \beta^a+\gamma^a\hat{e}^N\\
+    -\beta^b-\gamma^b\hat{e}^N & 0
+  \end{pmatrix}.
 \end{equation}
-which contemplates the torsional degrees of freedom appart from the usual metric ones.
 
-The decomposition \eqref{Nconnection} adds new metric-independent fields into the reduced theory, after the dimensional reduction is performed.
-For instance, $\alpha^{ab}$ is a rank-2 Lorentz-valued antisymmetric $0$-form, which adds a new spin-$1$ field into the matter content of the reduced theory.
-The $\beta^a$ piece of the extra dimensional spin connection, is a Lorentz-valued vector $1$-form which adds a spin-$2$, spin-$1$ and a spin-$0$ fields into the reduced theory.
-The last piece, $\gamma^a$, is a Lorentz-valued vector $0$-form, which adds a new spin-$1$ field into the picture.
-All these new fields admit the Fourier expansion in Eq.~\eqref{Fourier}, regarding their dependence on the extra dimension.
-Here, as was pointed out earlier, we will consider only the zero mode of such an expansion.
+The decomposition in Eq.~\eqref{Nconnection} adds new metric-independent fields.
+The 0-form $\alpha^{ab}$ is an antisymmetric tensor of spin-1.
+The 1-form $\beta^a = \beta^a{}_\mu \de{x}^\mu$ generically adds a spin-$2$, a spin-$1$ and a spin-$0$ fields.
+The last piece, $\gamma^a$, is a vector $0$-form of spin-1.
 
@@
   \label{curvadef}
-\hat{R}^{AB} &= \mbox{d}\hat{\omega}^{AB}+\hat{\omega}^A_{\ \ C}\wedge\hat{\omega}^{CB}, \\
+  \hat{R}^{AB} &= \mbox{d}\hat{\omega}^{AB}+\hat{\omega}^A_{\ \ C}\wedge\hat{\omega}^{CB} \\
+  &= \frac{1}{2} \hat{R}^{AB}{}_{\mu\nu} \, \de{x}^\mu \we \de{x}^\nu, \notag\\
   \label{tordef}
-\hat{T}^A &= \hat{\mbox{D}}\hat{e}^A=\mbox{d}\hat{e}^A+\hat{\omega}^A_{\ \ B}\wedge\hat{e}^B,
+  \hat{T}^A &= \mbox{d}\hat{e}^A+\hat{\omega}^A_{\ \ B}\wedge\hat{e}^B \\
+  &= \frac{1}{2} \hat{T}^{A}{}_{\mu\nu} \, \de{x}^\mu \we \de{x}^\nu. \notag
 \end{align}
-respectively. Using the definition of curvature in Eq.~\eqref{curvadef}, along with the KK ansatz for the spin connection~\eqref{Nconnection}, we find
-\begin{align}\label{R1}
+Using the definition of curvature in Eq.~\eqref{curvadef},  with the KK ansatz for the spin connection~\eqref{Nconnection}, we find
+\begin{align}
+  \label{R1}
   \hat{R}^{ab}&=R^{ab}+\sqrt{\phi}\alpha^{ab}F-\beta^a\wedge\beta^b\notag\\
@@
 \end{align}
-where we have used the fact that $\mbox{d}\hat{e}^N=\frac{1}{2}\mbox{d}\ln\phi\wedge\hat{e}^N+\sqrt{\phi}F$.
+
 Similarly from the definition of torsion in Eq.~\eqref{tordef}, we find its distinctive parts to be
@@
 
-\subsubsection{Bianchi identities}
+\subsection{Bianchi identities}
 
@@
 \end{align*}
-Considering the second Bianchi identity, gives nothing but its equivalent for the manifold $M_{N-1}$, this is
+The second Bianchi identity, gives nothing but its equivalent for the manifold $M_{N-1}$, this is
 \begin{equation*}
@@
 
-This is taken as a proof of the tensorial nature of these new fields under the $4$-dimensional Lorentz transformations, which have been already discussed in Sec.~\ref{KK}.
+This is taken as a proof of the tensorial nature of these new fields under the four-dimensional Lorentz transformations.
 

### 2016-09-13 KKosmology.tex
Modification of bibliography

@@
   \label{curvadef}
-  \hat{R}^{AB} &= \mbox{d}\hat{\omega}^{AB}+\hat{\omega}^A_{\ \ C}\wedge\hat{\omega}^{CB} \\
-  &= \frac{1}{2} \hat{R}^{AB}{}_{\mu\nu} \, \de{x}^\mu \we \de{x}^\nu, \notag\\
+  \hat{R}^{AB} &= \mbox{d}\hat{\omega}^{AB}+\hat{\omega}^A_{\ \ C}\wedge\hat{\omega}^{CB} = \frac{1}{2} \hat{R}^{AB}{}_{CD} \, \hat{e}^C \wedge \hat{e}^D,\\
   \label{tordef}
-  \hat{T}^A &= \mbox{d}\hat{e}^A+\hat{\omega}^A_{\ \ B}\wedge\hat{e}^B \\
-  &= \frac{1}{2} \hat{T}^{A}{}_{\mu\nu} \, \de{x}^\mu \we \de{x}^\nu. \notag
+  \hat{T}^A &= \mbox{d}\hat{e}^A+\hat{\omega}^A_{\ \ B}\wedge\hat{e}^B = \frac{1}{2} \hat{T}^{A}{}_{BC} \, \hat{e}^B \wedge \hat{e}^C. 
 \end{align}
@@
 \begin{widetext}
-  \begin{equation}\label{action5EGB}
-    I=\int\limits_{M_5}\hat{\epsilon}_{ABCDE}\Big(\frac{\alpha_0}{5}\hat{e}^A\wedge\hat{e}^B\wedge\hat{e}^C\wedge
+  \begin{equation}
+    \label{action5EGB}
+    I = \int_{M_5} \hat{\epsilon}_{ABCDE} \Big(\frac{\alpha_0}{5}\hat{e}^A\wedge\hat{e}^B\wedge\hat{e}^C\wedge
     \hat{e}^D\wedge\hat{e}^E
@@
   \end{equation}
-  where $\alpha_0$, $\alpha_1$ and $\alpha_2$ are dimensionful coupling constants. Variation with respect to the $5$-dimensional vielbein and spin connection give the equations
-  \begin{align}\label{delta e}
+  where $\alpha_0$, $\alpha_1$ and $\alpha_2$ are dimensionful coupling constants. Its variation with respect to the five-dimensional vielbein and spin connection give the equations
+  \begin{align}
+    \label{delta e}
     \hat{\epsilon}_{ABCDE}\Big(\alpha_0\hat{e}^A\wedge\hat{e}^B\wedge\hat{e}^C\wedge\hat{e}^D
-    +\alpha_1\hat{R}^{AB}\wedge\hat{e}^C\wedge\hat{e}^D
-    +\alpha_2\hat{R}^{AB}\wedge\hat{R}^{CD}\Big)&=0,
-    \\\label{delta w}
+    + \alpha_1\hat{R}^{AB}\wedge\hat{e}^C\wedge\hat{e}^D
+    + \alpha_2\hat{R}^{AB}\wedge\hat{R}^{CD}\Big)&=0,
+    \\
+    \label{delta w}
     \hat{\epsilon}_{ABCDE}\Big(\alpha_1\hat{e}^C\wedge\hat{e}^D+
@@
 
-In general, the Lovelock--Cartan theory in five dimensions allows the torsional term~\cite{Mardones:1990qc},
+The Lovelock--Cartan theory in five dimensions allows a unique torsional extension which is not in the Lovelock series~\cite{Mardones:1990qc},
 \begin{align}
   \label{boundary}
-  \mathcal{L} \propto \hat{T}_A\wedge \hat{R}^A_{\ B}\wedge\hat{e}^B,
+  \mathcal{L} \propto \hat{T}_A\wedge \hat{R}^A_{\ B}\wedge\hat{e}^B.
 \end{align}
-which can be written as a boundary term proportional to the Nieh-Yan invariant, adding no dynamics to the field equations.
-The action principle~\eqref{action5EGB}, along with the torsional term~\eqref{boundary}, describes the general Lovelock--Cartan theory in five dimensions.
+However, it can be written as a boundary term, adding no dynamics to the field equations.
 
@@
 \end{equation}
-This condition place us outside the Chern--Simons point since, otherwise, the field equations~\eqref{delta e} and~\eqref{delta w} remain invariant under $AdS_5$ gauge transformations, while the action principle~\eqref{action5EGB} can be recognized as its Chern--Simons form~\cite{Zanelli:2005sa,Troncoso:1999pk}.
+This condition place us outside the Chern--Simons point since, otherwise, the field equations~\eqref{delta e} and~\eqref{delta w} remain invariant under $AdS_5$ gauge transformations, while the action in Eq.~\eqref{action5EGB} can be recognized as its Chern--Simons form~\cite{Zanelli:2005sa,Troncoso:1999pk}.
 

### 2016-09-13 KKosmology.tex
Section III, discussed!

@@
 \usepackage{textcomp}
+\usepackage{float}
 \usepackage[%
@@
 \hypersetup{%
-  pdftitle={Cosmology from Kaluza--Klein reduction on five-dimensional Lovelock--Cartan theory},
+  pdftitle={Kaluza--Klein Cosmology from five-dimensional Lovelock--Cartan Theory},
   pdfauthor={Oscar Castillo-Felisola,}{Cristobal Corral,}{Simon del Pino,}{Francisca Ramirez.},
@@
 
-\title{Cosmology from  Kaluza--Klein reduction on five-dimensional Lovelock--Cartan Theory}
+\title{Kaluza--Klein Cosmology from five-dimensional Lovelock--Cartan Theory}
 
@@
 On the other hand, at cosmic scales, the experimental data obtained from Type Ia supernovae observations, indicates that our Universe is passing through a phase of accelerated expansion~\cite{Riess:1998cb}.
-This introduces the dark energy hypothesis to generate such a behavior.
+This introduces the dark energy hypothesis to generate such a behaviour.
 
@@
 
-The most general theory requiring the Lagrangian to be (i)~an invariant $N$-form under local Lorentz transformations; (ii)~a local polynomial of the vielbein, the Lorentz connection and their exterior derivatives; (iii)~constructed without the Hodge dual, is the so-called Lovelock--Cartan theory of gravity~\cite{Mardones:1990qc}.
-This theory is the natural generalization of the Lanczos--Lovelock action, when torsional degrees of freedom are present.
+The most general theory requiring the Lagrangian to be (i)~an invariant $N$-form under local Lorentz transformations; (ii)~a local polynomial of the vielbein, the Lorentz connection and their exterior derivatives; (iii)~constructed without the Hodge dual,\footnote{The Hodge dual maps $p$-forms into $(N-p)$-forms through $\star\left(\hat{e}^{A_1}\wedge ...
+\wedge\hat{e}^{A_p}\right) = \frac{1}{(N-p)!}\hat{\epsilon}^{A_1\ldots A_p}{}_{A_{p+1}...A_N}\,\hat{e}^{A_{p+1}}\wedge ...
+\wedge\hat{e}^{A_N}$.} is the Lovelock--Cartan theory of gravity~\cite{Mardones:1990qc}.
+This is the natural generalization of Lanczos--Lovelock when torsional degrees of freedom are present.
 
@@
     \hat{\epsilon}_{ABCDE}\Big(\alpha_1\hat{e}^C\wedge\hat{e}^D+
-    2\alpha_2\hat{R}^{CD}\Big)\wedge\hat{T}^E&=0,
+    2\alpha_2\hat{R}^{CD}\Big)\wedge\hat{T}^E&=0.
   \end{align}
-  respectively.
 \end{widetext}
+
+The last term in Eq.~\eqref{action5EGB} is the five-dimensional Gauss--Bonnet term for a Lorentz curvature written in exterior forms, which is analogue to Eq.~\eqref{GB} for the Riemannian case.
 
@@
   \label{boundary}
-  \mathcal{L} \propto \hat{T}_A\wedge \hat{R}^A_{\ B}\wedge\hat{e}^B.
+  \mathcal{L}_{T} \propto \hat{T}_A\wedge \hat{R}^A_{\ B}\wedge\hat{e}^B.
 \end{align}
@@
 \end{equation}
-This condition place us outside the Chern--Simons point since, otherwise, the field equations~\eqref{delta e} and~\eqref{delta w} remain invariant under $AdS_5$ gauge transformations, while the action in Eq.~\eqref{action5EGB} can be recognized as its Chern--Simons form~\cite{Zanelli:2005sa,Troncoso:1999pk}.
+This condition place us outside the Chern--Simons point.
+Otherwise, the field equations~\eqref{delta e} and~\eqref{delta w} are invariant under a larger gauge group ($AdS_5$), while the action~\eqref{action5EGB} becomes the Chern--Simons form for that group~\cite{Zanelli:2005sa,Troncoso:1999pk}.
 
-In terms of the KK ansatz presented in previous sections, the field equations can be decomposed into its distinctive parts.
-The equation of motion for the five-dimensional vielbein~\eqref{delta e}, leads to the following set of reduced field equations
+In terms of the KK ansatz presented in the previous section, the field equations can be decomposed into its distinctive parts, regarding
+its \(M_{4} \times S^1\) decomposition. The Eq.~\eqref{delta e} leads to 
 \begin{align}\label{equation}
@@
 \end{align}
-while the equation of motion for the spin connection~\eqref{delta w} gives
+while Eq.~\eqref{delta w} gives
 \begin{align}
@@
 isotropy and homogeneity of the involved fields.
-This can be achieved by considering the Lie derivative of each field, along the Killing vectors which generates the symmetries, to vanish.
-The Appendix~\ref{homotropic} is devoted to the details of how to reach the general ansatz we are presenting next.
-The $4$-dimensional vielbein compatible with the Friedman--Robertson--Walker metric is determined only by one function, the so-called scale factor $a(t)$, and their components read
+This can be achieved by imposing the Lie derivative of each field along the Killing vectors which generates the symmetries, to vanish.
+Appendix~\ref{homotropic} is devoted to the details of how to find the general ansatz.
+
+The four-dimensional vielbein compatible with the Friedman--Robertson--Walker metric is determined only by one function, the scale factor $a(t)$, and their components read
 \begin{align}
@@
 \end{align}
-where $k=+1,0,-1$ determines the spatial section of the four-dimensional manifold to be closed, flat or open, respectively.
+where $k=+1,0,-1$ determines the extrinsic curvature of the four-dimensional manifold to be closed, flat or open, respectively.
 
-The scalar field is determined by one time-dependent function, denoted by
+The scalar field is a time-dependent function, 
 \begin{equation}
-  \phi=\phi(t).
+  \phi=\phi(t),
 \end{equation}
+which is interpreted as the scale factor of the compact extra dimension.
 
@@
 
-For this ansatz, however, the $1$-form $A$ is $A_t(t)\text{d}t$, which can be shown to have vanishing field strength $F$.
+The $1$-form $A$ is $A(t)\text{d}t$, has vanishing field strength $F$.
 Thus, without loss of generality, it can be fixed to zero by means of an $U(1)$ gauge transformation or, equivalently, a diffeomorphism transformation along the $z$-direction.
@@
 Thus $b(t)=\beta(t)=0$.
-Otherwise, the system degenerates at the Chern--Simons point.
+Otherwise, the system is consistent only at the Chern--Simons point, where it degenerates.
 
@@
 \end{align}
-where dotted functions stand for its time-derivatives. For the sake of simplicity, we have defined $h(t)=\omega(t)-H(t)$, where $H=\dot{a}/a$ is the Hubble function.
+where dot stands for time-derivative. For the sake of simplicity, we have defined $h(t)=\omega(t)-H(t)$, where $H=\dot{a}/a$ is the Hubble function.
 
@@
 
-The system is integrated out developing two non-Riemannian branches, one for each value of the parameter $u_\pm$ defined to be
+The system is integrated and develops two non-Riemannian branches, one for each value of the parameter $u_\pm$ defined to be
 \begin{equation}\label{u}
@@
 where $\Delta$ was given in Eq.~\eqref{delta}.
+
 The Eqs.~\eqref{eqn1} to~\eqref{eqn2} are reduced to the Riemannian system when \mbox{$f=h=0$} and \mbox{$\gamma=-\dot{\phi}/2\phi$}.
@@
 
-Due to Eq.~\eqref{u}, the solutions will be valid in the region of parameter space where $\Delta>0$. The function $\omega(t)$ satisfies the equation
+Due to Eq.~\eqref{u}, the solutions will be valid in the region of the parameter space where $\Delta>0$. The function $\omega(t)$ satisfies the equation
 \begin{align}
@@
 where $t_0$ is an integration constant to be fixed.
-We relegate the time-dependence of the remaining fields to these expressions for $\omega$ and list their explicit time dependence in Appendix~\ref{solutions t}.
+We express the time-dependence of the remaining fields in terms of $\omega$ and list them explicitly in Appendix~\ref{solutions t}.
 The solutions read
@@
 
-For $u_\pm>0$ the solutions are periodic.
-In that case, the scale factor starts from a singular point and reaches a future one after a time $t_{\rm bounce}=\pi/\sqrt{u_\pm}$.
+For $u_\pm>0$ (see Fig.~\ref{uposreg}) the solutions are periodic.
+In that case, the scale factor $a(t)$ starts from a singular point and reaches a future one after a time $t_{\rm bounce}=\pi/\sqrt{u_\pm}$.
 Depending on the particular region on the parameter space, the scale factor undergoes an expanding and contracting age, allowing to reach a minimum value without collapsing, to expand again before a big crunch.
@@
 It also shows an inflationary phase at early times when the scale factor grows exponentially.
+The bouncing behaviour of the extra dimension is given by the scalar field in a phase difference of $\pi/2$ with respect to $a(t)$.
+
+\begin{figure}[H]
+  \includegraphics[scale=.45]{cond_u_pos.pdf}
+  \caption{The figure on the left denotes the region on the parameter space where $u_+>0$, while the figure on the right shows where $u_->0$. We have normalized $\alpha_1 = 1$.}
+  \label{uposreg}
+\end{figure}
+
+\begin{widetext}
+  \begin{figure}[H]
+    \includegraphics[scale=.45]{a_u_pos.pdf} \includegraphics[scale=.45]{phi_u_pos.pdf}
+    \caption{blablabla}
+    \label{upossol}
+  \end{figure}
+
+\end{widetext}
+
+% \begin{figure}[H]
+%  \includegraphics[scale=.45]{a_u_pos.pdf}
+% \label{aupos}
+% \end{figure}
+% 
+% \begin{figure}[H] 
+% \includegraphics[scale=.45]{phi_u_pos.pdf}
+%  \caption{These plots show the bouncing behaviour of the solutions in the distinctive regions of the parameter space.
+We have normalized each plot to its maximum value $a_{\rm max}$ and $\phi_{\rm max}$, respectively.}
+%  \label{phiupos}
+% \end{figure}
 
@@
 
-For $u_\pm\leq 0$, depending on the sign of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will infinitely expand while the scalar field reaches an asymtotic vanishing value for late cosmic times, after a period of expansion.
+For $u_\pm\leq 0$ (see Fig.~\ref{unegreg}), depending on the sign of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will infinitely expand while the scalar field reaches an asymtotic vanishing value for late cosmic times, after a period of expansion.
 This is interpreted as the late time behaviour of the radius of the compact manifold $S^1$, which for this family of solutions provides a natural explanation for its scale.
@@
 In the case of strictly negative $u_\pm$, for either contracting or expanding scale factors, their initial values remain finite and the solutions do not present an initial singularity.
-It has been reported that the GB term in a gravitational action can prevent the universe to expand from an initial singularity \cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if}, which is also the present case.
+It has been reported that the Gauss--Bonnet term can prevent the Universe to expand from an initial singularity \cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if}, which is also the present case.
 
-\subsubsection{Effective energy density and pressure}
+\begin{figure}[H]
+  \includegraphics[scale=.45]{cond_u_neg.pdf}
+  \caption{The figure on the left denotes the region on the parameter space where $u_+<0$, while the figure on the right shows where $u_-<0$. We have normalized $\alpha_1 = 1$.}
+  \label{unegreg}
+  \includegraphics[scale=.45]{a_u_neg.pdf} 
+  \includegraphics[scale=.45]{phi_u_neg.pdf}
+\end{figure}
+
+% \begin{figure}[H]
+%  \includegraphics[scale=.45]{a_u_neg.pdf} 
+%  \includegraphics[scale=.45]{phi_u_neg.pdf}
+%  \caption{Behaviour for $u_-<0$.}
+%  \label{unegsol}
+% \end{figure}
+
+\begin{figure}[H]
+  \includegraphics[scale=.45]{a_u-_0.pdf} 
+  \includegraphics[scale=.45]{phi_u-_0.pdf}
+  \caption{Behaviour for $u_- = 0$.}
+  \label{uzerosol}
+\end{figure}
+
+\subsection{Effective energy density and pressure}
 

### 2016-09-14 KKosmology.tex
Some modifications of the plots and their calls

@@
   f^2(\omega)&=\left(\omega^2+u_\pm\right)\frac{k}{a_0^2}\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2\left(\omega^2+u_\pm\right)}\right]\notag\\
-  &+\omega^2-Y_\pm,
+  &+\omega^2  + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2},
 \end{align}
-where we have defined the constant
-\begin{equation*}
-  Y_{\pm} = -\frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2},
-\end{equation*}
-and $a_0$ and $\phi_0$ are integration constants.
+where $a_0$ and $\phi_0$ are integration constants.
 
@@
 
-For $u_\pm>0$ (see Fig.~\ref{uposreg}) the solutions are periodic.
+For $u_\pm>0$ (see Fig.~\ref{cond_u_pos}) the solutions are periodic.
 In that case, the scale factor $a(t)$ starts from a singular point and reaches a future one after a time $t_{\rm bounce}=\pi/\sqrt{u_\pm}$.
@@
 \begin{figure}[H]
-  \includegraphics[scale=.45]{cond_u_pos.pdf}
+  \includegraphics[width=\linewidth]{cond_u_pos.pdf}
   \caption{The figure on the left denotes the region on the parameter space where $u_+>0$, while the figure on the right shows where $u_->0$. We have normalized $\alpha_1 = 1$.}
-  \label{uposreg}
+  \label{cond_u_pos}
 \end{figure}
-
-\begin{widetext}
 \begin{figure}[H]
-    \includegraphics[scale=.45]{a_u_pos.pdf} \includegraphics[scale=.45]{phi_u_pos.pdf}
+  \includegraphics[width=\linewidth]{a_u_pos.pdf}
   \caption{blablabla}
-    \label{upossol}
+  \label{a_u_pos}
 \end{figure}
-
-\end{widetext}
-
-% \begin{figure}[H]
-%  \includegraphics[scale=.45]{a_u_pos.pdf}
-% \label{aupos}
-% \end{figure}
-% 
-% \begin{figure}[H] 
-% \includegraphics[scale=.45]{phi_u_pos.pdf}
-%  \caption{These plots show the bouncing behaviour of the solutions in the distinctive regions of the parameter space.
-We have normalized each plot to its maximum value $a_{\rm max}$ and $\phi_{\rm max}$, respectively.}
-%  \label{phiupos}
-% \end{figure}
+\begin{figure}[H]
+  \includegraphics[scale=.45]{phi_u_pos.pdf}
+  \caption{Behavior of the $\phi$ field as a function of the scaled time, for $u_\pm > 0$.}
+  \label{phi_u_pos}
+\end{figure}
 
@@
 
-For $u_\pm\leq 0$ (see Fig.~\ref{unegreg}), depending on the sign of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will infinitely expand while the scalar field reaches an asymtotic vanishing value for late cosmic times, after a period of expansion.
+For $u_\pm\leq 0$ (see Fig.~\ref{cond_u_neg}), depending on the sign of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will infinitely expand while the scalar field reaches an asymtotic vanishing value for late cosmic times, after a period of expansion.
 This is interpreted as the late time behaviour of the radius of the compact manifold $S^1$, which for this family of solutions provides a natural explanation for its scale.
@@
   \caption{The figure on the left denotes the region on the parameter space where $u_+<0$, while the figure on the right shows where $u_-<0$. We have normalized $\alpha_1 = 1$.}
-  \label{unegreg}
+  \label{cond_u_neg}  
+\end{figure}
+\begin{figure}[H]
   \includegraphics[scale=.45]{a_u_neg.pdf}
+  \caption{}
+  \label{a_u_neg}
+\end{figure}
+\begin{figure}[H]
   \includegraphics[scale=.45]{phi_u_neg.pdf}
+  \caption{}
+  \label{phi_u_neg}
 \end{figure}
-
-% \begin{figure}[H]
-%  \includegraphics[scale=.45]{a_u_neg.pdf} 
-%  \includegraphics[scale=.45]{phi_u_neg.pdf}
-%  \caption{Behaviour for $u_-<0$.}
-%  \label{unegsol}
-% \end{figure}
 
@@
   \includegraphics[scale=.45]{a_u-_0.pdf} 
+  \caption{Behaviour for $u_- = 0$.}
+  \label{a_u_0}
+\end{figure}
+\begin{figure}[H]
   \includegraphics[scale=.45]{phi_u-_0.pdf}
   \caption{Behaviour for $u_- = 0$.}
-  \label{uzerosol}
+  \label{phi_u_0}
 \end{figure}
@@
   f^2(t)&=u_\pm\frac{k}{a_0^2}\sec^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cos^2\tau\right]\notag\\
-  &+u_\pm\tan^2\tau-Y_\pm.
+  &+u_\pm\tan^2\tau - \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
 \end{align}
@@
   f^2(t)&=\frac{k}{a_0^2(t-t_0)^2}\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2}(t-t_0)^2\right]\notag\\
-  &+(t-t_0)^{-2}-Y_\pm.
+  &+(t-t_0)^{-2} + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
 \end{align}
@@
   f^2(t)&=-u_\pm\frac{k}{a_0^2}\cosh^{-2}\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cosh^2\tau\right]\notag\\
-  &-u_\pm\tanh^2\tau-Y_\pm.
+  &-u_\pm\tanh^2\tau + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
 \end{align}

### 2016-09-14 KKosmology.tex
Bouncing solutions revised.

@@
 
+Hereon, we will consider $\alpha_1$ to be positive,\footnote{In all the plots, we fixed $\alpha_1 = 1$.} because this parameter admits the interpretation of the Newton's gravitational constant.
+
 \subsubsection{Bouncing solutions}
 
-For $u_\pm>0$ (see Fig.~\ref{cond_u_pos}) the solutions are periodic.
-In that case, the scale factor $a(t)$ starts from a singular point and reaches a future one after a time $t_{\rm bounce}=\pi/\sqrt{u_\pm}$.
-Depending on the particular region on the parameter space, the scale factor undergoes an expanding and contracting age, allowing to reach a minimum value without collapsing, to expand again before a big crunch.
-Otherwise, it reaches a maximum value before its collapse.
-It also shows an inflationary phase at early times when the scale factor grows exponentially.
-The bouncing behaviour of the extra dimension is given by the scalar field in a phase difference of $\pi/2$ with respect to $a(t)$.
-
+In Fig.~\ref{cond_u_pos} we show the allowed region, in the $\alpha_0$--$\alpha_2$ parameter space, for $u_\pm$ to be positive. 
 \begin{figure}[H]
   \includegraphics[width=\linewidth]{cond_u_pos.pdf}
-  \caption{The figure on the left denotes the region on the parameter space where $u_+>0$, while the figure on the right shows where $u_->0$. We have normalized $\alpha_1 = 1$.}
+  \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be positive defined. We have set $\alpha_1 = 1$.}
   \label{cond_u_pos}
 \end{figure}
+
+All the solutions for $u_\pm>0$ are periodic, as shown in the behavior of the scale factor in Fig.~\ref{a_u_pos}.
+In that case, the scale factor $a(t)$ starts from a singular point and reaches a future one after a time $t_{\rm bounce}=\pi/\sqrt{u_\pm}$.
+Depending on the particular region on the parameter space, the scale factor undergoes an expanding and contracting age, allowing a small bounce without collapsing, before a big crunch.
+This is the case of the third curve in Fig.~\ref{a_u_pos}.
+Otherwise, it expands and collapses in a \emph{simple} oscillatory way.
+One can also notice from the plot, that solutions parametrized by $u_+$ do not present inflationary phase, while those parametrized by $u_-$ do---the scale factor grows exponentially for a certain interval of time.
+Interestingly, the solutions with inflationary phases collapse in a damped form.
+Moreover, the bouncing behavior of the extra dimension is given by the scalar field in a phase difference of $\pi/2$ with respect to $a(t)$, as shown in Fig.~\ref{phi_u_pos}.
+It is worth to point that the \emph{small bounce} present in the scale factor is much prominent in the modulation of the extra dimension.
 \begin{figure}[H]
   \includegraphics[width=\linewidth]{a_u_pos.pdf}
-  \caption{blablabla}
+  \caption{Behavior of the scale factor, \(a(t)\), as a function of the scaled time, $\tau$, for $u_\pm > 0$.}
   \label{a_u_pos}
@@
   \includegraphics[scale=.45]{phi_u_pos.pdf}
-  \caption{Behavior of the $\phi$ field as a function of the scaled time, for $u_\pm > 0$.}
+  \caption{Behavior of the $\phi$ field as a function of the scaled time, $\tau$, for $u_\pm > 0$.}
   \label{phi_u_pos}
@@
 
-For $u_\pm\leq 0$ (see Fig.~\ref{cond_u_neg}), depending on the sign of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will infinitely expand while the scalar field reaches an asymtotic vanishing value for late cosmic times, after a period of expansion.
-This is interpreted as the late time behaviour of the radius of the compact manifold $S^1$, which for this family of solutions provides a natural explanation for its scale.
-Otherwise, in contrast to the previous behaviour, the scale factor contracts to an asymtotic vanishing value while the scalar field has no upper bound.
-In the case of strictly negative $u_\pm$, for either contracting or expanding scale factors, their initial values remain finite and the solutions do not present an initial singularity.
-It has been reported that the Gauss--Bonnet term can prevent the Universe to expand from an initial singularity \cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if}, which is also the present case.
+None of the other cases, $u_- = 0$ and $u_\pm < 0$, present bouncing behavior.
+The allowed moduli space for $u_+$ and $u_-$ to be negative is shown in Fig.~\ref{cond_u_neg}, while the restriction $u_- = 0$ sets $\alpha_2 = \tfrac{1}{12 \alpha_0}$.
+Nonetheless, depending of the values of the parameters there are three distinctive conduct: (i)~eternal expansion, (ii)~eternal contraction, and (iii)~initial expansion with a final collapse.
+
+The case $u_- = 0$ has two behaviors depending on whether $\alpha_0$ is positive or negative.
+As shown in Fig.~\ref{a_u_0}, for positive $\alpha_0$ the solutions fit into the third category above, while for negative $\alpha_0$ the solutions expand eternally.
+Notice that in the former case the Universe's fate is an asymptotic collapse, whilst in the later expands forever but shows an interesting evolution of the Hubble constant.
+\begin{figure}[H]
+  \includegraphics[scale=.45]{a_u-_0.pdf} 
+  \caption{Behaviour for $u_- = 0$.}
+  \label{a_u_0}
+\end{figure}
+\begin{figure}[H]
+  \includegraphics[scale=.45]{phi_u-_0.pdf}
+  \caption{Behaviour for $u_- = 0$.}
+  \label{phi_u_0}
+\end{figure}
+
+For $u_\pm\leq 0$ (see Fig.~\ref{cond_u_neg}), depending on the sign
+of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will
+infinitely expand while the scalar field reaches an asymtotic
+vanishing value for late cosmic times, after a period of
+expansion. This is interpreted as the late time behaviour of the
+radius of the compact manifold $S^1$, which for this family of
+solutions provides a natural explanation for its scale. Otherwise, in
+contrast to the previous behaviour, the scale factor contracts to an
+asymtotic vanishing value while the scalar field has no upper
+bound. In the case of strictly negative $u_\pm$, for either
+contracting or expanding scale factors, their initial values remain
+finite and the solutions do not present an initial singularity. It has
+been reported that the Gauss--Bonnet term can prevent the Universe to
+expand from an initial
+singularity~\cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if},
+which is also the present case.
 
@@
   \includegraphics[scale=.45]{cond_u_neg.pdf}
-  \caption{The figure on the left denotes the region on the parameter space where $u_+<0$, while the figure on the right shows where $u_-<0$. We have normalized $\alpha_1 = 1$.}
+  \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be positive defined. We have normalized $\alpha_1 = 1$.}
   \label{cond_u_neg}  
@@
   \label{phi_u_neg}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[scale=.45]{a_u-_0.pdf} 
-  \caption{Behaviour for $u_- = 0$.}
-  \label{a_u_0}
-\end{figure}
-\begin{figure}[H]
-  \includegraphics[scale=.45]{phi_u-_0.pdf}
-  \caption{Behaviour for $u_- = 0$.}
-  \label{phi_u_0}
 \end{figure}

### 2016-09-15 KKosmology.tex
Sections II and IV. Discussed!

--- 2016-09-14 KKosmology.tex
@@
   pdftitle={Kaluza--Klein Cosmology from five-dimensional Lovelock--Cartan Theory},
-  pdfauthor={Oscar Castillo-Felisola,}{Cristobal Corral,}{Simon del Pino,}{Francisca Ramirez.},
+  pdfauthor={Oscar Castillo-Felisola,}{Cristobal Corral,}{Simon del Pino.},
   pdfkeywords={Torsion,} {Cosmology,}{Kaluza--Klein reduction,}{Generalised Gravity.},
@@
 
-\author{Francisca \surname{Ram\'irez}.}
-\affiliation{\UTFSM.}
+%% \author{Francisca \surname{Ram\'irez}.}
+%% \affiliation{\UTFSM.}
 
@@
   We study the Kaluza--Klein dimensional reduction of the Lovelock--Cartan theory on a five-dimensional manifold, where the compact dimension has $S^1$ topology.
-We find cosmological solutions of the Friedman--Robertson--Walker class for the effective theory.
+We find cosmological solutions of the Friedmann--Robertson--Walker class for the effective theory.
   %%In the five-dimensional vacuum, this theory allows a nontrivial torsion in contrast to the Einstein--Cartan theory.
@@
 On the other hand, at cosmic scales, the experimental data obtained from Type Ia supernovae observations, indicates that our Universe is passing through a phase of accelerated expansion~\cite{Riess:1998cb}.
-This introduces the dark energy hypothesis to generate such a behaviour.
+This introduces the dark energy hypothesis to generate such a behavior.
 
@@
 On the one hand, classically, wormholes must be supported by a kind energy compatible with the cosmological hypothesis.
-Their existence in these extended models, implies the presence of degrees of freedom that could provide an explanation for the exotic matter/energy abundance in the Universe.
+Their existence in extended models, implies the presence of degrees of freedom that could provide an explanation for the exotic matter/energy abundance in the Universe.
 In the Riemannian five-dimensional Lanczos--Lovelock theory, exact wormholes solutions have been found in vacuum~\cite{Dotti:2006cp,Dotti:2007az}.
@@
 
-\subsection{Bianchi identities}
+\subsection{Bianchi identities\label{sec:bianchi}}
 
@@
 
-The four-dimensional vielbein compatible with the Friedman--Robertson--Walker metric is determined only by one function, the scale factor $a(t)$, and their components read
+The four-dimensional vielbein compatible with the Friedmann--Robertson--Walker metric is determined only by one function, the scale factor $a(t)$, and their components read
 \begin{align}
@@
 
+Hereon, we will consider $\alpha_1$ to be positive,\footnote{In all the plots, we fixed $\alpha_1 = 1$.} because this parameter admits the interpretation of the Newton's gravitational constant.
+
 \subsubsection{Bouncing solutions}
 
-For $u_\pm>0$ (see Fig.~\ref{cond_u_pos}) the solutions are periodic.
-In that case, the scale factor $a(t)$ starts from a singular point and reaches a future one after a time $t_{\rm bounce}=\pi/\sqrt{u_\pm}$.
-Depending on the particular region on the parameter space, the scale factor undergoes an expanding and contracting age, allowing to reach a minimum value without collapsing, to expand again before a big crunch.
-Otherwise, it reaches a maximum value before its collapse.
-It also shows an inflationary phase at early times when the scale factor grows exponentially.
-The bouncing behaviour of the extra dimension is given by the scalar field in a phase difference of $\pi/2$ with respect to $a(t)$.
-
+In Fig.~\ref{cond_u_pos} we show the allowed region, in the $\alpha_0$--$\alpha_2$ parameter space, for $u_\pm$ to be positive. 
 \begin{figure}[H]
   \includegraphics[width=\linewidth]{cond_u_pos.pdf}
-  \caption{The figure on the left denotes the region on the parameter space where $u_+>0$, while the figure on the right shows where $u_->0$. We have normalized $\alpha_1 = 1$.}
+  \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be positive defined.}
   \label{cond_u_pos}
 \end{figure}
+
+All the solutions for $u_\pm>0$ are periodic, as shown in the behavior of the scale factor in Fig.~\ref{a_u_pos}.
+In that case, the scale factor $a(t)$ starts from a singular point and reaches a future one after a time $t_{\rm bounce}=\pi/\sqrt{u_\pm}$.
+Depending on the particular region on the parameter space, the scale factor undergoes an expanding and contracting age, allowing a intermediate bounce without collapsing, before a big crunch.
+This is the case of the third curve in Fig.~\ref{a_u_pos}.
+Otherwise, it expands and collapses in a {simple} oscillatory way.
+\hl{One can also notice from the plot, that solutions parametrized by $u_+$ do not present inflationary phase, while those parametrized by $u_-$ do---the scale factor grows exponentially for a certain interval of time.
+Interestingly, the solutions with inflationary phases collapse in a damped form.}
+Moreover, the bouncing behavior of the extra dimension is given by the scalar field in a phase difference of $\pi/2$ with respect to $a(t)$, as shown in Fig.~\ref{phi_u_pos}.
+It is worth to point that the {intermediate bounce} present in the scale factor is much prominent in the modulation of the extra dimension.
 \begin{figure}[H]
   \includegraphics[width=\linewidth]{a_u_pos.pdf}
-  \caption{blablabla}
+  \caption{Behavior of the scale factor, \(a(t)\), as a function of the scaled time, $\tau$, for $u_\pm > 0$.}
   \label{a_u_pos}
@@
 \begin{figure}[H]
-  \includegraphics[scale=.45]{phi_u_pos.pdf}
-  \caption{Behavior of the $\phi$ field as a function of the scaled time, for $u_\pm > 0$.}
+  \includegraphics[width=\linewidth]{phi_u_pos.pdf}
+  \caption{Behavior of the $\phi$ field as a function of the scaled time, $\tau$, for $u_\pm > 0$.}
   \label{phi_u_pos}
@@
 
-For $u_\pm\leq 0$ (see Fig.~\ref{cond_u_neg}), depending on the sign of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will infinitely expand while the scalar field reaches an asymtotic vanishing value for late cosmic times, after a period of expansion.
-This is interpreted as the late time behaviour of the radius of the compact manifold $S^1$, which for this family of solutions provides a natural explanation for its scale.
-Otherwise, in contrast to the previous behaviour, the scale factor contracts to an asymtotic vanishing value while the scalar field has no upper bound.
-In the case of strictly negative $u_\pm$, for either contracting or expanding scale factors, their initial values remain finite and the solutions do not present an initial singularity.
-It has been reported that the Gauss--Bonnet term can prevent the Universe to expand from an initial singularity \cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if}, which is also the present case.
+None of the other cases, $u_- = 0$ and $u_\pm < 0$, present bouncing behavior.
+The allowed parameter space for $u_+$ and $u_-$ to be negative is shown in Fig.~\ref{cond_u_neg}, while the restriction $u_- = 0$ sets $\alpha_2 = \tfrac{1}{12 \alpha_0}$.
+Nonetheless, depending on the values of the parameters there are three distinctive conducts: (i)~eternal expansion, (ii)~eternal contraction, and (iii)~initial expansion with a final collapse.
+
+\paragraph{The case $u_- = 0$} has two behaviors depending on whether $\alpha_0$ is positive or negative.
+
+As shown in Fig.~\ref{a_u_0}, for positive $\alpha_2$ the solutions fit into the third category above, while for negative $\alpha_2$ the solutions expand eternally.
+Notice that in the former case the Universe's fate is an asymptotic collapse, whilst in the latter expands forever.
+\begin{figure}[H]
+  \includegraphics[width=\linewidth]{a_u-_0.pdf} 
+  \caption{Distinctive behaviors of the scale factor for $u_- = 0$.}
+  \label{a_u_0}
+\end{figure}
+\begin{figure}[H]
+  \includegraphics[width=\linewidth]{phi_u-_0.pdf}
+  \caption{Distinctive behavior of the $\phi$ field for $u_- = 0$.}
+  \label{phi_u_0}
+\end{figure}
+The $\phi$ field (see Fig.~\ref{phi_u_0}) grows infinitely for $\alpha_2 > 0$, or asymptotically goes to zero for $\alpha_2 < 0$.
+This latter behavior provides a {dynamical} compactification, which might serve as a mechanism to assure---at certain time---the decoupling of the zeroth mode from the Kaluza--Klein tower.
+
+\paragraph{The case $u_\pm < 0$} has solutions which either expand or collapse eternally, as shown in Fig.~\ref{a_u_neg}.
+The typical evolution of the Universe with $u_+ > 0$ is to grow infinitely, while for negative $u_-$ the expansion (contraction) corresponds to $\alpha_2$ positive (negative).
+Figure~\ref{phi_u_neg}, shows that the asymptotic {size} of the extra dimension, modulated by the $\phi$ field, is approximately reciprocal to the behavior of the scale factor but it is not monotonic.
+%% , depending on the sign
+%% of $\alpha_2$ and the $\pm$-sign in $u_\pm$, the scale factor will
+%% infinitely expand while the scalar field reaches an asymtotic
+%% vanishing value for late cosmic times, after a period of
+%% expansion. This is interpreted as the late time behavior of the
+%% radius of the compact manifold $S^1$, which for this family of
+%% solutions provides a natural explanation for its scale. Otherwise, in
+%% contrast to the previous behavior, the scale factor contracts to an
+%% asymtotic vanishing value while the scalar field has no upper
+%% bound.
+For all these solutions the scale factor remains finite at $t = 0$, and they do not present an initial singularity.
+It has been reported that the Gauss--Bonnet term can prevent the Universe to expand from an initial singularity~\cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if}, which is also this case.
 
 \begin{figure}[H]
-  \includegraphics[scale=.45]{cond_u_neg.pdf}
-  \caption{The figure on the left denotes the region on the parameter space where $u_+<0$, while the figure on the right shows where $u_-<0$. We have normalized $\alpha_1 = 1$.}
+  \includegraphics[width=\linewidth]{cond_u_neg.pdf}
+  \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be positive defined.}
   \label{cond_u_neg}  
-\end{figure}
-\begin{figure}[H]
-  \includegraphics[scale=.45]{a_u_neg.pdf}
-  \caption{}
-  \label{a_u_neg}
-\end{figure}
-\begin{figure}[H]
-  \includegraphics[scale=.45]{phi_u_neg.pdf}
-  \caption{}
-  \label{phi_u_neg}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[scale=.45]{a_u-_0.pdf} 
-  \caption{Behaviour for $u_- = 0$.}
-  \label{a_u_0}
+  \includegraphics[width=.95\linewidth]{a_u_neg.pdf}
+  \caption{Different behaviors for the scale factor, as function of the scaled time $\tau$, for $u_\pm < 0$.}
+  \label{a_u_neg}
 \end{figure}
+
+%\pagebreak
 \begin{figure}[H]
-  \includegraphics[scale=.45]{phi_u-_0.pdf}
-  \caption{Behaviour for $u_- = 0$.}
-  \label{phi_u_0}
+  \includegraphics[width=.95\linewidth]{phi_u_neg.pdf}
+  \caption{Different behaviors for the $\phi$ field, as function of the scaled time $\tau$, for $u_\pm < 0$.}
+  \label{phi_u_neg}
 \end{figure}
@@
 
-The Eq.~\eqref{equation} can be written, by means of the identity~\eqref{curvature decomp}, in the familiar form
+The Eq.~\eqref{equation} can be split into its Riemannian and non-Riemannian parts, by means of the identity~\eqref{curvature decomp}, taking the familiar form
 \begin{align}\label{einstein equation}
-  -\frac{1}{2}\epsilon_{abcd}\left(\tilde{R}^{bc} - \frac{\Lambda}{3}\,e^b\wedge e^c\right)e^d &= \kappa_{G}\,\tau^{eff}_a,
+  -\frac{1}{2}\epsilon_{abcd}\left(\tilde{R}^{bc} - \frac{\Lambda}{3}\,e^b\wedge e^c\right)e^d &= \kappa_{G}\,\tau^{\mathrm{eff}}_a,
 \end{align}
-where $\Lambda \equiv -6\alpha_0/\alpha_1$, $\kappa_{G} \equiv 2/\alpha_1$ and we have defined the effective energy-momentum $3$-form as
+with $\Lambda \equiv -6\alpha_0/\alpha_1$, $\kappa_{G} \equiv 2/\alpha_1$, and the effective energy-momentum $3$-form defined as
 \begin{align}
-  \tau_a^{eff}&= \frac{1}{2}\epsilon_{abcd}\bigg[\frac{1}{2}\alpha_1\left(\kappa^b_{\ l}\wedge\kappa^{lc}+\tilde{\text{D}}\kappa^{bc}\right)\wedge e^d\notag\\
-    &+\left(\text{D}\gamma^b+\frac{1}{2}\gamma^b\text{d}\ln\phi\right)\wedge\left(\frac{1}{2}\alpha_1e^c\wedge e^d+\alpha_2R^{cd}\right)\bigg].
+  \tau_a^{\mathrm{eff}}&= \tfrac{1}{2}\epsilon_{abcd}\Big[\tfrac{1}{2}\alpha_1\left(\kappa^b_{\ l}\wedge\kappa^{lc}+\tilde{\text{D}}\kappa^{bc}\right)\wedge e^d\\
+    & \quad +\left(\text{D}\gamma^b+\tfrac{1}{2}\gamma^b\text{d}\ln\phi\right)\wedge\left(\tfrac{1}{2}\alpha_1e^c\wedge e^d+\alpha_2R^{cd}\right)\Big]. \notag
 \end{align}
-The contributions of $F$, $\alpha^{ab}$ and $\beta^a$ have not been taken into account since for all practical purposes of this article, they vanish.
-The left hand side of the Eq.~\eqref{einstein equation} is the Eintein $3$-form, which is dual to the Einstein tensor with cosmological constant in four dimensions.
-We identify its right hand side as an effective energy-momentum $3$-form, which behaves as a perfect fluid.
-Such an energy-momentum contains the torsional and the higher dimensional degrees of freedom as part of the gravitating fields in four dimensions.
-From this, we identify the energy density $\rho$ and the pressure $p$ through
+The contributions of $F$, $\alpha^{ab}$ and $\beta^a$ have not been taken into account since, for all practical purposes of this article, they vanish.
+%% The left hand side of the Eq.~\eqref{einstein equation} is the Einstein $3$-form, which is dual to the Einstein tensor with cosmological constant in four dimensions.
+The left hand side of Eq.~\eqref{einstein equation} is the 3-form which, in components, yields the four-dimensional Einstein equations with cosmological constant.
+We identify the right hand side as an effective energy-momentum $3$-form, which behaves as a perfect fluid. It contains the torsional and the higher dimensional degrees of freedom.
+
+The energy density $\rho$ and the pressure $p$ are obtained through the identities
 \begin{align}
-  \tau_0^{eff}&=-\frac{1}{3!}\,\rho\,\epsilon_{0ijk}e^i\wedge e^j\wedge e^k,\\
-  \tau_i^{eff}&=-\frac{1}{2}\,p\,\epsilon_{0ijk}e^0\wedge e^j\wedge e^k,
+  \tau_0^{\mathrm{eff}} &= -\frac{1}{3!} \, \rho \, \epsilon_{0ijk} \; e^i\wedge e^j\wedge e^k,\\
+  \tau_i^{\mathrm{eff}} &= -\frac{1}{2} \,p \, \epsilon_{0ijk} \; e^0\wedge e^j\wedge e^k,
 \end{align} 
-where
-\begin{align}\label{rho}
-  \rho&=-\frac{3}{2}\alpha_1\left(h^2-f^2+2Hh\right)\notag\\
-  &+3\omega\gamma\left[\frac{1}{2}\alpha_1+\alpha_2\left(\omega^2+\frac{k}{a^2}-f^2\right)\right],\\
+giving
+\begin{align}
+  \label{rho}
+  \rho &= -\frac{3}{2}\alpha_1\left(h^2-f^2+2Hh\right) \\
+  & \quad + 3\omega\gamma\left[\frac{1}{2}\alpha_1+\alpha_2\left(\omega^2+\frac{k}{a^2}-f^2\right)\right], \notag
+  \\
   \label{presion}
-  p&=\alpha_1\left[\dot{h}+Hh+\frac{1}{2}\left(h^2-f^2+2Hh\right)\right]\notag\\
-  &-2\omega\gamma\left[\frac{1}{2}\alpha_1+\alpha_2\left(\dot{\omega}+H\omega\right)\right]\notag\\
-  &+\left(\dot{\gamma}+\frac{1}{2}\gamma\dot{\Phi}\right)\left[\frac{1}{2}\alpha_1+\alpha_2\left(\omega^2+\frac{k}{a^2}-f^2\right)\right].
+  p &= \alpha_1\left[\dot{h}+Hh+\frac{1}{2}\left(h^2-f^2+2Hh\right)\right]\\
+  & \quad - 2\omega\gamma\left[\frac{1}{2}\alpha_1+\alpha_2\left(\dot{\omega}+H\omega\right)\right]\notag\\
+  & \quad +\left(\dot{\gamma}+\frac{1}{2}\gamma\dot{\Phi}\right)\left[\frac{1}{2}\alpha_1+\alpha_2\left(\omega^2+\frac{k}{a^2}-f^2\right)\right]. \notag
 \end{align}
-Using the solutions obtained in the previous section , it can be shown that
+Using the solutions obtained in the previous section, these expressions are
 \begin{align}
+  \label{our-rho}
   \rho &= \frac{3}{2}\alpha_1\left(H^2 + \frac{k}{a^2}\right) + 3\alpha_0, \\
+  \label{our-p}
   p &= -\alpha_1\left(\dot{H} + \frac{3}{2}\,H^2 + \frac{k}{2a^2}\right) - 3\alpha_0,
 \end{align}
-satisfying the continuity equation $\dot{\rho}+3H\left(\rho+p\right)=0$.
+and satisfy the continuity equation $\dot{\rho}+3H\left(\rho+p\right)=0$.
 
 For $u_\pm<0$, the energy density remains finite at the begining.
-From expressions \eqref{rho} and \eqref{presion} one can see that the induced energy density and pressure are not positive definite quantities.
+From the Eqs.~\eqref{our-rho} and~\eqref{our-p} one can see that the induced energy density and pressure are not positive definite quantities.
 In fact, the Universe undergoes through an accelerating expansion phase due to the presence of torsion and the extra dimensional fields.
@@
 
-It is known that KK theory, in its simplest version, gives rise to matter fields as a manifestation of the extra dimensional metric degrees of freedom.
+In this paper we have presented the dimensional reduction of the five-dimensional Lovelock--Cartan theory introduced in Ref.~\cite{Mardones:1990qc}, under the assumption that the compact dimension has $S^1$ topology.
+An interesting feature of these theories is that, unlike the Einstein--Cartan theory, torsion is allowed to propagate.
+
+In the generic reduction, the effective theory has a spin-2 particle, a spin-1 $U(1)$ gauge boson, and a spin-0 scalar particle coming from the decomposition of the metric, while the decomposition of the spin connection introduces three extra fields: two spin-1 particles, $\alpha^{ab}$ and $\gamma^a$, and the 1-form $\beta^a$ yields a new spin-2, a spin-1 and a spin-0 particles.
+These new fields do not transform under the gauge group (see Sec.~\ref{sec:bianchi}).
+
+We used the most general ansatz compatible with the cosmological symmetries (see Appendix~\ref{homotropic}), to find solutions of the Friedmann--Robertson--Walker class in the effective theory.
+The equations of motion [Eqs.~\eqref{eqn1}--\eqref{eqn2}] ensures that solutions with nontrivial torsion have vanishing $\alpha^{ab}$ and $\beta^a$.
+
+The solutions are parametrized by a variable $u_\pm$, related to the fundamental couplings of the theory through the Eq.~\eqref{u}, and there are three different sectors depending on whether this parameter is positive, negative or zero.
+
+The behavior of the Universes described by our solutions is four-fold: (i)~eternal expansion, (ii)~eternal contraction, (iii)~initial expansion with a final collapse, and (iv)~bouncing.
+We show the typical evolution of the scale factor, for each case, in Figs.~\ref{a_u_pos}, ~\ref{a_u_0} and~\ref{a_u_neg}.
+The size of the extra dimension is dynamical, and there are sectors in the parameter space where there exists an asymptotic dynamical mechanics of compactification, see Figs.~\ref{phi_u_0} and~\ref{phi_u_neg}.
+
+%% It is known that KK theory, in its simplest version, gives rise to matter fields as a manifestation of the extra dimensional metric degrees of freedom.
 The kind of matter arising in the reduced spacetime is subordinated to the topology taken for the extra compact manifold.
-In general, those fields will gravitate and modulate the behaviour of spacetime in its time evolution.
+In general, those fields will gravitate and modulate the behavior of spacetime in its time evolution.
 
-We find cosmological solutions of the Friedmann--Robertson--Walker class for the dimensionally reduced Lovelock--Cartan theory in five dimensions~\cite{Mardones:1990qc}.
+%% We find cosmological solutions of the Friedmann--Robertson--Walker class for the dimensionally reduced Lovelock--Cartan theory in five dimensions~\cite{Mardones:1990qc}.
 This theory allows torsion to propagate in vacuum, in contrast to the Einstein--Cartan theory.
@@
 \begin{align*}
-  \mathcal{L}_{\text{Interaction}}&=-\frac{i}{8}T_{\alpha\mu\nu}\bar{\Psi}\Gamma^{\alpha\mu\nu}\Psi =\frac{3}{2}f\bar{\Psi}\Gamma^0\Gamma^5\Psi
+  \mathcal{L}_{\text{Int}}&=-\frac{i}{8}T_{\alpha\mu\nu}\bar{\Psi}\Gamma^{\alpha\mu\nu}\Psi =\frac{3}{2}f\bar{\Psi}\Gamma^0\Gamma^5\Psi
 \end{align*}
@@
 Similar results where found in~\cite{Toloza:2013wi} for another cosmological model with torsion which shares some of the couplings of the reduced action.
+
 \begin{acknowledgments}
-  The authors would like to thank O. Miskovic, A. Toloza and J. Zanelli for enlightening discussions in this work.
+  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, O.~Miskovic, A.~Toloza and J.~Zanelli for enlightening discussions in this work.
+This work has been partially supported by the PAI project No.~79140040 (CONICYT--Chile).
 \end{acknowledgments}
@@
   f^2(t)&=u_\pm\frac{k}{a_0^2}\sec^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cos^2\tau\right]\notag\\
-  &+u_\pm\tan^2\tau - \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
+  &+u_\pm\tan^2\tau + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
 \end{align}

### 2016-09-15 KKosmology.tex
Finished the Section V.

@@
 \begin{align}
+  \label{taua}
   \tau_a^{\mathrm{eff}}&= \tfrac{1}{2}\epsilon_{abcd}\Big[\tfrac{1}{2}\alpha_1\left(\kappa^b_{\ l}\wedge\kappa^{lc}+\tilde{\text{D}}\kappa^{bc}\right)\wedge e^d\\
@@
 The size of the extra dimension is dynamical, and there are sectors in the parameter space where there exists an asymptotic dynamical mechanics of compactification, see Figs.~\ref{phi_u_0} and~\ref{phi_u_neg}.
+
+After the decomposition into Riemannian and non-Riemannian parts of the fields, the Eq.~\eqref{equation} takes the form of the usual Einstein's equations where the extra fields, coming from the extra dimensional and torsional degrees of freedom, originate an effective energy-momentum tensor.
+It is important to mention that this tensor is nonstandard, in the sense that it inherit a coupling between matter and curvature---see Eq.~\eqref{taua} for the definition---, due to the non-minimal coupling among the four-dimensional gravity and the ``matter'' fields.
+Moreover, the energy density and the pressure are not positive definite, which allows us to interpret these induced matter as exotic, providing a playground for finding gravitational explanation to dark matter/energy.
 

### 2016-09-16 KKosmology.tex
Additional modifications by Criz and Simon

--- 2016-09-15 KKosmology.tex
@@
 have driven the community to think that GR is not the ultimate gravitational theory. 
-The \emph{dark sector} of the Universe is composed by two different forms of ``matter'', whose effects are detected at different scales.
+The \emph{dark sector} of the Universe is composed by two kinds of degrees of freedom, whose effects are detected at different scales.
 On the one hand, at galactic scales, the gravitational lensing produced by the local distribution of energy, suggests the presence of an exotic form of matter, unseen by the current light-based telescopes~\cite{Sofue:2000jx}, for this it is named dark matter.
@@
 Hereon, we will consider $\alpha_1$ to be positive,\footnote{In all the plots, we fixed $\alpha_1 = 1$.} because this parameter admits the interpretation of the Newton's gravitational constant.
+It is worth mentioning the existence of models of gravity which are free of the linear curvature term.
+These models are refer to as Pure Lovelock (PL) gravities~\cite{Cai:2006pq} and consider the cosmological term and the polynomial of highest order in the curvature.
+Even though the analysis that follows will consider a positive non-zero value for $\alpha_1$, the solutions listed above are also valid in the regime where $\alpha_1=0$, that constitutes the non-Riemannian extension of PL theories.
 
@@
 Otherwise, it expands and collapses in a {simple} oscillatory way.
-\hl{One can also notice from the plot, that solutions parametrized by $u_+$ do not present inflationary phase, while those parametrized by $u_-$ do---the scale factor grows exponentially for a certain interval of time.
+%\hl{One can also notice from the plot, that solutions parametrized by $u_+$ do not present inflationary phase, while those parametrized by $u_-$ do---the scale factor grows exponentially for a certain interval of time.
 Interestingly, the solutions with inflationary phases collapse in a damped form.}
@@
 \begin{align}
+  \label{taua}
   \tau_a^{\mathrm{eff}}&= \tfrac{1}{2}\epsilon_{abcd}\Big[\tfrac{1}{2}\alpha_1\left(\kappa^b_{\ l}\wedge\kappa^{lc}+\tilde{\text{D}}\kappa^{bc}\right)\wedge e^d\\
-    & \quad +\left(\text{D}\gamma^b+\tfrac{1}{2}\gamma^b\text{d}\ln\phi\right)\wedge\left(\tfrac{1}{2}\alpha_1e^c\wedge e^d+\alpha_2R^{cd}\right)\Big]. \notag
+    & \quad +\left(\text{D}\gamma^b+\tfrac{1}{2}\gamma^b\text{d}\ln\phi\right)\wedge\left(\tfrac{1}{2}\alpha_1e^c\wedge e^d+\alpha_2R^{cd}\right)\Big], \notag
 \end{align}
+where $\kappa^{ab}$ is the contorsion $1$-form defined as the torsional correction to the spin connection (see Eq.~\eqref{spin separation}).
 The contributions of $F$, $\alpha^{ab}$ and $\beta^a$ have not been taken into account since, for all practical purposes of this article, they vanish.
@@
 
-We used the most general ansatz compatible with the cosmological symmetries (see Appendix~\ref{homotropic}), to find solutions of the Friedmann--Robertson--Walker class in the effective theory.
+We used the most general ansatz compatible with the cosmological symmetries (see Appendix~\ref{homotropic}) to find solutions of the Friedmann--Robertson--Walker class in the effective theory.
 The equations of motion [Eqs.~\eqref{eqn1}--\eqref{eqn2}] ensures that solutions with nontrivial torsion have vanishing $\alpha^{ab}$ and $\beta^a$.
 
-The solutions are parametrized by a variable $u_\pm$, related to the fundamental couplings of the theory through the Eq.~\eqref{u}, and there are three different sectors depending on whether this parameter is positive, negative or zero.
+The solutions are parametrized by a variable $u_\pm$, related to the fundamental couplings of the theory through the Eq.~\eqref{u}, and exhibit three different sectors depending on whether this parameter is positive, negative or zero.
 
@@
 We show the typical evolution of the scale factor, for each case, in Figs.~\ref{a_u_pos}, ~\ref{a_u_0} and~\ref{a_u_neg}.
-The size of the extra dimension is dynamical, and there are sectors in the parameter space where there exists an asymptotic dynamical mechanics of compactification, see Figs.~\ref{phi_u_0} and~\ref{phi_u_neg}.
+The size of the extra dimension is dynamical, and there are sectors in the parameter space where there exists an asymptotic dynamical mechanics of compactification (see Figs.~\ref{phi_u_0} and~\ref{phi_u_neg}).
+
+After the decomposition of the curvature into Riemannian and non-Riemannian parts, the Eq.~\eqref{equation} takes the form of the usual Einstein's equations where the extra fields, coming from the extra dimensional and torsional degrees of freedom, originate an effective energy-momentum tensor.
+It is important to mention that this tensor is nonstandard, in the sense that it inherit a coupling between matter and curvature---see Eq.~\eqref{taua} for the definition---, due to the non-minimal coupling among the four-dimensional gravity and the induced fields.
+Moreover, the energy density and the pressure are not positive definite, which allows us to interpret these induced matter as exotic, providing a playground for finding gravitational explanation to dark matter/energy.
 
@@
 \begin{align*}
-  \mathcal{L}_{\text{Int}}&=-\frac{i}{8}T_{\alpha\mu\nu}\bar{\Psi}\Gamma^{\alpha\mu\nu}\Psi =\frac{3}{2}f\bar{\Psi}\Gamma^0\Gamma^5\Psi
+  \mathcal{L}_{\text{Int}}&=-\frac{i}{8}T_{\alpha\mu\nu}\bar{\Psi}\Gamma^{\alpha\mu\nu}\Psi =\frac{3}{2}f\bar{\Psi}\Gamma^0\Gamma^5\Psi.
 \end{align*}
@@
 For imaginary values of $f(t)$, the corresponding violation of current conservation can be interpreted as particle creation.
-Similar results where found in~\cite{Toloza:2013wi} for another cosmological model with torsion which shares some of the couplings of the reduced action.
+Similar results where found in~\cite{Toloza:2013wi} for other cosmological model with torsion which shares some of the couplings of a reduced action.
 
 \begin{acknowledgments}
-  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, O.~Miskovic, A.~Toloza and J.~Zanelli for enlightening discussions in this work.
-This work has been partially supported by the PAI project No.~79140040 (CONICYT--Chile).
+  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, A.~Toloza and J.~Zanelli for enlightening discussions on this work.
+O.~C-F.
+has been partially supported by the PAI project No.~79140040 (CONICYT--Chile).
+The Centro Cient\'ifico y Tecnol\'ogico de Valpara\'iso (CCTVal) is funded by the Chilean Goverment through the Centers of Excellence Base Financing Program of CONICYT.
 \end{acknowledgments}
@@
 
-In Riemann--Cartan geometry, both the vielbein $e^A$ and the spin connection $\omega^{AB}$ are independent fields.
+In Riemann--Cartan geometry, both the vielbein $e^A$ and the spin connection $\omega^{AB}$ are independent features of the manifold.
 We will drop the hats since we are not dealing with the KK decomposition.
-Stationary variations of a gravitational action principle must be taken with respect to both fields, allowing in general the presence of torsion.
-The spin connection, however, can be decomposed in a Riemannian part $\tilde{\omega}^{AB}$ satisfying $\tilde{\mbox{D}}e^A=0$, and a contorsion piece $\kappa^{AB}=-\kappa^{BA}$, such that
+The spin connection, however, can be decomposed in a Riemannian part, which is metric dependent, $\tilde{\omega}^{AB}$ satisfying $\tilde{\mbox{D}}e^A=0$, and a contorsion piece $\kappa^{AB}=-\kappa^{BA}$, such that
 \begin{equation}\label{spin separation}
@@
 \end{equation}
-where $\tilde{R}^{AB}$ is the Riemannian curvature built up from $\tilde{\omega}^{AB}$.
+where $\tilde{R}^{AB}$ is the Riemannian curvature constructed with $\tilde{\omega}^{AB}$.
 
@@
 
-The cosmological principle demands the spacial section of spacetime to be isotropic and homogeneus.
+The cosmological principle demands the spatial section of spacetime to be isotropic and homogeneus.
 This means that the fields involved in the model must be compatible with this assumption.
-A spacetime is an isotropic spacetime with respect to certain point $P$ if after a rotation with respect to an axis passing through $P$, all the geometrical properties remain invariant, thus the spacetime looks the same in all directions.
+A spacetime is isotropic with respect to certain point $P$ if after a rotation with respect to an axis passing through $P$, all the geometrical properties remain invariant, thus the spacetime looks the same in all directions.
 Homogeneity is understood such that spacetime looks the same from every point $P$.
@@
 
-The set of Killing vectors $\{\zeta^\lambda_i\}$ are the generators of $SO(3)$, which generate the spatial rotations in three dimensions, $\xi_i=\epsilon_{ijk}x_j\partial_k$, and the Killing vectors assosiated with spacial translations $\mathcal{P}_i=\sqrt{1-kr^2}\partial_i$, satisfy the algebra
+The set of Killing vectors $\{\zeta^\lambda_i\}$ are the generators of $SO(3)$, which generate the spatial rotations in three dimensions, $\xi_i=\epsilon_{ijk}x_j\partial_k$, and the Killing vectors assosiated with spatial translations $\mathcal{P}_i=\sqrt{1-kr^2}\partial_i$, satisfy the algebra
 \begin{align}
@@
 These requirements on the fields are translated to a set of first order differential equations whose most general solutions determines our ansatz structure from Eq.~\eqref{vielbein cosmo} to \eqref{gamma cosmo}.
-The fields strengths for isotropic--homogeneus vielbein and spin connection are written next.
+The field strengths for isotropic--homogeneus vielbein and spin connection are written next.
 The curvature components are
@@
 
-\section{Time dependence of the solutions\label{solutions t}}
+\section{Time-dependence of the solutions\label{solutions t}}
 

### 2016-09-17 KKosmology.tex
New location of plots
Modification of bibliography, PACS, keywords

@@
 We find cosmological solutions of the Friedmann--Robertson--Walker class for the effective theory.
-  %%In the five-dimensional vacuum, this theory allows a nontrivial torsion in contrast to the Einstein--Cartan theory.
 The torsional degrees of freedom induce a nonvanishing energy-momentum tensor in four dimensions.
-We find solutions describing accelerated expansion phases as well as contracting universes.
+We find solutions describing expanding, contracting and bouncing universes.
 The model shows a dynamical compactification of the extra dimension in some regions of the parameter space.
@@
 
-\pacs{02.40.Ma,04.50.Kd,04.90.+e}
-\keywords{Affine Gravity, Torsion, Generalised Gravity.}
+\pacs{02.40.-k,04.50.Cd,04.50.Kd,98.80.Jk}
+\keywords{Lovelock Gravity, Kaluza--Klein reduction, Extra dimensions, Cosmology.}
 
@@
 
-%The theory of General Relativity (GR) has been proven to be one of the most successful description of gravity nowadays.
-In this framework, the gravitational interaction is understood as the geometrical manifestation of the spacetime.
-The experimental status of General Relativity (GR), regarding the solar system tests~\cite{Will:2014kxa} and the detection of signals consistent with the merge of two black holes by the LIGO collaboration~\cite{Abbott:2016blz,Abbott:2016nmj}, have settled it as the most successful theory of gravity so far.
-However, the difficulties on finding out explanations for the so called \emph{dark sector} of the Universe,
-% and a \emph{consistent} quantum theory of gravitation,
-have driven the community to think that GR is not the ultimate gravitational theory. 
+The experimental status of General Relativity (GR), regarding the solar system tests~\cite{Will:2014kxa} and the detection of signals consistent with the merge of two black holes by the LIGO collaboration~\cite{Abbott:2016blz,*Abbott:2016nmj}, have settled it as the most successful theory of gravity so far.
+However, the difficulties on finding out explanations for the so-called \emph{dark sector} of the Universe, have driven the community to think that GR is not the ultimate gravitational theory.
 The \emph{dark sector} of the Universe is composed by two kinds of degrees of freedom, whose effects are detected at different scales.
@@
 This idea opened the possibility of a novel geometrical understanding of interactions, where the gauge group arise as a consequence of the topology of the spatial compact manifold in a higher dimensional spacetime.
-The idea of higher dimensions comes naturally in diverse physical models, as for example: supersymmetry and supergravity, string theory~\cite{Green:1987sp,*Green:1987mn}, and novel proposals by Arkani-Hammed, Dimopoulos and Dvali~\cite{ArkaniHamed:1998rs,*Antoniadis:1998ig}, and Randall and Sundrum~\cite{Randall:1999ee,*Randall:1999vf} models, as attempts of solving the hierarchy problem.
+The idea of higher dimensions comes naturally in diverse physical models, as for example: supersymmetry and supergravity~\cite{supergravity}, string theory~\cite{Green:1987sp,*Green:1987mn}, and novel proposals by Arkani-Hammed, Dimopoulos and Dvali~\cite{ArkaniHamed:1998rs,*Antoniadis:1998ig}, and Randall and Sundrum~\cite{Randall:1999ee,*Randall:1999vf} models, as attempts of solving the hierarchy problem.
 
@@
 In higher dimensions, particular combinations of higher order terms in the curvature can be added to the gravitational action, giving rise to second order field equations for the metric.
-The most general theory in arbitrary dimensions, which preserves the features of the four-dimensional Einstein--Hilbert, is described by the Lanczos--Lovelock action~\cite{Lanczos:1938sf,Lovelock:1971yv}.
+The most general theory in arbitrary dimensions, which preserves the features of the four-dimensional Einstein--Hilbert, is described by the Lanczos--Lovelock action~\cite{Lanczos:1938sf,*Lovelock:1971yv}.
 Such a theory, has no ghosts~\cite{Zumino:1985dp} and it has the same degrees of freedom as the Einstein--Hilbert Lagrangian in arbitrary dimensions~\cite{Henneaux:1990au}.
@@
 In dimensions higher than four it contributes to the field equations.
-%% Equivalently, the previous $N$-dimensional Lagrangian can be written in exterior forms as
-%% \begin{equation}\label{GB2}
-%% \mathcal{L}_{GB} = \epsilon_{a_1...a_N} \tilde{R}^{a_1a_2}\wedge\tilde{R}^{a_3a_4}\wedge e^{a_5}\wedge...\wedge e^{a_N}.
-%% \end{equation}
 This Lagrangian was also identified as the low-energy correction for a spin two field in string theory~\cite{Zwiebach:1985uq}.
-%%Thus, the consideration of further Lanczos--Lovelock terms in a higher dimensional manifold can be interpreted as the semiclassical corrections to the gravity action.
 
+\hl{Los siguientes dos parrafos no cuadran con la tonica anterior, modificar?}
 On the one hand, classically, wormholes must be supported by a kind energy compatible with the cosmological hypothesis.
 Their existence in extended models, implies the presence of degrees of freedom that could provide an explanation for the exotic matter/energy abundance in the Universe.
-In the Riemannian five-dimensional Lanczos--Lovelock theory, exact wormholes solutions have been found in vacuum~\cite{Dotti:2006cp,Dotti:2007az}.
+In the Riemannian five-dimensional Lanczos--Lovelock theory, exact wormholes solutions have been found in vacuum~\cite{Dotti:2006cp,*Dotti:2007az}.
 The same kind of solutions were found with matter~\cite{Mehdizadeh:2015jra}, in higher order Lanczos--Lovelock models~\cite{Mehdizadeh:2015dta}, and in higher order compactified Lanczos--Lovelock theories with torsion~\cite{Canfora:2008ka}.
-Additionally, the compactification of higher Lanczos--Lovelock terms has been considered in Refs.~\cite{MuellerHoissen:1985mm,MuellerHoissen:1989yv} and their cosmology in Refs.~\cite{MuellerHoissen:1985ij,Deruelle:1986iv,Deruelle:1989fj}.
+Additionally, the compactification of higher Lanczos--Lovelock terms has been considered in Refs.~\cite{MuellerHoissen:1985mm,*MuellerHoissen:1989yv} and their cosmology in Refs.~\cite{MuellerHoissen:1985ij,Deruelle:1986iv,Deruelle:1989fj}.
 
-On the other hand, the renaissance of the Riemann--Cartan geometries came through the pioneer works of Sciama~\cite{Sciama:1962} and Kibble~\cite{Kibble:1961ba} on Poincar\'e gauge theories (see also Refs.~\cite{Hehl:1976kj,Blagojevic:2002du}).
+On the other hand, the renaissance of the Riemann--Cartan geometries came through the pioneer works of Sciama~\cite{Sciama:1962} and Kibble~\cite{Kibble:1961ba} on Poincar\'e gauge theories (see also Refs.~\cite{Blagojevic:2002du,*Hehl:1976kj}).
 In such a framework, the torsion appears as the field strength of translations, sourced by the spin-current.
@@
 The theory admits a nonvanishing torsion in vacuum due to the presence of the Gauss--Bonnet term, in contrast to the five-dimensional Einstein--Cartan theory.
-After the dimensional reduction, the extra degrees of freedom generate an energy density and pressure as four-dimensional matter fields.
+\hl{This are our results, in the intro?}After the dimensional reduction, the extra degrees of freedom generate an energy density and pressure as four-dimensional matter fields.
 In some cases, the solutions avoid the appearance of initial singularities and drive the accelerated expansion of the Universe, while the radius of the compact manifold goes to zero.
@@
 
-The Kaluza--Klein (KK) ansatz for the metric relies on the premise that the compact dimension of $M_N$ is orthogonal to the rest of the manifold at each point (see for example Ref.~\cite{Blagojevic:2002du}).
+The Kaluza--Klein (KK) ansatz for the metric relies on the premise that the compact dimension of $M_N$ is orthogonal to the rest of the manifold at each point (see for example Ref.~\cite{Blagojevic:2002du,*Hehl:1976kj}).
 This leads to the following metric structure
@@
 This condition place us outside the Chern--Simons point.
-Otherwise, the field equations~\eqref{delta e} and~\eqref{delta w} are invariant under a larger gauge group ($AdS_5$), while the action~\eqref{action5EGB} becomes the Chern--Simons form for that group~\cite{Zanelli:2005sa,Troncoso:1999pk}.
+Otherwise, the field equations~\eqref{delta e} and~\eqref{delta w} are invariant under a larger gauge group ($AdS_5$), while the action~\eqref{action5EGB} becomes the Chern--Simons form for that group~\cite{Zanelli:2005sa,*Troncoso:1999pk}.
 
@@
 The Eqs.~\eqref{eqn1} to~\eqref{eqn2} are reduced to the Riemannian system when \mbox{$f=h=0$} and \mbox{$\gamma=-\dot{\phi}/2\phi$}.
-The details of such a model can be seen in Refs.~\cite{Deruelle:1986iv,Deruelle:2003ck,Henriques:1986jw,Ishihara:1986if,Kleidis:1997mu}.
+The details of such a model can be seen in Refs.~\cite{Deruelle:1986iv,Deruelle:2003ck,Henriques:1986jw,*Ishihara:1986if,Kleidis:1997mu}.
 
@@
 It is worth mentioning the existence of models of gravity which are free of the linear curvature term.
-These models are refer to as Pure Lovelock (PL) gravities~\cite{Cai:2006pq} and consider the cosmological term and the polynomial of highest order in the curvature.
-Even though the analysis that follows will consider a positive non-zero value for $\alpha_1$, the solutions listed above are also valid in the regime where $\alpha_1=0$, that constitutes the non-Riemannian extension of PL theories.
-
-\subsubsection{Bouncing solutions}
-
-In Fig.~\ref{cond_u_pos} we show the allowed region, in the $\alpha_0$--$\alpha_2$ parameter space, for $u_\pm$ to be positive. 
+These models are refer to as pure Lovelock gravities~\cite{Cai:2006pq} and consider the cosmological term and the polynomial of highest order in the curvature.
+Even though the analysis that follows will consider a positive non-zero value for $\alpha_1$, the solutions listed above are also valid in the regime where $\alpha_1=0$, that constitutes the non-Riemannian extension of pure Lovelock theories.
 \begin{figure}[H]
@@
 \end{figure}
+
+\subsubsection{Bouncing solutions}
+
+In Fig.~\ref{cond_u_pos} we show the allowed region, in the $\alpha_0$--$\alpha_2$ parameter space, for $u_\pm$ to be positive.
 
@@
 \begin{figure}[H]
-  \includegraphics[width=\linewidth]{a_u_pos.pdf}
+  \includegraphics[width=.95\linewidth]{a_u_pos.pdf}
   \caption{Behavior of the scale factor, \(a(t)\), as a function of the scaled time, $\tau$, for $u_\pm > 0$.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\linewidth]{phi_u_pos.pdf}
+  \includegraphics[width=.95\linewidth]{phi_u_pos.pdf}
   \caption{Behavior of the $\phi$ field as a function of the scaled time, $\tau$, for $u_\pm > 0$.}
@@
 For all these solutions the scale factor remains finite at $t = 0$, and they do not present an initial singularity.
-It has been reported that the Gauss--Bonnet term can prevent the Universe to expand from an initial singularity~\cite{Deruelle:1986iv,Henriques:1986jw,Ishihara:1986if}, which is also this case.
+It has been reported that the Gauss--Bonnet term can prevent the Universe to expand from an initial singularity~\cite{Deruelle:1986iv,Henriques:1986jw,*Ishihara:1986if}, which is also this case.
 

### 2016-09-17 KKosmology.tex
Explanation of a_max and phi_max

@@
 \begin{figure}[H]
-  \includegraphics[width=\linewidth]{cond_u_neg.pdf}
+  \includegraphics[width=.95\linewidth]{cond_u_neg.pdf}
   \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be positive defined.}
@@
   \includegraphics[width=.95\linewidth]{a_u_neg.pdf}
-  \caption{Different behaviors for the scale factor, as function of the scaled time $\tau$, for $u_\pm < 0$.}
+  \caption{Different behaviors for the scale factor, as function of the scaled time $\tau$, for $u_\pm < 0$.
+We have normalized using the maximum value of the scale factor, $a_{\mathrm{max}}$, in the plotted region.}
   \label{a_u_neg}
@@
   \includegraphics[width=.95\linewidth]{phi_u_neg.pdf}
-  \caption{Different behaviors for the $\phi$ field, as function of the scaled time $\tau$, for $u_\pm < 0$.}
+  \caption{Different behaviors for the $\phi$ field, as function of the scaled time $\tau$, for $u_\pm < 0$.
+We have normalized using the maximum value of the $\phi$ field, $\phi_{\mathrm{max}}$, in the plotted region.}
   \label{phi_u_neg}

### 2016-09-17 KKosmology.tex
Format changes... not that important.

@@
 Due to the topology of the manifold, we can expand the depenence of the fields on the extra coordinate, $z$, in Fourier series as  
-\begin{align}\label{Fourier}
-  \hat{\varrho}(\hat{x}^\Gamma)&=\sum_n\varrho_{(n)}(x)e^{i n z},
-\end{align}
+\begin{equation}
+  \label{Fourier}
+  \hat{\varrho}(\hat{x}^\Gamma)=\sum_n\varrho_{(n)}(x)e^{i n z},
+\end{equation}
 where $x$ denotes the coordinates of the $(N-1)$-manifold, collectively. Henceforth, we will focus in the $n=0$ mode of the expasion, also refered to as the low-energy sector.
@@
   \label{curvadef}
-  \hat{R}^{AB} &= \mbox{d}\hat{\omega}^{AB}+\hat{\omega}^A_{\ \ C}\wedge\hat{\omega}^{CB} = \frac{1}{2} \hat{R}^{AB}{}_{CD} \, \hat{e}^C \wedge \hat{e}^D,\\
+  \hat{R}^{AB} &= \mbox{d}\hat{\omega}^{AB}+\hat{\omega}^A_{\ \ C}\wedge\hat{\omega}^{CB} = \frac{1}{2} \hat{R}^{AB}{}_{CD} \; \hat{e}^C \wedge \hat{e}^D,\\
   \label{tordef}
-  \hat{T}^A &= \mbox{d}\hat{e}^A+\hat{\omega}^A_{\ \ B}\wedge\hat{e}^B = \frac{1}{2} \hat{T}^{A}{}_{BC} \, \hat{e}^B \wedge \hat{e}^C. 
+  \hat{T}^A &= \mbox{d}\hat{e}^A+\hat{\omega}^A_{\ \ B}\wedge\hat{e}^B = \frac{1}{2} \hat{T}^{A}{}_{BC} \; \hat{e}^B \wedge \hat{e}^C. 
 \end{align}
@@
   \hat{R}^{ab}&=R^{ab}+\sqrt{\phi}\alpha^{ab}F-\beta^a\wedge\beta^b\notag\\
-  &+\left(\mbox{D}\alpha^{ab}+\frac{1}{2}\alpha^{ab}\mbox{d}\ln\phi-2\beta^{[a}\gamma^{b]}\right)\wedge\hat{e}^N,\\
+  & \quad +\left(\mbox{D}\alpha^{ab}+\frac{1}{2}\alpha^{ab}\mbox{d}\ln\phi-2\beta^{[a}\gamma^{b]}\right)\wedge\hat{e}^N,\\
   \label{R2}
   \hat{R}^{Na}&=-\left(\mbox{D}\beta^a+\sqrt{\phi}\gamma^a F\right)\notag\\
-  &+\left(\alpha^a_{\ b}\beta^b-\mbox{D}\gamma^a-\frac{1}{2}\gamma^a\mbox{d}\ln\phi\right)\wedge\hat{e}^N.
+  & \quad +\left(\alpha^a_{\ b}\beta^b-\mbox{D}\gamma^a-\frac{1}{2}\gamma^a\mbox{d}\ln\phi\right)\wedge\hat{e}^N.
 \end{align}
@@
 Taking the exterior covariant derivative over the $N$-dimensional curvature and torsion, the Bianchi identities are
-\begin{align*}
-  \hat{\text{D}}\hat{R}^{AB} &= 0 &\mbox{and}& &\hat{\text{D}}\hat{T}^A &= \hat{R}^A_{\ \ B}\wedge\hat{e}^B.
-\end{align*}
+\begin{equation}
+  \hat{\text{D}}\hat{R}^{AB} = 0 \quad \text{and} \quad \hat{\text{D}}\hat{T}^A = \hat{R}^A_{\ \ B}\wedge\hat{e}^B.
+\end{equation}
 A careful decomposition of the first one into its distinctive parts, gives the Bianchi identity for the curvature of the reduced spacetime, together with the second derivative rules
-\begin{align*}
+\begin{equation}
+  \begin{aligned}
     \text{D}R^{ab} &=0, & \text{D}\mbox{D}\alpha^{ab} &=R^a_{\ \ l}\alpha^{lb}+R^b_{\ \ l}\alpha^{al},\\
     \text{D}\text{D}\beta^a &= R^a_{\ \ b}\beta^{b}, & \text{D}\text{D}\gamma^a &= R^a_{\ \ b}\gamma^{b}.
-\end{align*}
+  \end{aligned}
+\end{equation}
 The second Bianchi identity, gives nothing but its equivalent for the manifold $M_{N-1}$, this is
@@
   where $\alpha_0$, $\alpha_1$ and $\alpha_2$ are dimensionful coupling constants. Its variation with respect to the five-dimensional vielbein and spin connection give the equations
-  \begin{align}
+  \begin{gather}
     \label{delta e}
@@
     + \alpha_1\hat{R}^{AB}\wedge\hat{e}^C\wedge\hat{e}^D
-    + \alpha_2\hat{R}^{AB}\wedge\hat{R}^{CD}\Big)&=0,
+    + \alpha_2\hat{R}^{AB}\wedge\hat{R}^{CD}\Big)=0,
     \\
@@
     \hat{\epsilon}_{ABCDE}\Big(\alpha_1\hat{e}^C\wedge\hat{e}^D+
-    2\alpha_2\hat{R}^{CD}\Big)\wedge\hat{T}^E&=0.
-  \end{align}
+    2\alpha_2\hat{R}^{CD}\Big)\wedge\hat{T}^E=0.
+  \end{gather}
 \end{widetext}
@@
 In this work, we will focus on the region in the parameter space where
-\begin{equation}\label{delta}
+\begin{equation}
+  \label{delta}
   \Delta\equiv\alpha_1^2-4\alpha_0\alpha_2\neq 0.
@@
 its \(M_{4} \times S^1\) decomposition. The Eq.~\eqref{delta e} leads to 
-\begin{align}\label{equation}
-  \epsilon_{abcd}\Big[\alpha_0 e^b\wedge e^c\wedge e^d + \frac{1}{2}\alpha_1\left(M^{bc}-L^b\wedge e^c\right)\wedge e^d&\notag\\
-    -\alpha_2\left(L^b\wedge M^{cd} + K^b\wedge N^{cd}\right)\Big]&=0,\\
-  \epsilon_{abcd}K^b\wedge\left(\alpha_1e^c\wedge e^d + 2\alpha_2M^{cd}\right)&=0,\\
-  \epsilon_{abcd}\big(\alpha_0 e^a\wedge e^b\wedge e^c\wedge e^d&\notag\\
-  +\alpha_1M^{ab}\wedge e^c\wedge e^d+\alpha_2M^{ab}\wedge M^{cd}\big)&=0,\\
-  \epsilon_{abcd}\left(\alpha_1N^{ab}\wedge e^c\wedge e^d+2\alpha_2N^{ab}M^{cd}\right)&=0,
-\end{align}
+\setlength\multlinegap{0pt}
+\begin{gather}
+  \begin{multlined}
+    \label{equation}
+    \epsilon_{abcd}\Big[\alpha_0 e^b\wedge e^c\wedge e^d + \frac{1}{2}\alpha_1\left(M^{bc}-L^b\wedge e^c\right)\wedge e^d\\
+      -\alpha_2\left(L^b\wedge M^{cd} + K^b\wedge N^{cd}\right)\Big]=0,
+  \end{multlined}
+  \\
+  \epsilon_{abcd}K^b\wedge\left(\alpha_1e^c\wedge e^d + 2\alpha_2M^{cd}\right)=0,
+  \\
+  \begin{multlined}
+    \epsilon_{abcd}\big(\alpha_0 e^a\wedge e^b\wedge e^c\wedge e^d \\
+    +\alpha_1M^{ab}\wedge e^c\wedge e^d+\alpha_2M^{ab}\wedge M^{cd}\big)=0,
+  \end{multlined}
+  \\
+  \epsilon_{abcd}\left(\alpha_1N^{ab}\wedge e^c\wedge e^d+2\alpha_2N^{ab}M^{cd}\right)=0,
+\end{gather}
 while Eq.~\eqref{delta w} gives
-\begin{align}
-  \epsilon_{abcd}\Big[\alpha_1\left(e^c\wedge e^d\wedge Z-2e^c\wedge T^d\right)
-    +2\alpha_2\Big(N^{cd}\wedge W&\notag\\
+\setlength\multlinegap{0pt}
+\begin{gather}
+  \begin{multlined}
+    \epsilon_{abcd}\Bigl[\alpha_1\left(e^c\wedge e^d\wedge Z-2e^c\wedge T^d\right)
+      +2\alpha_2\big(N^{cd}\wedge W\\
       +M^{cd}\wedge Z+2L^c\wedge T^d
-    +2K^c\wedge V^d\Big)\Big]&=0,\\
-  \epsilon_{abcd}\left[\alpha_1e^c\wedge e^ d\wedge W+2\alpha_2\left(M^{cd}\wedge W+2K^c\wedge T^d\right)\right]&=0,\\
-  \epsilon_{abcd}\left(\alpha_1e^b\wedge e^c+2\alpha_2M^{bc}\right)\wedge T^d&=0,\\
-  \epsilon_{abcd}\left[\alpha_1e^b\wedge e^c\wedge V^d+2\alpha_2\Big(M^{bc}\wedge V^d+N^{bc}\wedge T^d\Big)\right]&=0.
-\end{align}
+      +2K^c\wedge V^d\big)\Bigr]=0,
+  \end{multlined}
+  \\
+  \begin{multlined}
+    \epsilon_{abcd}\Bigl[\alpha_1e^c\wedge e^ d\wedge W \\
+      + 2\alpha_2\left(M^{cd}\wedge W+2K^c\wedge T^d\right)\Bigr]=0,
+  \end{multlined}
+  \\
+  \epsilon_{abcd}\left(\alpha_1e^b\wedge e^c+2\alpha_2M^{bc}\right)\wedge T^d=0,
+  \\
+  \begin{multlined}
+    \epsilon_{abcd}\Bigl[\alpha_1e^b\wedge e^c\wedge V^d \\
+      + 2\alpha_2\big(M^{bc}\wedge V^d+N^{bc}\wedge T^d\big)\Bigr]=0.
+  \end{multlined}
+\end{gather}
 The fields $M^{ab}$, $N^{ab}$, $L^a$, $K^a$, $W$, $V^a$ and $Z$ are defined from the Eq.~\eqref{R1} to~\eqref{T2}, such that
-\begin{align*}
+\begin{equation}
+  \begin{aligned}
     \hat{R}^{ab}&=M^{ab}+N^{ab}\wedge\hat{e}^5, & \hat{R}^{5a}&=K^a+L^a\wedge\hat{e}^5,\\
     \hat{T}^a&=T^a+V^a\wedge\hat{e}^5, & \hat{T}^5&=W+Z\wedge\hat{e}^5.
-\end{align*}
+  \end{aligned}
+\end{equation}
 
@@
 The four-dimensional vielbein compatible with the Friedmann--Robertson--Walker metric is determined only by one function, the scale factor $a(t)$, and their components read
-\begin{align}
+\begin{equation}
+  \begin{aligned}
     \label{vielbein cosmo}
-  e^0&=\mbox{d}t, & e^1&=\frac{a(t)}{\sqrt{1-kr^2}}\mbox{d}r,\\
-  e^2&=a(t)r\mbox{d}\theta, & e^3&=a(t)r\sin\varphi\mbox{d}\varphi, 
-\end{align}
+    e^0&=\mbox{d}t, & e^1&=\frac{a(t)}{\sqrt{1-kr^2}} \, \mbox{d}r,\\
+    e^2&=a(t)r \, \mbox{d}\theta, & e^3&=a(t) r \sin(\varphi) \,\mbox{d}\varphi, 
+  \end{aligned}
+\end{equation}
 where $k=+1,0,-1$ determines the extrinsic curvature of the four-dimensional manifold to be closed, flat or open, respectively.
@@
 \begin{align}
-  \omega^{0i}&=\omega(t) e^i,\\
-  \omega^{12}&=-\frac{\sqrt{1-kr^2}}{a(t)r}e^2-f(t)e^3,\\
-  \omega^{13}&=-\frac{\sqrt{1-kr^2}}{a(t)r}e^3+f(t)e^2,\\
-  \omega^{23}&=-\frac{\cot\theta}{a(t)r}e^3-f(t)e^1.
+  \omega^{0i}&=\omega(t) \, e^i,\\
+  \omega^{12}&=-\frac{\sqrt{1-kr^2}}{a(t)r} \, e^2 - f(t) \, e^3,\\
+  \omega^{13}&=-\frac{\sqrt{1-kr^2}}{a(t)r} \, e^3 + f(t) \, e^2,\\
+  \omega^{23}&=-\frac{\cot\theta}{a(t)r} \, e^3 - f(t) \, e^1.
 \end{align}
@@
 The nonvanishing components of the spin connection induced by the compact manifold are
-\begin{align}
-  \beta^0=-b(t)e^0,&\ \beta^i=\beta(t)e^i,\\
+\begin{gather}
+  \beta^0 = -b(t) \, e^0,\quad \beta^i = \beta(t) \, e^i,\\
   \label{gamma cosmo}
-  \gamma^0&=-\gamma(t).
-\end{align}
+  \gamma^0 =-\gamma(t).
+\end{gather}
 
@@
 The final set of equations is the following:
-\begin{align}
+\setlength\multlinegap{0pt}
+\begin{gather}
   \label{eqn1}
-  h\left[\alpha_1+2\alpha_2\left(\omega^2+\frac{k}{a^2}-f^2\right)\right]+4\alpha_2f\left(\dot{f}+Hf\right)&=0,\\
-  2\alpha_0+\alpha_1\left(\dot{\omega}+2\omega^2+\frac{k}{a^2}-f^2\right)\notag\\
-  +2\alpha_2\left(\dot{\omega}+\omega^2\right)\left(\omega^2+\frac{k}{a^2}-f^2\right)&=0,\\
-  2\alpha_2\left(\frac{\dot{\phi}}{2\phi}+\gamma\right)f\left(\dot{f}+Hf\right)
-  +h^2\left(\alpha_1-2\alpha_2\omega\gamma\right)&=0,\\
-  \omega\left(\frac{\dot{\phi}}{2\phi}+\gamma\right)+\dot{\gamma}+\frac{\dot{\phi}}{2\phi}\gamma-\frac{\alpha_1}{2\alpha_2}&=0,\\
-  \left(\omega^2+\frac{k}{a^2}-f^2\right)\left(\alpha_1-2\alpha_2\omega\gamma\right)-\alpha_1\omega\gamma+2\alpha_0&=0,\\
+  \begin{multlined}
+    h\left[\alpha_1+2\alpha_2\left(\omega^2+\frac{k}{a^2}-f^2\right)\right] \\
+    +4\alpha_2f\left(\dot{f}+Hf\right)=0,
+  \end{multlined}
+  \\
+  \begin{multlined}
+    2\alpha_0+\alpha_1\left(\dot{\omega}+2\omega^2+\frac{k}{a^2}-f^2\right)\\
+    +2\alpha_2\left(\dot{\omega}+\omega^2\right)\left(\omega^2+\frac{k}{a^2}-f^2\right)=0,
+  \end{multlined}
+  \\
+  2\alpha_2\left(\tfrac{\dot{\phi}}{2\phi}+\gamma\right)f\left(\dot{f}+Hf\right)
+  +h^2\left(\alpha_1-2\alpha_2\omega\gamma\right)=0,
+  \\
+  \omega\left(\frac{\dot{\phi}}{2\phi}+\gamma\right)+\dot{\gamma}+\frac{\dot{\phi}}{2\phi}\gamma-\frac{\alpha_1}{2\alpha_2}=0,
+  \\
+  \left(\omega^2+\tfrac{k}{a^2}-f^2\right)\left(\alpha_1-2\alpha_2\omega\gamma\right)-\alpha_1\omega\gamma+2\alpha_0=0,
+  \\
   \label{eqn2}
   \left(\dot{\omega}+\omega^2\right)\left(\alpha_1-2\alpha_2\omega\gamma\right)-\alpha_1
-  \omega\gamma+3\alpha_0-\frac{\alpha^2_1}{4\alpha_2}&=0,
-\end{align}
+  \omega\gamma+3\alpha_0 - \frac{\alpha^2_1}{4\alpha_2}=0,
+\end{gather}
 where dot stands for time-derivative. For the sake of simplicity, we have defined $h(t)=\omega(t)-H(t)$, where $H=\dot{a}/a$ is the Hubble function.
@@
 The system is integrated and develops two non-Riemannian branches, one for each value of the parameter $u_\pm$ defined to be
-\begin{equation}\label{u}
+\begin{equation}
+  \label{u}
   u_\pm=\frac{2\alpha_1\pm\sqrt{6\Delta}}{4\alpha_2},
@@
 
-Due to Eq.~\eqref{u}, the solutions will be valid in the region of the parameter space where $\Delta>0$. The function $\omega(t)$ satisfies the equation
-\begin{align}
-  \dot{\omega}+\omega^2+u_{\pm} &= 0,
-\end{align}
-which, for the three significantly different values of $u_\pm$, has the following solutions
-\begin{align}
-  u_\pm>0&:\omega(t)=-\sqrt{u_\pm}\tan\left[\sqrt{u_\pm}\left(t-t_0\right)\right],\\
-  u_\pm=0&:\omega(t)=\left(t-t_0\right)^{-1},\\
-  u_\pm<0&:\omega(t)=\sqrt{-u_\pm}\tanh\left[\sqrt{-u_\pm}\left(t-t_0\right)\right],
-\end{align}
+Due to Eq.~\eqref{u}, the solutions will be valid in the region of the parameter space where $\Delta>0$.
+The function $\omega(t)$ satisfies the equation\footnote{Notice that in our choice of parameters $u_+$ cannot vanish.}
+\begin{equation}
+  \dot{\omega}+\omega^2+u_{\pm} = 0,
+\end{equation}
+which, for the three significantly different values of $u_\pm$, has the following solutions\footnote{Notice that in our choice of parameters $u_+$ cannot vanish.}
+\begin{equation}
+  \omega(t) =
+  \begin{cases}
+    -\sqrt{u_\pm}\tan\left[\sqrt{u_\pm}\left(t-t_0\right)\right] & u_\pm > 0 \\
+    \left(t-t_0\right)^{-1} & u_- = 0 \\
+    \sqrt{-u_\pm}\tanh\left[\sqrt{-u_\pm}\left(t-t_0\right)\right] & u_\pm < 0
+  \end{cases}
+\end{equation}
 where $t_0$ is an integration constant to be fixed.
@@
   f^2(\omega)&=\left(\omega^2+u_\pm\right)\frac{k}{a_0^2}\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2\left(\omega^2+u_\pm\right)}\right]\notag\\
-  &+\omega^2  + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2},
+  & \quad +\omega^2  + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2},
 \end{align}
@@
   The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, A.~Toloza and J.~Zanelli for enlightening discussions on this work.
-O.~C-F.
-has been partially supported by the PAI project No.~79140040 (CONICYT--Chile).
-The Centro Cient\'ifico y Tecnol\'ogico de Valpara\'iso (CCTVal) is funded by the Chilean Goverment through the Centers of Excellence Base Financing Program of CONICYT.
+\mbox{O.~C.-F.} has been partially supported by the PAI project No.~79140040 (CONICYT--Chile).
+The Centro Cient\'ifico y Tecnol\'ogico de Valpara\'iso (CCTVal) is funded by the Chilean Government through the Centers of Excellence Basal Financing Program of CONICYT.
 \end{acknowledgments}

### 2016-09-18 KKosmology.tex
More format changes

@@
   \label{killing metric}
-  \text{\textsterling}_i g_{\mu\nu}&=\zeta^\lambda_i\partial_\lambda g_{\mu\nu}+\partial_\mu\zeta^\lambda_i g_{\lambda\nu}+\partial_\nu\zeta^\lambda_i g_{\mu\lambda}=0,\\
+  \text{\textsterling}_i g_{\mu\nu}&=\zeta^\lambda_i\partial_\lambda g_{\mu\nu}+\partial_\mu\zeta^\lambda_i g_{\lambda\nu}+\partial_\nu\zeta^\lambda_i g_{\mu\lambda}=0,
+  \\
   \text{\textsterling}_i T^\alpha_{\ \mu\nu}&=\zeta^\lambda_i\partial_\lambda T^\alpha_{\ \mu\nu}-\zeta^\alpha_i\partial_\lambda T^\lambda_{\ \mu\nu}
-  +\zeta^\lambda_i\partial_\mu T^\alpha_{\ \lambda\nu}\notag\\
+  +\zeta^\lambda_i\partial_\mu T^\alpha_{\ \lambda\nu} \notag \\
   &\quad +\zeta^\lambda_i\partial_\nu T^\alpha_{\ \mu\lambda}=0,
@@
 \end{align*} 
-and whose Killing equations are analogous to the tensorial one,~\eqref{killing metric} and vectorial one,~\eqref{killing vector}.
+and whose Killing equations are analogous to the tensorial, Eq.~\eqref{killing metric}, and vectorial one, Eq.~\eqref{killing vector}.
 
-The set of Killing vectors $\{\zeta^\lambda_i\}$ are the generators of $SO(3)$, which generate the spatial rotations in three dimensions, $\xi_i=\epsilon_{ijk}x_j\partial_k$, and the Killing vectors assosiated with spatial translations $\mathcal{P}_i=\sqrt{1-kr^2}\partial_i$, satisfy the algebra
-\begin{align}
-  \left[\xi_i,\xi_j\right]&=\epsilon_{ijk}\xi_k,\\
-  \left[\mathcal{P}_i,\mathcal{P}_j\right]&=-k\epsilon_{ijk}\xi_k,\\
-  \left[\xi_i,\mathcal{P}_j\right]&=\epsilon_{ijk}\xi_k.
-\end{align}
-These requirements on the fields are translated to a set of first order differential equations whose most general solutions determines our ansatz structure from Eq.~\eqref{vielbein cosmo} to \eqref{gamma cosmo}.
+The set of Killing vectors $\{\zeta^\lambda_i\}$ are the generators of $SO(3)$, which generate the spatial rotations in three dimensions, $\mathcal{J}_i=\epsilon_{ijk}x_j\partial_k$, and the Killing vectors assosiated with spatial translations $\mathcal{P}_i=\sqrt{1-kr^2}\partial_i$, satisfy the algebra
+\begin{align*}
+  \left[\mathcal{J}_i,\mathcal{J}_j\right]&=\epsilon_{ijk}\mathcal{J}_k,\\
+  \left[\mathcal{P}_i,\mathcal{P}_j\right]&=-k\epsilon_{ijk}\mathcal{J}_k,\\
+  \left[\mathcal{J}_i,\mathcal{P}_j\right]&=\epsilon_{ijk}\mathcal{P}_k.
+\end{align*}
+These requirements on the fields are translated to a set of first order differential equations whose most general solutions determines our ansatz structure from Eqs.~\eqref{vielbein cosmo} to~\eqref{gamma cosmo}.
 The field strengths for isotropic--homogeneus vielbein and spin connection are written next.
 The curvature components are
-\begin{align}
-  R^{0i}&=\left(\dot{\omega}+H\omega\right)e^0\wedge e^i+f\omega\epsilon^i_{\ jk}e^j\wedge e^k,\\
-  R^{ij}&=\left(\omega^2+\frac{k}{a^2}-f^2\right)e^i\wedge e^j
-  -\left(\dot{f}+Hf\right)\epsilon^{ij}_{\ \ k}e^0\wedge e^k,
-\end{align}
+\begin{align*}
+  R^{0i}&=\left(\dot{\omega}+H\omega\right) \, e^0\wedge e^i+f\omega\epsilon^i{}_{jk} \, e^j\wedge e^k,\\
+  R^{ij}&=\left(\omega^2+\tfrac{k}{a^2}-f^2\right) \, e^i\wedge e^j
+  -\left(\dot{f}+Hf\right)\epsilon^{ij}{}_{k} \, e^0\wedge e^k,
+\end{align*}
 and the torsion components
 \begin{equation}\label{homotorsion}
-  T^0=0,\quad T^i=-he^0\wedge e^i+f\epsilon^i_{\ jk}e^j\wedge e^k,
+  T^0=0,\quad T^i=-h \, e^0\wedge e^i+f\epsilon^i{}_{jk} \, e^j\wedge e^k,
 \end{equation}
-where $h(t)$ and $H(t)$ were defined in Section \ref{5EGB}.
+where $h(t)$ and $H(t)$ were defined in Sec.~\ref{5EGB}.
 
@@
 
-We supply here the time-dependent expressions for the cosmological solutions, given the differents values of $u_\pm$, in terms of the dimensionless parameter~$\tau=\sqrt{|u_\pm|}(t-t_0)$.
-\medskip
+We supply here the time-dependent expressions for the cosmological solutions, given the differents values of $u_\pm$, in terms of the dimensionless parameter \mbox{$\tau=\sqrt{|u_\pm|}(t-t_0)$}.
 
-$\bullet\ u_\pm>0$:
+\begin{description}
+\item[$\bm{u_\pm>0}$]
   \begin{align}
     \gamma(t)&=-\sqrt{u_\pm}\cot\tau,\\
-  \phi(t)&=\phi_0\sin^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2 u_\pm}\cos^2\tau\right],\\
-  a(t)&=\frac{a_0}{\sqrt{u_\pm}}|\cos\tau|\exp\left[\frac{\pm\sqrt{6\Delta}}{24\alpha_2 u_\pm}\cos^2\tau\right],\\
-  f^2(t)&=u_\pm\frac{k}{a_0^2}\sec^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cos^2\tau\right]\notag\\
-  &+u_\pm\tan^2\tau + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
+    \phi(t)&=\phi_0\sin^2\tau\exp\left[\tfrac{\mp\sqrt{6\Delta}}{4\alpha_2 u_\pm}\cos^2\tau\right],\\
+    a(t)&=\tfrac{a_0}{\sqrt{u_\pm}}|\cos\tau|\exp\left[\tfrac{\pm\sqrt{6\Delta}}{24\alpha_2 u_\pm}\cos^2\tau\right],\\
+    f^2(t)&=u_\pm\tfrac{k}{a_0^2}\sec^2\tau\exp\left[\tfrac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cos^2\tau\right]\notag\\
+    &+u_\pm\tan^2\tau + \tfrac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
   \end{align}
 
-$\bullet\ u_\pm=0$:
+\item[$\bm{u_\pm=0}$]
   \begin{align}
     \gamma(t)&=0,\\
-  \phi(t)&=\phi_0\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2}(t-t_0)^2\right],\\
-  a(t)&=a_0|t-t_0|\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2}(t-t_0)^2\right],\\
-  f^2(t)&=\frac{k}{a_0^2(t-t_0)^2}\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2}(t-t_0)^2\right]\notag\\
-  &+(t-t_0)^{-2} + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
+    \phi(t)&=\phi_0\exp\left[\tfrac{\mp\sqrt{6\Delta}}{4\alpha_2}(t-t_0)^2\right],\\
+    a(t)&=a_0|t-t_0|\exp\left[\tfrac{\pm\sqrt{6\Delta}} {24\alpha_2}(t-t_0)^2\right],\\
+    f^2(t)&=\tfrac{k}{a_0^2(t-t_0)^2}\exp\left[\tfrac{\mp\sqrt{6\Delta}}{12\alpha_2}(t-t_0)^2\right]\notag\\
+    &+(t-t_0)^{-2} + \tfrac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
   \end{align}
 
-$\bullet\ u_\pm<0$:
+\item[$\bm{u_\pm<0}$]
   \begin{align}
     \gamma(t)&=-\sqrt{-u_\pm}\coth\tau,\\
-  \phi(t)&=\phi_0\sinh^2\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2u_\pm}\cosh^2\tau\right],\\
-  a(t)&=\frac{a_0}{\sqrt{-u_\pm}}\cosh\tau\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2u_\pm}\cosh^2\tau\right],\\
-  f^2(t)&=-u_\pm\frac{k}{a_0^2}\cosh^{-2}\tau\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cosh^2\tau\right]\notag\\
-  &-u_\pm\tanh^2\tau + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
+    \phi(t)&=\phi_0\sinh^2\tau\exp\left[\tfrac{\mp\sqrt{6\Delta}}{4\alpha_2u_\pm}\cosh^2\tau\right],\\
+    a(t)&=\tfrac{a_0}{\sqrt{-u_\pm}}\cosh\tau\exp\left[\tfrac{\pm\sqrt{6\Delta}} {24\alpha_2u_\pm}\cosh^2\tau\right],\\
+    f^2(t)&=-u_\pm\tfrac{k}{a_0^2}\cosh^{-2}\tau\exp\left[\tfrac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cosh^2\tau\right]\notag\\
+    &-u_\pm\tanh^2\tau + \tfrac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
   \end{align}
+\end{description}
 %%%%%%%%% BIBLIOGRAPHY %%%%%%%%%

### 2016-09-21 KKosmology.tex
drop a duplicated footnote
Changes by Criz and SdP

@@
 
-%% \author{Francisca \surname{Ram\'irez}.}
-%% \affiliation{\UTFSM.}
+\author{Francisca \surname{Ram\'irez}.}
+\affiliation{\UTFSM.}
 
-%% --------- Abstract
 \begin{abstract}
@@
 In four dimensions, the Einstein--Hilbert action with cosmological constant is the most general theory which leads to second order field equations for the metric.
-In higher dimensions, particular combinations of higher order terms in the curvature can be added to the gravitational action, giving rise to second order field equations for the metric.
+In higher dimensions, however, particular combinations of higher order terms in the curvature can be added to the gravitational action, giving rise to second order field equations for the metric.
 The most general theory in arbitrary dimensions, which preserves the features of the four-dimensional Einstein--Hilbert, is described by the Lanczos--Lovelock action~\cite{Lanczos:1938sf,*Lovelock:1971yv}.
@@
 The Gauss--Bonnet term, in four dimensions, adds no dynamics to the metric, since it represents a topological invariant proportional to the Euler characteristic class, which can be written locally as a boundary term.
-In dimensions higher than four it contributes to the field equations.
+Nevertheless, in dimensions higher than four it contributes to the field equations.
 This Lagrangian was also identified as the low-energy correction for a spin two field in string theory~\cite{Zwiebach:1985uq}.
 
-\hl{Los siguientes dos parrafos no cuadran con la tonica anterior, modificar?}
-On the one hand, classically, wormholes must be supported by a kind energy compatible with the cosmological hypothesis.
-Their existence in extended models, implies the presence of degrees of freedom that could provide an explanation for the exotic matter/energy abundance in the Universe.
-In the Riemannian five-dimensional Lanczos--Lovelock theory, exact wormholes solutions have been found in vacuum~\cite{Dotti:2006cp,*Dotti:2007az}.
-The same kind of solutions were found with matter~\cite{Mehdizadeh:2015jra}, in higher order Lanczos--Lovelock models~\cite{Mehdizadeh:2015dta}, and in higher order compactified Lanczos--Lovelock theories with torsion~\cite{Canfora:2008ka}.
+The Riemannian five-dimensional Lanczos--Lovelock theory has been widely studied in the literature.
+For instance, exact wormhole solutions have been found in vacuum~\cite{Dotti:2006cp,*Dotti:2007az} and coupled with matter fields~\cite{Mehdizadeh:2015jra}.
+In higher order Lanczos--Lovelock models, this class of exact solutions have been reported in~\cite{Mehdizadeh:2015dta} and in higher order compactified Lanczos--Lovelock theories with torsion in~\cite{Canfora:2008ka}.
 Additionally, the compactification of higher Lanczos--Lovelock terms has been considered in Refs.~\cite{MuellerHoissen:1985mm,*MuellerHoissen:1989yv} and their cosmology in Refs.~\cite{MuellerHoissen:1985ij,Deruelle:1986iv,Deruelle:1989fj}.
+It is worth mentioning that, classically, wormholes must be supported by a kind energy compatible with the cosmological hypothesis.
+Their existence in these extended models implies the presence of degrees of freedom that could provide an explanation for the exotic matter/energy abundance in the Universe.
 
-On the other hand, the renaissance of the Riemann--Cartan geometries came through the pioneer works of Sciama~\cite{Sciama:1962} and Kibble~\cite{Kibble:1961ba} on Poincar\'e gauge theories (see also Refs.~\cite{Blagojevic:2002du,*Hehl:1976kj}).
-In such a framework, the torsion appears as the field strength of translations, sourced by the spin-current.
+On the other hand, it is well-known that GR assume the torsion-free condition {\it a priori}.
+However, as E.~Cartan first considered, it is possible to take metricity and parallelism as truly independent concepts~\cite{Cartan1922,*Cartan1924,*Cartan1925}.
+Such kind of geometries are known as Riemann--Cartan and offer the natural framework for Poincar\'e gauge theories, where the torsion appears as the field strength of translations, sourced by the spin-current~\cite{Kibble:1961ba,Sciama:1962,Hehl:1976kj}.
 %In such a framework, the vielbein is interpreted as the gauge connection of translations and the torsion as its field strength.
@@
 
-In the framework of the Kaluza--Klein theories in higher dimensional Riemann--Cartan geometries, phenomenology of the extra dimensional torsion was found in Ref.~\cite{Kalinowski:1980da}, metric dependent torsional models of extra dimensions was studied in~\cite{Shankar:2012vd}, along with its consequences in cosmology~\cite{Chen:2009ep}.
+In the framework of the Kaluza--Klein theories in higher dimensional Riemann--Cartan geometries, phenomenology of the extra dimensional torsion was found in Ref.~\cite{Kalinowski:1980da}, metric dependent torsion in extra dimensions was studied in~\cite{Shankar:2012vd}, along with its consequences in cosmology~\cite{Chen:2009ep}.
 Compatification of higher dimensional Brans--Dicke models with torsion was considered in Ref.~\cite{German:1993bq}.
@@
 The theory admits a nonvanishing torsion in vacuum due to the presence of the Gauss--Bonnet term, in contrast to the five-dimensional Einstein--Cartan theory.
-\hl{This are our results, in the intro?}After the dimensional reduction, the extra degrees of freedom generate an energy density and pressure as four-dimensional matter fields.
+After the dimensional reduction, the extra degrees of freedom generate an energy density and pressure as four-dimensional matter fields.
 In some cases, the solutions avoid the appearance of initial singularities and drive the accelerated expansion of the Universe, while the radius of the compact manifold goes to zero.
 
-This work is organized as follows: In Sec.~\ref{KK} we introduce the torsional degrees of freedom compatible with the Kaluza--Klein geometry, and fix our notation and conventions.
+This work is organized as follows: In Sec.~\ref{KK} we review the Kaluza--Klein geometry in the first order formalism and fix our notation and conventions.
 In Sec.~\ref{5EGB}, we study the dynamics of the general Lovelock--Cartan action in five-dimensional spacetime and its dimensional reduction.
@@
 
-\section{Kaluza--Klein Ansatz in Riemann--Cartan Spacetimes\label{KK}}
+\section{Kaluza--Klein Geomatry in Riemann--Cartan Spacetimes\label{KK}}
 
@@
 Every quantity defined on $M_N$ will be denoted by hats $\hat{x}$.
-Capital Greek characters (coordinate indices) and capital Latin characters (Lorentz indices) run over the $N$ dimensions, while lower case run in the $(N-1)$-dimensional reduced manifold.
+Capital Greek characters (coordinate indices) and capital Latin characters (Lorentz indices) run over the $N$ dimensions, i.~e.
+$A = 0,...,N-2,N$, while lower case run in the $(N-1)$-dimensional reduced manifold, i.~e.
+$a = 0,...,N-2$.
 
-The differential structure on $M_N$ is determined by two fields.
-The spin connection 1-form, $\hat{\omega}^{AB} = \hat{\omega}^{AB}{}_{\Gamma}\,\text{d}\hat{x}^\Gamma$, describes the affine structure of the $N$-dimensional manifold, and the vielbein 1-form, $\hat{e}^A=\hat{e}^{A}{}_{\Gamma}\,\text{d}\hat{x}^\Gamma$, defines the metric structure of the same manifold through the relation $\hat{g}_{\Gamma\Delta} = \hat{e}^{A}{}_{\Gamma}\hat{e}^{B}{}_{\Delta}\hat{\eta}_{AB}$.
-The vielbein is the mapping that relates coordinate indices with the Lorentz indices.
-In our convention, both Lorentzian metrics \(\hat{\eta}\) and \(\eta\) have positive signature, and the Levi-Civita symbols are such that
-\[\hat{\epsilon}_{a_1 \cdots a_{N-1} N} \equiv \epsilon_{a_1 \cdots a_{N-1}},\]
+The differential structure of $M_N$ is determined by two fields.
+The spin connection 1-form, $\hat{\omega}^{AB} = \hat{\omega}^{AB}{}_{\Gamma}\,\text{d}\hat{x}^\Gamma$, describes the affine structure of the $N$-dimensional manifold and the vielbein 1-form, $\hat{e}^A=\hat{e}^{A}{}_{\Gamma}\,\text{d}\hat{x}^\Gamma$, defines the metric structure of the same manifold through the relation $\hat{g}_{\Gamma\Delta} = \hat{e}^{A}{}_{\Gamma}\hat{e}^{B}{}_{\Delta}\hat{\eta}_{AB}$, where $\hat{g}_{\Gamma\Delta}$ is the spacetime metric, while $\hat{\eta}_{AB} = \diag(-,+,...,+)$ is the Minkowski metric.
+The vielbein maps coordinate into Lorentz indices.
+In our convention, the Levi-Civita symbol is such that $\hat{\epsilon}_{01...(N-2)N} = 1$ and the reduced symbol is defined by fixing the last index, i.~e.
+\[\hat{\epsilon}_{a_1 \cdots a_{N-1} N} \equiv \epsilon_{a_1 \cdots a_{N-1}}.\]
 %% \[\hat{\epsilon}_{0 1 \cdots N-1} = {\epsilon}_{0 1 \cdots N-1} = 1,\]
-which allows us to define the reduced symbol by fixing the last index.
-Additionally, all Riemannian fields (torsion-free) will be explicited by a tilde, as it was done in Eq.~\eqref{GB} for the curvature.
+Additionally, all the Riemannian fields (torsion-free) will be explicited by a tilde, as it was done in Eq.~\eqref{GB} for the curvature.
 %% In our convention, the flat Lorentzian metric will be $\hat{\eta}_{AB}=\diag(-,+,\cdots,+)$, and the $N$-dimensional Levi-Civita symbol $\hat{\epsilon}_{A_1 \cdots A_N}$ is reduced by defining
@@
 \end{equation}
-where $x$ denotes the coordinates of the $(N-1)$-manifold, collectively. Henceforth, we will focus in the $n=0$ mode of the expasion, also refered to as the low-energy sector.
+where $x$ denotes the coordinates of the $(N-1)$-manifold, collectively. Henceforth, we will focus on the $n=0$ mode of the expasion, also refered to as the low-energy sector.
 
@@
 
-The Kaluza--Klein (KK) ansatz for the metric relies on the premise that the compact dimension of $M_N$ is orthogonal to the rest of the manifold at each point (see for example Ref.~\cite{Blagojevic:2002du,*Hehl:1976kj}).
+The Kaluza--Klein (KK) ansatz for the metric lies on the premise that the compact dimension of $M_N$ is orthogonal to the rest of the manifold at each point.
 This leads to the following metric structure
@@
 \end{equation}
-and its inverse, $\hat{E}_A{}^{\Gamma}$, defined such that $\hat{E}_A{}^{\Gamma}\hat{e}^A{}_{\Delta}=\delta^\Gamma_{\Delta}$, reads
+and its inverse, $\hat{E}_A{}^{\Gamma}$, defined such that $\hat{E}_A{}^{\Gamma}\hat{e}^A{}_{\Delta}=\delta^\Gamma_{\Delta}$ and $\hat{E}_A{}^{\Gamma}\hat{e}^B{}_{\Gamma}=\delta^A_B$, reads
 \begin{equation}
@@
 \end{equation}
-and it shows that the compact manifold $S^1$ has its own tangent space, $T_pS^1$, at each point $p\in M_N$, and is independent of $T_pM_{N-1}$ in the sense that they do not mix.
+where $A_a = E_a{}^\mu A_\mu$.
+This shows that the compact manifold $S^1$ has its own tangent space, $T_pS^1$, at each point $p\in M_N$, and is independent of $T_pM_{N-1}$ in the sense that they do not mix.
 %It seems reasonable to think that if we want two manifolds to be orthogonal to one another, they will not mix vectors after the tangent mapping.
@@
 \end{equation}
-where $\tilde{\omega}^{ab}$ is the Riemannian spin connection of the reduced manifold and $F_{ab}$ are the components of the field strength of $A_b$ defined as
+where $\tilde{\omega}^{ab}$ is the Riemannian spin connection of the reduced manifold and $F_{ab}$ are the components of the field strength of $A = A_m\, e^m$, defined by
 \begin{equation}
@@
 
-Nevertheless, in a Riemann--Cartan geometry $M_N$ is not entirely described by $\hat{\tilde{\omega}}^{AB}$, but by a more general spin connection, independent of the metric degrees of freedom (see Appendix~\ref{Riemann-Cartan} for details).
+In a Riemann--Cartan geometry $M_N$ is not entirely described by $\hat{\tilde{\omega}}^{AB}$, but by a more general spin connection independent of the metric degrees of freedom (see Appendix~\ref{Riemann-Cartan} for details).
 We can assume a general connection 1-form of the same type (regarding its $M_{N-1}\times S^1$ decomposition) as in Eq.~\eqref{DRiemannconnection}.
@@
 
-Similarly from the definition of torsion in Eq.~\eqref{tordef}, we find its distinctive parts to be
+Similarly from the definition of torsion in Eq.~\eqref{tordef}, we find its distinctive parts to be~\footnote{Henceforth we will refer to as distinctive parts regarding the $M_{N-1}\times S^1$ decomposition.}
 \begin{align}\label{T1}
@@
 \end{equation}
-The second Bianchi identity, gives nothing but its equivalent for the manifold $M_{N-1}$, this is
+This is taken as a proof of the tensorial nature of these new fields under the four-dimensional Lorentz transformations.
+The second Bianchi identity gives nothing but its equivalent for $M_{N-1}$, this is
 \begin{equation*}
@@
 \end{equation*}
-
-This is taken as a proof of the tensorial nature of these new fields under the four-dimensional Lorentz transformations.
 
@@
 \wedge\hat{e}^{A_N}$.} is the Lovelock--Cartan theory of gravity~\cite{Mardones:1990qc}.
-This is the natural generalization of Lanczos--Lovelock when torsional degrees of freedom are present.
+This is the natural generalization of Lanczos--Lovelock action when torsional degrees of freedom are present.
+Its simplest realization is the Einstein--Cartan model and their dimensional reduction is indistinguishable from the Riemannian case (see for instance~\cite{German:1993bq,Aros:2007nn}).
 
-In five dimensions, the action principle of such a theory is given by  
+In five dimensions, the Lovelock--Cartan theory is given by the following action principle 
 \begin{widetext}
@@
 
-In terms of the KK ansatz presented in the previous section, the field equations can be decomposed into its distinctive parts, regarding
-its \(M_{4} \times S^1\) decomposition. The Eq.~\eqref{delta e} leads to 
+In terms of the KK ansatz presented in the previous section, the field equations can be decomposed into its distinctive parts. The Eq.~\eqref{delta e} leads to 
 \setlength\multlinegap{0pt}
@@
 
-In order to look for cosmological solutions, we demand the symmetries assumed by the cosmological principle, i.e.
+In order to look for cosmological solutions, we demand the symmetries assumed by the cosmological principle, i.~e.
 isotropy and homogeneity of the involved fields.
@@
 
-The $1$-form $A$ is $A(t)\text{d}t$, has vanishing field strength $F$.
+For this ansatz, the $1$-form $A = A(t)\text{d}t$ has vanishing field strength $F$.
 Thus, without loss of generality, it can be fixed to zero by means of an $U(1)$ gauge transformation or, equivalently, a diffeomorphism transformation along the $z$-direction.
@@
 
-Due to Eq.~\eqref{u}, the solutions will be valid in the region of the parameter space where $\Delta>0$.
-The function $\omega(t)$ satisfies the equation\footnote{Notice that in our choice of parameters $u_+$ cannot vanish.}
+Due to Eq.~\eqref{u}, the solutions will be valid in the region of the parameter space where $\Delta>0$. The function $\omega(t)$ satisfies the equation
 \begin{equation}
@@
 \end{equation}
-which, for the three significantly different values of $u_\pm$, has the following solutions\footnote{Notice that in our choice of parameters $u_+$ cannot vanish.}
+which, for the three significantly different values of $u_\pm$, has the following solutions
 \begin{equation}
@@
   \begin{cases}
-    -\sqrt{u_\pm}\tan\left[\sqrt{u_\pm}\left(t-t_0\right)\right] & u_\pm > 0 \\
-    \left(t-t_0\right)^{-1} & u_- = 0 \\
-    \sqrt{-u_\pm}\tanh\left[\sqrt{-u_\pm}\left(t-t_0\right)\right] & u_\pm < 0
+    -\sqrt{u_\pm}\tan\left[\sqrt{u_\pm}\left(t-t_0\right)\right], & u_\pm > 0 \\
+    \left(t-t_0\right)^{-1}, & u_\pm = 0 \\
+    \sqrt{-u_\pm}\tanh\left[\sqrt{-u_\pm}\left(t-t_0\right)\right], & u_\pm < 0
   \end{cases}
@@
 
-Hereon, we will consider $\alpha_1$ to be positive,\footnote{In all the plots, we fixed $\alpha_1 = 1$.} because this parameter admits the interpretation of the Newton's gravitational constant.
+Hereon, we will consider $\alpha_1$ to be positive,\footnote{In all the plots we normalize $\alpha_1 = 1$.} because this parameter admits the interpretation of the Newton's gravitational constant.
 It is worth mentioning the existence of models of gravity which are free of the linear curvature term.
@@
 Even though the analysis that follows will consider a positive non-zero value for $\alpha_1$, the solutions listed above are also valid in the regime where $\alpha_1=0$, that constitutes the non-Riemannian extension of pure Lovelock theories.
+
+\subsubsection{Bouncing solutions}
+
+In Fig.~\ref{cond_u_pos} we show the allowed region, in the $\alpha_0$--$\alpha_2$ parameter space, for $u_\pm$ to be positive.
+
 \begin{figure}[H]
@@
 \end{figure}
-
-\subsubsection{Bouncing solutions}
-
-In Fig.~\ref{cond_u_pos} we show the allowed region, in the $\alpha_0$--$\alpha_2$ parameter space, for $u_\pm$ to be positive.
 
@@
 Moreover, the bouncing behavior of the extra dimension is given by the scalar field in a phase difference of $\pi/2$ with respect to $a(t)$, as shown in Fig.~\ref{phi_u_pos}.
-It is worth to point that the {intermediate bounce} present in the scale factor is much prominent in the modulation of the extra dimension.
+%It is worth to point that the {intermediate bounce} present in the scale factor is much prominent in the modulation of the extra dimension.
 \begin{figure}[H]
   \includegraphics[width=.95\linewidth]{a_u_pos.pdf}
-  \caption{Behavior of the scale factor, \(a(t)\), as a function of the scaled time, $\tau$, for $u_\pm > 0$.}
+  \caption{Behavior of the scale factor, \(a(t)\), as a function of the scaled time, $\tau$, for $u_\pm > 0$. We have normalized using the maximum value of the scale factor, $a_{\mathrm{max}}$.}
   \label{a_u_pos}
@@
   \includegraphics[width=.95\linewidth]{phi_u_pos.pdf}
-  \caption{Behavior of the $\phi$ field as a function of the scaled time, $\tau$, for $u_\pm > 0$.}
+  \caption{Behavior of the $\phi$ field as a function of the scaled time, $\tau$, for $u_\pm > 0$. We have normalized using the maximum value of the scalar field, $\phi_{\mathrm{max}}$.}
   \label{phi_u_pos}
@@
 
-None of the other cases, $u_- = 0$ and $u_\pm < 0$, present bouncing behavior.
-The allowed parameter space for $u_+$ and $u_-$ to be negative is shown in Fig.~\ref{cond_u_neg}, while the restriction $u_- = 0$ sets $\alpha_2 = \tfrac{1}{12 \alpha_0}$.
-Nonetheless, depending on the values of the parameters there are three distinctive conducts: (i)~eternal expansion, (ii)~eternal contraction, and (iii)~initial expansion with a final collapse.
+%None of the other cases, $u_- = 0$ and $u_\pm < 0$, present bouncing behavior.
+The allowed parameter space for $u_+$ and $u_-$ to be negative is shown in Fig.~\ref{cond_u_neg}, while the restriction $u_- = 0$ sets $\alpha_2 = \tfrac{1}{12 \alpha_0}$.~\footnote{Notice that in our choice of parameters $u_+$ cannot vanish.}
 
-\paragraph{The case $u_- = 0$} has two behaviors depending on whether $\alpha_0$ is positive or negative.
+\begin{figure}[H]
+  \includegraphics[width=.95\linewidth]{cond_u_neg.pdf}
+  \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be positive defined.}
+  \label{cond_u_neg}  
+\end{figure}
 
+In the regime where $u_\pm \leq 0$, there are three distinctive conducts: (i)~eternal expansion, (ii)~eternal contraction, and (iii)~initial expansion followed by an eternal contraction.
+In all these three scenarios, the size of the extra dimension, modulated by the $\phi(t)$ field, is reciprocal to the scale factor $a(t)$ regarding their expansive/contractive asymptotic behavior.
+
+\paragraph{The case $u_- = 0$} has two behaviors depending on whether $\alpha_2$ is positive or negative.
 As shown in Fig.~\ref{a_u_0}, for positive $\alpha_2$ the solutions fit into the third category above, while for negative $\alpha_2$ the solutions expand eternally.
-Notice that in the former case the Universe's fate is an asymptotic collapse, whilst in the latter expands forever.
+%Notice that in the former case the Universe's fate is an asymptotic collapse, whilst in the latter expands forever.
 \begin{figure}[H]
@@
 \paragraph{The case $u_\pm < 0$} has solutions which either expand or collapse eternally, as shown in Fig.~\ref{a_u_neg}.
-The typical evolution of the Universe with $u_+ > 0$ is to grow infinitely, while for negative $u_-$ the expansion (contraction) corresponds to $\alpha_2$ positive (negative).
-Figure~\ref{phi_u_neg}, shows that the asymptotic {size} of the extra dimension, modulated by the $\phi$ field, is approximately reciprocal to the behavior of the scale factor but it is not monotonic.
+The typical evolution of the Universe with $u_+ < 0$ is to grow infinitely, while for negative $u_-$ the expansion (contraction) corresponds to $\alpha_2$ positive (negative).
+Figure~\ref{phi_u_neg} shows the behaviour of $\phi$ for this case.%, which presents an asymptotic contraction (expansion) when the scale factor expands (contracts) eternally.
 %% , depending on the sign
@@
 %% bound.
-For all these solutions the scale factor remains finite at $t = 0$, and they do not present an initial singularity.
+For all these solutions the scale factor remains finite at $t = 0$, presenting no initial singularity.
 It has been reported that the Gauss--Bonnet term can prevent the Universe to expand from an initial singularity~\cite{Deruelle:1986iv,Henriques:1986jw,*Ishihara:1986if}, which is also this case.
-
-\begin{figure}[H]
-  \includegraphics[width=.95\linewidth]{cond_u_neg.pdf}
-  \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be positive defined.}
-  \label{cond_u_neg}  
-\end{figure}
 
@@
 
-The Eq.~\eqref{equation} can be split into its Riemannian and non-Riemannian parts, by means of the identity~\eqref{curvature decomp}, taking the familiar form
+The Eq.~\eqref{equation} can be written into its familiar form, by means of the identity~\eqref{curvature decomp}, as
 \begin{align}\label{einstein equation}
@@
 The contributions of $F$, $\alpha^{ab}$ and $\beta^a$ have not been taken into account since, for all practical purposes of this article, they vanish.
-%% The left hand side of the Eq.~\eqref{einstein equation} is the Einstein $3$-form, which is dual to the Einstein tensor with cosmological constant in four dimensions.
-The left hand side of Eq.~\eqref{einstein equation} is the 3-form which, in components, yields the four-dimensional Einstein equations with cosmological constant.
-We identify the right hand side as an effective energy-momentum $3$-form, which behaves as a perfect fluid. It contains the torsional and the higher dimensional degrees of freedom.
+The left hand side of the Eq.~\eqref{einstein equation} is the $3$-form whose Hodge dual yields to the Einstein equations with cosmological constant in four dimensions.
+%%The left hand side of Eq.~\eqref{einstein equation} is the 3-form which, in components, yields the four-dimensional Einstein equations with cosmological constant.
+We identify the right hand side as a $3$-form effective energy-momentum, which behaves as a perfect fluid. It contains the torsional and the higher dimensional degrees of freedom.
 
@@
 
-In this paper we have presented the dimensional reduction of the five-dimensional Lovelock--Cartan theory introduced in Ref.~\cite{Mardones:1990qc}, under the assumption that the compact dimension has $S^1$ topology.
+In this work we have presented the dimensional reduction of the five-dimensional Lovelock--Cartan theory introduced in Ref.~\cite{Mardones:1990qc}, under the assumption that the compact dimension has $S^1$ topology.
 An interesting feature of these theories is that, unlike the Einstein--Cartan theory, torsion is allowed to propagate.
@@
 We used the most general ansatz compatible with the cosmological symmetries (see Appendix~\ref{homotropic}) to find solutions of the Friedmann--Robertson--Walker class in the effective theory.
-The equations of motion [Eqs.~\eqref{eqn1}--\eqref{eqn2}] ensures that solutions with nontrivial torsion have vanishing $\alpha^{ab}$ and $\beta^a$.
+The equations of motion for the cosmological ansatz ensure that the solutions with nontrivial torsion have vanishing $\alpha^{ab}$ and $\beta^a$.
 
-The solutions are parametrized by a variable $u_\pm$, related to the fundamental couplings of the theory through the Eq.~\eqref{u}, and exhibit three different sectors depending on whether this parameter is positive, negative or zero.
+The solutions are parametrized by $u_\pm$, related to the fundamental couplings of the theory through the Eq.~\eqref{u}, and exhibit three different sectors depending on whether this parameter is positive, negative or zero.
 
-The behavior of the Universes described by our solutions is four-fold: (i)~eternal expansion, (ii)~eternal contraction, (iii)~initial expansion with a final collapse, and (iv)~bouncing.
-We show the typical evolution of the scale factor, for each case, in Figs.~\ref{a_u_pos}, ~\ref{a_u_0} and~\ref{a_u_neg}.
-The size of the extra dimension is dynamical, and there are sectors in the parameter space where there exists an asymptotic dynamical mechanics of compactification (see Figs.~\ref{phi_u_0} and~\ref{phi_u_neg}).
+The behavior of the Universes described by our solutions is four-fold: (i)~eternal expansion, (ii)~eternal contraction, (iii)~initial expansion with asymptotic contraction, and (iv)~bouncing.
+%We show the typical evolution of the scale factor for each case in Figs.~\ref{a_u_pos}, ~\ref{a_u_0} and~\ref{a_u_neg}.
+The size of the extra dimension is driven by the scalar field $\phi(t)$ and its behavior at late times is compatible with a dynamical mechanism for compactification, at certain regions of the parameter space (see Figs.~\ref{phi_u_0} and~\ref{phi_u_neg}).
 
-After the decomposition of the curvature into Riemannian and non-Riemannian parts, the Eq.~\eqref{equation} takes the form of the usual Einstein's equations where the extra fields, coming from the extra dimensional and torsional degrees of freedom, originate an effective energy-momentum tensor.
-It is important to mention that this tensor is nonstandard, in the sense that it inherit a coupling between matter and curvature---see Eq.~\eqref{taua} for the definition---, due to the non-minimal coupling among the four-dimensional gravity and the induced fields.
-Moreover, the energy density and the pressure are not positive definite, which allows us to interpret these induced matter as exotic, providing a playground for finding gravitational explanation to dark matter/energy.
+% takes the form of the usual Einstein equations where the fields coming from the extra dimensional and torsional degrees of freedom, originate an effective energy-momentum tensor. 
+The field equation~\eqref{equation} admits the interpretation of the Einstein equations, where the energy-momentum $3$-form is induced by the torsion and the extra dimensional degrees of freedom.
+It is important to mention that the induced energy-momentum form is nonstandard, in the sense that it inherit a coupling between matter and curvature---see Eq.~\eqref{taua} for the definition---, due to the nonminimal coupling in the reduced theory.
+Moreover, the energy density and the pressure are not positive definite, which allows us to interpret these induced fields as exotic, providing a playground for a gravitational explanation for the accelerated expansion of the Universe.
 
@@
 \begin{acknowledgments}
-  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, A.~Toloza and J.~Zanelli for enlightening discussions on this work.
+  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, R.~Gannouji, A.~Toloza and J.~Zanelli for enlightening discussions on this work.
 \mbox{O.~C.-F.} has been partially supported by the PAI project No.~79140040 (CONICYT--Chile).
-The Centro Cient\'ifico y Tecnol\'ogico de Valpara\'iso (CCTVal) is funded by the Chilean Government through the Centers of Excellence Basal Financing Program of CONICYT.
+The Centro Cient\'ifico Tecnol\'ogico de Valpara\'iso (CCTVal) is funded by the Chilean Government through the Centers of Excellence Basal Financing Program of CONICYT.
 \end{acknowledgments}
@@
 
-In Riemann--Cartan geometry, both the vielbein $e^A$ and the spin connection $\omega^{AB}$ are independent features of the manifold.
-We will drop the hats since we are not dealing with the KK decomposition.
-The spin connection, however, can be decomposed in a Riemannian part, which is metric dependent, $\tilde{\omega}^{AB}$ satisfying $\tilde{\mbox{D}}e^A=0$, and a contorsion piece $\kappa^{AB}=-\kappa^{BA}$, such that
+In Riemann--Cartan geometry, both the vielbein $e^A$ and the spin connection $\omega^{AB}$ are independent features of the manifold.~\footnote{Here, we will drop the hats since we will not deal with the KK decomposition.} The spin connection can be decomposed in a Riemannian part, which is metric dependent, $\tilde{\omega}^{AB}$ satisfying $\tilde{\mbox{D}}e^A=0$, and a contorsion piece $\kappa^{AB}=-\kappa^{BA}$, such that
 \begin{equation}\label{spin separation}
@@
 These two assumptions are translated in the Killing equations, which are the vanishing of the Lie derivatives of the fields along the vectors which generate the symmetries $\{\zeta^\lambda_{i}\}$.
+
+The set of Killing vectors $\{\zeta^\lambda_i\}$ are the generators of $SO(3)$, which generate the spatial rotations in three dimensions, $\mathcal{J}_i=\epsilon_{ijk}x_j\partial_k$, and the Killing vectors assosiated with spatial translations $\mathcal{P}_i=\sqrt{1-kr^2}\partial_i$, satisfy the algebra
+\begin{align*}
+  \left[\mathcal{J}_i,\mathcal{J}_j\right]&=\epsilon_{ijk}\mathcal{J}_k,\\
+  \left[\mathcal{P}_i,\mathcal{P}_j\right]&=-k\epsilon_{ijk}\mathcal{J}_k,\\
+  \left[\mathcal{J}_i,\mathcal{P}_j\right]&=\epsilon_{ijk}\mathcal{P}_k.
+\end{align*}
+
 In particular the Killing equations for the metric tensor and the torsion tensor are
@@
 and whose Killing equations are analogous to the tensorial, Eq.~\eqref{killing metric}, and vectorial one, Eq.~\eqref{killing vector}.
+These requirements on the fields are translated to a set of first order differential equations whose most general solutions determines our ansatz structure from Eqs.~\eqref{vielbein cosmo} to~\eqref{gamma cosmo}.
 
-The set of Killing vectors $\{\zeta^\lambda_i\}$ are the generators of $SO(3)$, which generate the spatial rotations in three dimensions, $\mathcal{J}_i=\epsilon_{ijk}x_j\partial_k$, and the Killing vectors assosiated with spatial translations $\mathcal{P}_i=\sqrt{1-kr^2}\partial_i$, satisfy the algebra
-\begin{align*}
-  \left[\mathcal{J}_i,\mathcal{J}_j\right]&=\epsilon_{ijk}\mathcal{J}_k,\\
-  \left[\mathcal{P}_i,\mathcal{P}_j\right]&=-k\epsilon_{ijk}\mathcal{J}_k,\\
-  \left[\mathcal{J}_i,\mathcal{P}_j\right]&=\epsilon_{ijk}\mathcal{P}_k.
-\end{align*}
-These requirements on the fields are translated to a set of first order differential equations whose most general solutions determines our ansatz structure from Eqs.~\eqref{vielbein cosmo} to~\eqref{gamma cosmo}.
-The field strengths for isotropic--homogeneus vielbein and spin connection are written next.
-The curvature components are
+The components of the Lorentz curvature and torsion for the isotropic--homogeneus ansatz are 
 \begin{align*}
@@
 \end{align*}
-and the torsion components
+and 
 \begin{equation}\label{homotorsion}
@@
 \end{equation}
-where $h(t)$ and $H(t)$ were defined in Sec.~\ref{5EGB}.
+respectively. The functions $a(t)$, $\omega(t)$, $f(t)$, $h(t)$ and $H(t)$ were defined in Sec.~\ref{cosmos}.
 
@@
 
-We supply here the time-dependent expressions for the cosmological solutions, given the differents values of $u_\pm$, in terms of the dimensionless parameter \mbox{$\tau=\sqrt{|u_\pm|}(t-t_0)$}.
+We supply here the time-dependent expressions for the cosmological solutions, given the different values of $u_\pm$, in terms of the dimensionless parameter \mbox{$\tau=\sqrt{|u_\pm|}(t-t_0)$}.
+\vspace{.1cm}
 
@@
     f^2(t)&=u_\pm\tfrac{k}{a_0^2}\sec^2\tau\exp\left[\tfrac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cos^2\tau\right]\notag\\
-    &+u_\pm\tan^2\tau + \tfrac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
+    &\quad +u_\pm\tan^2\tau + \tfrac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
   \end{align}
@@
     f^2(t)&=\tfrac{k}{a_0^2(t-t_0)^2}\exp\left[\tfrac{\mp\sqrt{6\Delta}}{12\alpha_2}(t-t_0)^2\right]\notag\\
-    &+(t-t_0)^{-2} + \tfrac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
+    &\quad +(t-t_0)^{-2} + \tfrac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
   \end{align}
@@
     f^2(t)&=-u_\pm\tfrac{k}{a_0^2}\cosh^{-2}\tau\exp\left[\tfrac{\mp\sqrt{6\Delta}}{12\alpha_2u_\pm}\cosh^2\tau\right]\notag\\
-    &-u_\pm\tanh^2\tau + \tfrac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
+    &\quad -u_\pm\tanh^2\tau + \tfrac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2}.
   \end{align}

### 2016-10-06 KKosmology.tex
Last revision altogether

@@
 
-\author{Cristobal \surname{Corral}}
+\author{Crist\'obal \surname{Corral}}
 \email{cristobal.corral@usm.cl}
@@
 \author{Francisca \surname{Ram\'irez}.}
+\email{francisca.ramirez@alumnos.usm.cl}
 \affiliation{\UTFSM.}
@@
 \begin{abstract}
-  We study the Kaluza--Klein dimensional reduction of the Lovelock--Cartan theory on a five-dimensional manifold, where the compact dimension has $S^1$ topology.
-We find cosmological solutions of the Friedmann--Robertson--Walker class for the effective theory.
-The torsional degrees of freedom induce a nonvanishing energy-momentum tensor in four dimensions.
+  We study the Kaluza--Klein dimensional reduction of the Lovelock--Cartan theory in five-dimensional spacetime, with a compact dimension of $S^1$ topology.
+We find cosmological solutions of the Friedmann--Robertson--Walker class in the reduced spacetime.
+The torsion and the fields arising from the dimensional reduction induce a nonvanishing energy-momentum tensor in four dimensions.
 We find solutions describing expanding, contracting and bouncing universes.
@@
 
-The experimental status of General Relativity (GR), regarding the solar system tests~\cite{Will:2014kxa} and the detection of signals consistent with the merge of two black holes by the LIGO collaboration~\cite{Abbott:2016blz,*Abbott:2016nmj}, have settled it as the most successful theory of gravity so far.
-However, the difficulties on finding out explanations for the so-called \emph{dark sector} of the Universe, have driven the community to think that GR is not the ultimate gravitational theory.
-The \emph{dark sector} of the Universe is composed by two kinds of degrees of freedom, whose effects are detected at different scales.
-On the one hand, at galactic scales, the gravitational lensing produced by the local distribution of energy, suggests the presence of an exotic form of matter, unseen by the current light-based telescopes~\cite{Sofue:2000jx}, for this it is named dark matter.
-Such an abundance in the galaxy is also compatible with the velocity profile of stars at its outter regions and it is needed to achieve the current structures formation.
-This matter would interact mostly (if not only) through gravity.
-On the other hand, at cosmic scales, the experimental data obtained from Type Ia supernovae observations, indicates that our Universe is passing through a phase of accelerated expansion~\cite{Riess:1998cb}.
-This introduces the dark energy hypothesis to generate such a behavior.
+The experimental status of General Relativity (GR), regarding the solar system tests~\cite{Will:2014kxa} and the detection of signals consistent with the merge of two black holes by the LIGO collaboration~\cite{Abbott:2016blz,*Abbott:2016nmj}, have settled it as the most successful theory of gravity.
+However, the difficulties on finding out explanations for the so-called \emph{dark sector} of the universe, have driven the community to think that GR is not the ultimate gravitational theory.
+
+The \emph{dark sector} of the universe is composed by two kinds of degrees of freedom.
+On the one hand, the gravitational lensing produced by the local distribution of energy, suggests the presence of an exotic form of matter unseen by the current light-based telescopes~\cite{Wittman:2000tc}.
+For this it is named dark matter and it would interact mostly (if not only) through gravity.
+Such an abundance at the galactic scale is compatible with the velocity profile of stars at its outter regions~\cite{Sofue:2000jx}.
+At the cosmic scale, it plays a key role in the origin and evolution of structures (see, for instance Ref.~\cite{DelPopolo:2008mr}).
+On the other hand, the experimental data obtained from Type Ia Supernovae observations, indicates that our universe is passing through a phase of accelerated expansion~\cite{Riess:1998cb}.
+This behavior suggests the existence of an exotic form of energy, called dark energy, which constitutes roughly the $70\%$ of the current content of our universe.
 
@@
 This idea opened the possibility of a novel geometrical understanding of interactions, where the gauge group arise as a consequence of the topology of the spatial compact manifold in a higher dimensional spacetime.
-The idea of higher dimensions comes naturally in diverse physical models, as for example: supersymmetry and supergravity~\cite{supergravity}, string theory~\cite{Green:1987sp,*Green:1987mn}, and novel proposals by Arkani-Hammed, Dimopoulos and Dvali~\cite{ArkaniHamed:1998rs,*Antoniadis:1998ig}, and Randall and Sundrum~\cite{Randall:1999ee,*Randall:1999vf} models, as attempts of solving the hierarchy problem.
+The idea of higher dimensions comes naturally in diverse physical models, as for example: supersymmetry and supergravity~\cite{supergravity}, string theory~\cite{Green:1987sp,*Green:1987mn}, and novel proposals by Arkani-Hammed, Dimopoulos and Dvali~\cite{ArkaniHamed:1998rs,*Antoniadis:1998ig}, and Randall and Sundrum~\cite{Randall:1999ee,*Randall:1999vf} models, as attempts to solve the hierarchy problem.
 
 In four dimensions, the Einstein--Hilbert action with cosmological constant is the most general theory which leads to second order field equations for the metric.
-In higher dimensions, however, particular combinations of higher order terms in the curvature can be added to the gravitational action, giving rise to second order field equations for the metric.
-The most general theory in arbitrary dimensions, which preserves the features of the four-dimensional Einstein--Hilbert, is described by the Lanczos--Lovelock action~\cite{Lanczos:1938sf,*Lovelock:1971yv}.
+In higher dimensions, however, particular combinations of higher order terms in the curvature can be added to the gravitational action, whose variation with respect to the metric yields also second order field equations.
+The most general theory in arbitrary dimensions, which preserves this feature of the four-dimensional Einstein--Hilbert, is named the Lanczos--Lovelock action~\cite{Lanczos:1938sf,*Lovelock:1971yv}.
 Such a theory, has no ghosts~\cite{Zumino:1985dp} and it has the same degrees of freedom as the Einstein--Hilbert Lagrangian in arbitrary dimensions~\cite{Henneaux:1990au}.
@@
 $\tilde{R}_{\mu\nu}$ and $\tilde{R}$ are the Ricci tensor and Ricci scalar, respectively.
-
-The Gauss--Bonnet term, in four dimensions, adds no dynamics to the metric, since it represents a topological invariant proportional to the Euler characteristic class, which can be written locally as a boundary term.
+In four dimensions, the Gauss--Bonnet term adds no dynamics to the metric, since it represents a topological invariant proportional to the Euler characteristic class, which can be written locally as a boundary term.
 Nevertheless, in dimensions higher than four it contributes to the field equations.
@@
 For instance, exact wormhole solutions have been found in vacuum~\cite{Dotti:2006cp,*Dotti:2007az} and coupled with matter fields~\cite{Mehdizadeh:2015jra}.
-In higher order Lanczos--Lovelock models, this class of exact solutions have been reported in~\cite{Mehdizadeh:2015dta} and in higher order compactified Lanczos--Lovelock theories with torsion in~\cite{Canfora:2008ka}.
+In higher order Lanczos--Lovelock models, this class of exact solutions have been reported in~\cite{Mehdizadeh:2015dta} and in the compactified theory with torsion~\cite{Canfora:2008ka}.
 Additionally, the compactification of higher Lanczos--Lovelock terms has been considered in Refs.~\cite{MuellerHoissen:1985mm,*MuellerHoissen:1989yv} and their cosmology in Refs.~\cite{MuellerHoissen:1985ij,Deruelle:1986iv,Deruelle:1989fj}.
-It is worth mentioning that, classically, wormholes must be supported by a kind energy compatible with the cosmological hypothesis.
-Their existence in these extended models implies the presence of degrees of freedom that could provide an explanation for the exotic matter/energy abundance in the Universe.
+It is worth mentioning that, classically, wormholes must be supported by a kind of energy compatible with the cosmological hypothesis.
+Their existence in these extended models implies the presence of degrees of freedom that could provide an explanation for the exotic matter/energy abundance in the universe.
 
@@
 %For instance, the pioneer works of Kopczy\'nski showed how the introduction of a spinning dust, acting as a source of the curvature and torsion, avoids the appearance of cosmological singularities~\cite{Kopczynski:1972,Kopczynski:1973}.
-In the context of the early evolution of the Universe, the spin density can drive the inflationary phase within the Einstein--Cartan theory, along with the avoidance of singularities~\cite{Gasperini:1986mv}.
-It has been shown in Ref.~\cite{Shie:2008ms} that torsional degrees of freedom might produce the current phase of accelerated expansion of the Universe.
+In the context of the early evolution of the universe, the spin density can drive the inflationary phase within the Einstein--Cartan theory, along with the avoidance of singularities~\cite{Gasperini:1986mv}.
+It has been shown in Ref.~\cite{Shie:2008ms} that torsional degrees of freedom might produce the current phase of accelerated expansion of the universe.
 Additionally, the nonminimal coupling of the Euler characteristic form with a scalar field in Riemann--Cartan spacetimes, allows the torsional degrees of freedom to behave as an exotic induced-matter content, customary for the phase of accelerated expansion~\cite{Toloza:2013wi}.
@@
 The theory admits a nonvanishing torsion in vacuum due to the presence of the Gauss--Bonnet term, in contrast to the five-dimensional Einstein--Cartan theory.
-After the dimensional reduction, the extra degrees of freedom generate an energy density and pressure as four-dimensional matter fields.
-In some cases, the solutions avoid the appearance of initial singularities and drive the accelerated expansion of the Universe, while the radius of the compact manifold goes to zero.
+The new degrees of freedom coming from the higher dimensional geometry, generate an induced energy-momentum form in the reduced theory.
+In some cases, the solutions avoid the appearance of initial singularities and drive the accelerated expansion of the universe, while the radius of the compact manifold goes to zero.
 
 This work is organized as follows: In Sec.~\ref{KK} we review the Kaluza--Klein geometry in the first order formalism and fix our notation and conventions.
-In Sec.~\ref{5EGB}, we study the dynamics of the general Lovelock--Cartan action in five-dimensional spacetime and its dimensional reduction.
+In Sec.~\ref{5EGB}, we study the dynamics of the general Lovelock--Cartan action in a five-dimensional spacetime and its dimensional reduction.
 In Sec.~\ref{cosmos}, we look for cosmological solutions of the Friedmann--Robertson--Walker class with nontrivial torsion.
@@
 
-\section{Kaluza--Klein Geomatry in Riemann--Cartan Spacetimes\label{KK}}
+\section{Kaluza--Klein Geometry in Riemann--Cartan Spacetimes\label{KK}}
 
@@
 Capital Greek characters (coordinate indices) and capital Latin characters (Lorentz indices) run over the $N$ dimensions, i.~e.
-$A = 0,...,N-2,N$, while lower case run in the $(N-1)$-dimensional reduced manifold, i.~e.
-$a = 0,...,N-2$.
+\mbox{$A = 0,...,N-2,N$}, while the lower case ones run in the $(N-1)$-dimensional reduced manifold, i.~e.
+\mbox{$a = 0,...,N-2$}.
 
@@
     -A^{\delta} & \phi^{-1}+A^2
-  \end{pmatrix}.
+  \end{pmatrix},
 \end{equation}
-It introduces a scalar field $\phi$ and a vector field $A_\mu$ as new gravitational degrees of freedom. The vielbein which holds this structure for the metric has the following form
+defined such that $\hat{g}_{\Gamma\Delta}\hat{g}^{\Delta\Lambda}=\delta^\Lambda_\Gamma$.
+This metric introduces a scalar field $\phi$ and a vector field $A_\mu$ as new gravitational degrees of freedom.
+The vielbein which holds this structure for the metric has the following form
 \begin{equation}
@@
 \end{equation}
-and its inverse, $\hat{E}_A{}^{\Gamma}$, defined such that $\hat{E}_A{}^{\Gamma}\hat{e}^A{}_{\Delta}=\delta^\Gamma_{\Delta}$ and $\hat{E}_A{}^{\Gamma}\hat{e}^B{}_{\Gamma}=\delta^A_B$, reads
+and its inverse, $\hat{E}_A{}^{\Gamma}$, defined such that $\hat{E}_A{}^{\Gamma}\hat{e}^A{}_{\Delta}=\delta^\Gamma_{\Delta}$ and $\hat{E}_A{}^{\Gamma}\hat{e}^B{}_{\Gamma}=\delta^B_A$, reads
 \begin{equation}
@@
 
-In a Riemann--Cartan geometry $M_N$ is not entirely described by $\hat{\tilde{\omega}}^{AB}$, but by a more general spin connection independent of the metric degrees of freedom (see Appendix~\ref{Riemann-Cartan} for details).
+In a Riemann--Cartan geometry, $M_N$ is not entirely described by $\hat{\tilde{\omega}}^{AB}$, but by a more general spin connection independent of the metric degrees of freedom (see Appendix~\ref{Riemann-Cartan} for details).
 We can assume a general connection 1-form of the same type (regarding its $M_{N-1}\times S^1$ decomposition) as in Eq.~\eqref{DRiemannconnection}.
@@
   \hat{R}^{Na}&=-\left(\mbox{D}\beta^a+\sqrt{\phi}\gamma^a F\right)\notag\\
-  & \quad +\left(\alpha^a_{\ b}\beta^b-\mbox{D}\gamma^a-\frac{1}{2}\gamma^a\mbox{d}\ln\phi\right)\wedge\hat{e}^N.
+  & \quad +\left(\alpha^a_{\ b}\beta^b-\mbox{D}\gamma^a-\frac{1}{2}\gamma^a\mbox{d}\ln\phi\right)\wedge\hat{e}^N,
 \end{align}
-
+where $R=\text{d}\omega+\omega\wedge\omega$ is the curvature of the reduced spacetime $M_{N-1}$.
 Similarly from the definition of torsion in Eq.~\eqref{tordef}, we find its distinctive parts to be~\footnote{Henceforth we will refer to as distinctive parts regarding the $M_{N-1}\times S^1$ decomposition.}
@@
   \label{T2}
-  \hat{T}^N &= \sqrt{\phi}F-\beta_b\wedge e^b+\left(\frac{1}{2}\mbox{d}\ln\phi+\gamma_be^b\right)\wedge\hat{e}^N.
+  \hat{T}^N &= \sqrt{\phi}F-\beta_b\wedge e^b+\left(\frac{1}{2}\mbox{d}\ln\phi+\gamma_be^b\right)\wedge\hat{e}^N,
 \end{align}
+where $T=\text{d}e+\omega\wedge e$ is the torsion $2$-form of $M_{N-1}$.
 
@@
 \begin{equation}
-  \hat{\text{D}}\hat{R}^{AB} = 0 \quad \text{and} \quad \hat{\text{D}}\hat{T}^A = \hat{R}^A_{\ \ B}\wedge\hat{e}^B.
+  \hat{\text{D}}\hat{R}^{AB} = 0 \quad \text{and} \quad \hat{\text{D}}\hat{T}^A = \hat{R}^A_{\ B}\wedge\hat{e}^B.
 \end{equation}
@@
   \begin{aligned}
-    \text{D}R^{ab} &=0, & \text{D}\mbox{D}\alpha^{ab} &=R^a_{\ \ l}\alpha^{lb}+R^b_{\ \ l}\alpha^{al},\\
-    \text{D}\text{D}\beta^a &= R^a_{\ \ b}\beta^{b}, & \text{D}\text{D}\gamma^a &= R^a_{\ \ b}\gamma^{b}.
+    \text{D}R^{ab} &=0, & \text{D}\mbox{D}\alpha^{ab} &=R^a_{\ l}\alpha^{lb}+R^b_{\ l}\alpha^{al},\\
+    \text{D}\text{D}\beta^a &= R^a_{\ b}\beta^{b}, & \text{D}\text{D}\gamma^a &= R^a_{\ b}\gamma^{b}.
   \end{aligned}
 \end{equation}
-This is taken as a proof of the tensorial nature of these new fields under the four-dimensional Lorentz transformations.
-The second Bianchi identity gives nothing but its equivalent for $M_{N-1}$, this is
+This is taken as a proof of the tensorial nature of these new fields under the four-dimensional Lorentz transformations.~\footnote{The tensorial nature of these fields can be also derived from the decomposition of the transformation rule of $\hat{\omega}^{AB}$ under the Lorentz group.} The second Bianchi identity gives its equivalent for $M_{N-1}$, this is
 \begin{equation*}
-  \mbox{D}T^a=R^a_{\ \ b}\wedge e^b.
+  \mbox{D}T^a=R^a_{\ b}\wedge e^b.
 \end{equation*}
@@
     \label{delta e}
-    \hat{\epsilon}_{ABCDE}\Big(\alpha_0\hat{e}^A\wedge\hat{e}^B\wedge\hat{e}^C\wedge\hat{e}^D
-    + \alpha_1\hat{R}^{AB}\wedge\hat{e}^C\wedge\hat{e}^D
-    + \alpha_2\hat{R}^{AB}\wedge\hat{R}^{CD}\Big)=0,
+    \hat{\epsilon}_{ABCDE}\Big(\alpha_0\hat{e}^B\wedge\hat{e}^C\wedge\hat{e}^D\wedge\hat{e}^E
+    + \alpha_1\hat{R}^{BC}\wedge\hat{e}^D\wedge\hat{e}^E
+    + \alpha_2\hat{R}^{BC}\wedge\hat{R}^{DE}\Big)=0,
     \\
@@
 The last term in Eq.~\eqref{action5EGB} is the five-dimensional Gauss--Bonnet term for a Lorentz curvature written in exterior forms, which is analogue to Eq.~\eqref{GB} for the Riemannian case.
+Exact solutions with torsion where reported in~\cite{Canfora:2007ux}.
 
@@
 This condition place us outside the Chern--Simons point.
-Otherwise, the field equations~\eqref{delta e} and~\eqref{delta w} are invariant under a larger gauge group ($AdS_5$), while the action~\eqref{action5EGB} becomes the Chern--Simons form for that group~\cite{Zanelli:2005sa,*Troncoso:1999pk}.
+Otherwise, the field equations~\eqref{delta e} and~\eqref{delta w} are invariant under a larger gauge group ($AdS_5$), while the action~\eqref{action5EGB} becomes the Chern--Simons form for that group~\cite{Zanelli:2016cs,*Troncoso:1999pk}.
+Exact solutions with torsion in Chern--Simons gravity are given in~\cite{Aros:2007nn,Banados:2003cz}.
 
@@
 Appendix~\ref{homotropic} is devoted to the details of how to find the general ansatz.
-
-The four-dimensional vielbein compatible with the Friedmann--Robertson--Walker metric is determined only by one function, the scale factor $a(t)$, and their components read
+The metric for this case is given by a Friedmann--Robertson--Walker ansatz whose line element reads
+\begin{align*}
+\text{d}s^2=-\text{d}t^2+a^2(t)\left(\frac{\text{d}r^2}{1-kr^2}+r^2\text{d}\theta^2+r^2\sin^2\varphi\text{d}\varphi^2\right),
+\end{align*}
+where $a(t)$ is the scale factor and $k=+1,0,-1$ determines the spatial section of the four-manifold to be closed, flat or open, respectively.
+The four-dimensional vielbein compatible with this metric reads
 \begin{equation}
@@
     e^0&=\mbox{d}t, & e^1&=\frac{a(t)}{\sqrt{1-kr^2}} \, \mbox{d}r,\\
-    e^2&=a(t)r \, \mbox{d}\theta, & e^3&=a(t) r \sin(\varphi) \,\mbox{d}\varphi, 
+    e^2&=a(t)r \, \mbox{d}\theta, & e^3&=a(t) r \sin\varphi \,\mbox{d}\varphi.
   \end{aligned}
 \end{equation}
-where $k=+1,0,-1$ determines the extrinsic curvature of the four-dimensional manifold to be closed, flat or open, respectively.
-
 The scalar field is a time-dependent function, 
@@
 
-The final set of equations is the following:
+After a simple algebra, the system becomes
 \setlength\multlinegap{0pt}
@@
 
-\subsection{Solutions}
+\subsection{Solutions\label{sec:sol}}
 
-The system is integrated and develops two non-Riemannian branches, one for each value of the parameter $u_\pm$ defined to be
+The system develops two non-Riemannian branches, one for each value of the parameter
 \begin{equation}
@@
 The Eqs.~\eqref{eqn1} to~\eqref{eqn2} are reduced to the Riemannian system when \mbox{$f=h=0$} and \mbox{$\gamma=-\dot{\phi}/2\phi$}.
-The details of such a model can be seen in Refs.~\cite{Deruelle:1986iv,Deruelle:2003ck,Henriques:1986jw,*Ishihara:1986if,Kleidis:1997mu}.
+The details of such a model can be seen in Refs.~\cite{Deruelle:1986iv,Deruelle:2003ck,Henriques:1986jw,*Ishihara:1986if,Kleidis:1997mu,Canfora:2016umq}.
 
@@
   \gamma(\omega)&=\frac{u_\pm}{\omega},\\
-  \phi(\omega)&=\frac{\phi_0\,\omega^2}{\omega^2+u_\pm}\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2\left(\omega^2+u_\pm\right)}\right],\\
-  a(\omega)&=\frac{a_0}{\sqrt{\omega^2+u_\pm}}\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2\left(\omega^2+u_\pm\right)}\right],\\
-  f^2(\omega)&=\left(\omega^2+u_\pm\right)\frac{k}{a_0^2}\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2\left(\omega^2+u_\pm\right)}\right]\notag\\
+  \phi(\omega)&=\frac{\phi_0\,\omega^2}{\left|\omega^2+u_\pm\right|}\exp\left[\frac{\mp\sqrt{6\Delta}}{4\alpha_2\left(\omega^2+u_\pm\right)}\right],\\
+  a(\omega)&=\frac{a_0}{\sqrt{\left|\omega^2+u_\pm\right|}}\exp\left[\frac{\pm\sqrt{6\Delta}} {24\alpha_2\left(\omega^2+u_\pm\right)}\right],\\
+  \label{f2sol}
+  f^2(\omega)&=\left|\omega^2+u_\pm\right|\frac{k}{a_0^2}\exp\left[\frac{\mp\sqrt{6\Delta}}{12\alpha_2\left(\omega^2+u_\pm\right)}\right]\notag\\
   & \quad +\omega^2  + \frac{3\alpha_1\pm\sqrt{6\Delta}}{6\alpha_2},
@@
   \includegraphics[width=\linewidth]{cond_u_pos.pdf}
-  \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be positive defined.}
+  \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be positive definite.}
   \label{cond_u_pos}
@@
   \includegraphics[width=.95\linewidth]{cond_u_neg.pdf}
-  \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be positive defined.}
+  \caption{Allowed regions, in the $\alpha_0$--$\alpha_2$ plane, for $u_+$ and $u_-$ to be negative definite.}
   \label{cond_u_neg}  
@@
 As shown in Fig.~\ref{a_u_0}, for positive $\alpha_2$ the solutions fit into the third category above, while for negative $\alpha_2$ the solutions expand eternally.
-%Notice that in the former case the Universe's fate is an asymptotic collapse, whilst in the latter expands forever.
+%Notice that in the former case the universe's fate is an asymptotic collapse, whilst in the latter expands forever.
 \begin{figure}[H]
@@
 
-\paragraph{The case $u_\pm < 0$} has solutions which either expand or collapse eternally, as shown in Fig.~\ref{a_u_neg}.
-The typical evolution of the Universe with $u_+ < 0$ is to grow infinitely, while for negative $u_-$ the expansion (contraction) corresponds to $\alpha_2$ positive (negative).
-Figure~\ref{phi_u_neg} shows the behaviour of $\phi$ for this case.%, which presents an asymptotic contraction (expansion) when the scale factor expands (contracts) eternally.
+\paragraph{The case $u_\pm < 0$} has solutions which either expand or contract eternally, as shown in Fig.~\ref{a_u_neg}.
+The typical evolution of the universe with $u_+ < 0$ is to grow infinitely, while for negative $u_-$ the expansion (contraction) corresponds to $\alpha_2$ positive (negative).
+Figure~\ref{phi_u_neg} shows the behavior of $\phi$ for this case.
+%, which presents an asymptotic contraction (expansion) when the scale factor expands (contracts) eternally.
 %% , depending on the sign
@@
 For all these solutions the scale factor remains finite at $t = 0$, presenting no initial singularity.
-It has been reported that the Gauss--Bonnet term can prevent the Universe to expand from an initial singularity~\cite{Deruelle:1986iv,Henriques:1986jw,*Ishihara:1986if}, which is also this case.
+It has been reported that the Gauss--Bonnet term can prevent the universe to expand from an initial singularity~\cite{Deruelle:1986iv,Henriques:1986jw,*Ishihara:1986if}, which is also this case.
 
@@
 
-The Eq.~\eqref{equation} can be written into its familiar form, by means of the identity~\eqref{curvature decomp}, as
+The Eq.~\eqref{equation} can be written in a familiar form, by means of~\eqref{curvature decomp}, as
 \begin{align}\label{einstein equation}
@@
 \end{align}
-with $\Lambda \equiv -6\alpha_0/\alpha_1$, $\kappa_{G} \equiv 2/\alpha_1$, and the effective energy-momentum $3$-form defined as
+with $\Lambda \equiv -6\alpha_0/\alpha_1$, $\kappa_{G} \equiv 2/\alpha_1$ and 
 \begin{align}
@@
 where $\kappa^{ab}$ is the contorsion $1$-form defined as the torsional correction to the spin connection (see Eq.~\eqref{spin separation}).
-The contributions of $F$, $\alpha^{ab}$ and $\beta^a$ have not been taken into account since, for all practical purposes of this article, they vanish.
-The left hand side of the Eq.~\eqref{einstein equation} is the $3$-form whose Hodge dual yields to the Einstein equations with cosmological constant in four dimensions.
+The contributions of $F$, $\alpha^{ab}$ and $\beta^a$ have not been taken into account since, for the purposes of this article, they vanish.
+The left hand side of the Eq.~\eqref{einstein equation} is the $3$-form whose Hodge dual yields the Einstein equations with cosmological constant in four dimensions.
 %%The left hand side of Eq.~\eqref{einstein equation} is the 3-form which, in components, yields the four-dimensional Einstein equations with cosmological constant.
-We identify the right hand side as a $3$-form effective energy-momentum, which behaves as a perfect fluid. It contains the torsional and the higher dimensional degrees of freedom.
+We identify the right hand side as an energy-momentum $3$-form induced by the geometry, which behaves as a perfect fluid. It contains the torsional and the higher dimensional degrees of freedom.
 
@@
 From the Eqs.~\eqref{our-rho} and~\eqref{our-p} one can see that the induced energy density and pressure are not positive definite quantities.
-In fact, the Universe undergoes through an accelerating expansion phase due to the presence of torsion and the extra dimensional fields.
+In fact, the universe undergoes through an accelerating expansion phase due to the presence of torsion and the extra dimensional fields.
 
@@
 
-We used the most general ansatz compatible with the cosmological symmetries (see Appendix~\ref{homotropic}) to find solutions of the Friedmann--Robertson--Walker class in the effective theory.
-The equations of motion for the cosmological ansatz ensure that the solutions with nontrivial torsion have vanishing $\alpha^{ab}$ and $\beta^a$.
-
+We used the most general ansatz compatible with the cosmological symmetries (see Appendix~\ref{homotropic}) to find solutions of the Friedmann--Robertson--Walker class in the four-dimensional theory.
+The field equations for the cosmological ansatz ensure that the solutions with nontrivial torsion have vanishing $\alpha^{ab}$ and $\beta^a$.
 The solutions are parametrized by $u_\pm$, related to the fundamental couplings of the theory through the Eq.~\eqref{u}, and exhibit three different sectors depending on whether this parameter is positive, negative or zero.
-
-The behavior of the Universes described by our solutions is four-fold: (i)~eternal expansion, (ii)~eternal contraction, (iii)~initial expansion with asymptotic contraction, and (iv)~bouncing.
+The behavior of the universes described by our solutions is four-fold: (i)~eternal expansion, (ii)~eternal contraction, (iii)~initial expansion with asymptotic contraction, and (iv)~bouncing.
 %We show the typical evolution of the scale factor for each case in Figs.~\ref{a_u_pos}, ~\ref{a_u_0} and~\ref{a_u_neg}.
+In some cases, the solutions do not start from an initial singularity.
 The size of the extra dimension is driven by the scalar field $\phi(t)$ and its behavior at late times is compatible with a dynamical mechanism for compactification, at certain regions of the parameter space (see Figs.~\ref{phi_u_0} and~\ref{phi_u_neg}).
@@
 The field equation~\eqref{equation} admits the interpretation of the Einstein equations, where the energy-momentum $3$-form is induced by the torsion and the extra dimensional degrees of freedom.
-It is important to mention that the induced energy-momentum form is nonstandard, in the sense that it inherit a coupling between matter and curvature---see Eq.~\eqref{taua} for the definition---, due to the nonminimal coupling in the reduced theory.
-Moreover, the energy density and the pressure are not positive definite, which allows us to interpret these induced fields as exotic, providing a playground for a gravitational explanation for the accelerated expansion of the Universe.
+It is important to mention that the induced energy-momentum form is nonstandard, in the sense that it inherit a coupling between matter and curvature (see Eq.~\eqref{taua} for the definition).
+Moreover, the energy density and the pressure are not positive definite, %which allows us to interpret these induced fields as exotic,
+providing a playground for a gravitational explanation for the accelerated expansion of the universe.
 
@@
 
-There are periods of time where $f^2(t)$ becomes negative.
-As we can see from Eq.~\eqref{homotorsion}, $f$ corresponds to the completely antisymmetric part of torsion and is unseen by classical particles following geodesics.
-It does not couple minimally to spin-0 or spin-1 bosons, but it does appear as an effective interaction term in the Dirac Lagrangian written in a curved spacetime~\cite{Hehl:1976kj}
+As we can see from Eq.~\eqref{homotorsion}, $f$ corresponds to the completely antisymmetric part of torsion.
+Its solution given in Eq.~\eqref{f2sol} shows that it can be imaginary for certain periods of time.
+Additionally, it is the only function sensitive to the value of $k$.
+Thus, if a reality condition is imposed over $f$ for certain values of the coupling constants, it would provide a distinction criteria among open, flat and closed universes.
+In the light of this and given the experimental evidence supporting the flatness of the universe (see Ref.~\cite{Spergel:2006hy}) is that we restrict the discussion to the case where $k=0$.~\footnote{The cases where $k = \pm1$ can be studied analogously.} A reality condition over $f$ in Eq.~\eqref{f2sol} is satisfied within the region
 \begin{align*}
-  \mathcal{L}_{\text{Int}}&=-\frac{i}{8}T_{\alpha\mu\nu}\bar{\Psi}\Gamma^{\alpha\mu\nu}\Psi =\frac{3}{2}f\bar{\Psi}\Gamma^0\Gamma^5\Psi.
+ & & 3\alpha_1 \pm \sqrt{6\Delta} &\geq 0 & &\mbox{and}& \alpha_2 &>0, \\
+ & \mbox{or,} & 3\alpha_1 \pm \sqrt{6\Delta} &\leq 0 & &\mbox{and}& \alpha_2&<0.
 \end{align*}
-It couples to the chiral current.
-For imaginary values of $f(t)$, the corresponding violation of current conservation can be interpreted as particle creation.
-Similar results where found in~\cite{Toloza:2013wi} for other cosmological model with torsion which shares some of the couplings of a reduced action.
+This should be taken into account to restrict the parameter region studied in Sec.~\ref{sec:sol}.  
+A similar behavior for this function was found in Ref.~\cite{Toloza:2013wi}.
+Notice that the antisymmetric part of torsion is unseen by classical particles following geodesics and it does not couple minimally to spin-0 or spin-1 bosons.
+However, it appears as an effective interaction term for the Dirac Lagrangian in spacetimes with torsion~\cite{Hehl:1976kj}
+\begin{align*}
+  \mathcal{L}_{\text{Int}}&=-\frac{i}{8}T_{\alpha\mu\nu}\bar{\Psi}\Gamma^{\alpha\mu\nu}\Psi =\frac{3}{2}f\bar{\Psi}\Gamma^0\Gamma^5\Psi,
+\end{align*}
+which couples to the fermionic axial current.
+For imaginary values of $f(t)$, the authors in Ref.~\cite{Toloza:2013wi} argued that the loss of unitarity and the violation of the current conservation can be interpreted as particle creation.
+
+The vacuum cosmological solutions presented here provide a useful arena to isolate the effects of the induced fields and probe their consequences in physically important scenarios.
+In that sense, and since they are absent in the cosmological ansatz, a four-dimensional isotropic configuration might contribute to interpret the $\alpha^{ab}$ and $\beta^a$ fields as gravitational hairs in black hole solutions, or wormhole supporting matter.
+
+The gravitational principles we take for granted can be relaxed to construct a logically consistent theory.
+This approach might provide an explanation for the phenomena at the limit of the current understanding of the universe, such as its content and evolution.
+This work shows that the renounce to the metric description of gravity, attached to a higher dimensional spacetime, can radically affect the way we perceive the gravitational degrees of freedom and, therefore, what our measurements read as the energy content of the universe.
+
+%It is clear that the principles we take for granted to build a gravity action can be relaxed to construct a logically consistent theory.
+To renounce to certain standard assumtions such as the metric description of gravity, can as well provide explanation for phenomena at the limit of the current understanding of the universe, such as its content and evolution.
+This work shows that this idea, attached to a higher dimensional spacetime can radically affect the way we perceive the gravitational degrees of freedom, and therefore, what our measurements read as the energy content of the universe.
+
+%The evidence of dark matter/energy in the universe,
+
+%The relaxation of the torsion-free condition allows to address the cosmological behaviour of the reduced four-dimensional theory.
+
+%Despite this, the regime where  provides
 
 \begin{acknowledgments}
-  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, R.~Gannouji, A.~Toloza and J.~Zanelli for enlightening discussions on this work.
-\mbox{O.~C.-F.} has been partially supported by the PAI project No.~79140040 (CONICYT--Chile).
+  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, S.~Kovalenko, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, R.~Gannouji, A.~Toloza, A.~Cisterna and J.~Zanelli for valuable comments on this work.
+\mbox{O.~C.-F.} has been partially supported by the PAI project No.~79140040 (CONICYT--Chile) and F.~R.
+by the Fondecyt project No.~1160423.
 The Centro Cient\'ifico Tecnol\'ogico de Valpara\'iso (CCTVal) is funded by the Chilean Government through the Centers of Excellence Basal Financing Program of CONICYT.
@@
 \end{equation}
-On the other hand, the curvature $2$-form also suffers corrections with respect to the Riemannian one due to the presence of torsional degrees of freedom.
+On the other hand, the curvature $2$-form also suffers corrections with respect to the Riemannian one, due to the presence of torsional degrees of freedom.
 This can be seen explictly by taking the definition of curvature in Eq.~\eqref{curvadef} and~\eqref{spin separation} to find
@@
 This means that the fields involved in the model must be compatible with this assumption.
-A spacetime is isotropic with respect to certain point $P$ if after a rotation with respect to an axis passing through $P$, all the geometrical properties remain invariant, thus the spacetime looks the same in all directions.
-Homogeneity is understood such that spacetime looks the same from every point $P$.
+A spacetime is isotropic with respect to certain point $P$ if after a rotation with respect to an axis passing through $P$, all the geometrical properties remain invariant.
+Thus, the spacetime looks the same in all directions.
+Homogeneity is understood such that the spacetime looks the same from every point $P$.
 These two assumptions are translated in the Killing equations, which are the vanishing of the Lie derivatives of the fields along the vectors which generate the symmetries $\{\zeta^\lambda_{i}\}$.
@@
   \label{killing metric}
-  \text{\textsterling}_i g_{\mu\nu}&=\zeta^\lambda_i\partial_\lambda g_{\mu\nu}+\partial_\mu\zeta^\lambda_i g_{\lambda\nu}+\partial_\nu\zeta^\lambda_i g_{\mu\lambda}=0,
+  \text{\textsterling}_i g_{\mu\nu}&=\zeta^\lambda_i\partial_\lambda g_{\mu\nu} + g_{\lambda\nu}\partial_\mu\zeta^\lambda_i  + g_{\mu\lambda}\partial_\nu\zeta^\lambda_i = 0,
   \\
-  \text{\textsterling}_i T^\alpha_{\ \mu\nu}&=\zeta^\lambda_i\partial_\lambda T^\alpha_{\ \mu\nu}-\zeta^\alpha_i\partial_\lambda T^\lambda_{\ \mu\nu}
-  +\zeta^\lambda_i\partial_\mu T^\alpha_{\ \lambda\nu} \notag \\
-  &\quad +\zeta^\lambda_i\partial_\nu T^\alpha_{\ \mu\lambda}=0,
+  \text{\textsterling}_i T^\alpha_{\ \mu\nu}&=\zeta^\lambda_i\partial_\lambda T^\alpha_{\ \mu\nu} - T^\lambda_{\ \mu\nu} \partial_\lambda \zeta^\alpha_i + T^\alpha_{\ \lambda\nu} \partial_\mu\zeta^\lambda_i  \notag \\
+  &\quad + T^\alpha_{\ \mu\lambda} \partial_\nu\zeta^\lambda_i = 0,
 \end{align}
@@
   \label{killing vector}
-  \text{\textsterling}_i A_\mu&=\zeta^\lambda_i\partial_\lambda A_\mu+\zeta^\lambda_i\partial_\mu A_\lambda=0,\\
+  \text{\textsterling}_i A_\mu&=\zeta^\lambda_i\partial_\lambda A_\mu + A_\lambda \partial_\mu\zeta^\lambda_i = 0,\\
   \text{\textsterling}_i \phi &=\zeta^\lambda_i\partial_\lambda\phi=0,

### 2016-10-06 KKosmology.tex
After the comments by R. Olea

@@
 In four dimensions, the Gauss--Bonnet term adds no dynamics to the metric, since it represents a topological invariant proportional to the Euler characteristic class, which can be written locally as a boundary term.
-Nevertheless, in dimensions higher than four it contributes to the field equations.
-This Lagrangian was also identified as the low-energy correction for a spin two field in string theory~\cite{Zwiebach:1985uq}.
+Nevertheless, it has been shown that this term can be relevant for conserved charges considerations in spacetimes with local AdS asymptotic~\cite{Aros:1999id}.
+Moreover, as it was reported in Ref.~\cite{Miskovic:2009bm}, the inclusion of topological invariants of the Euler class is equivalent to the program of Holographic Renormalization in the context of AdS/CFT.
+In dimensions higher than four it contributes to the field equations and it was also identified as the low-energy correction for a spin two field in string theory~\cite{Zwiebach:1985uq}.
 
@@
 Exact solutions with torsion in Chern--Simons gravity are given in~\cite{Aros:2007nn,Banados:2003cz}.
+Its holographic properties in the Riemannian case where studied in Ref.~\cite{Banados:2005rz} and in \cite{Banados:2006fe} when torsion is considered.
 
@@
 \begin{acknowledgments}
-  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, S.~Kovalenko, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, R.~Gannouji, A.~Toloza, A.~Cisterna and J.~Zanelli for valuable comments on this work.
+  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, S.~Kovalenko, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, R.~Olea, R.~Gannouji, A.~Toloza, A.~Cisterna and J.~Zanelli for valuable comments on this work.
 \mbox{O.~C.-F.} has been partially supported by the PAI project No.~79140040 (CONICYT--Chile) and F.~R.

### 2016-11-14 KKosmology.tex
First changes after the Referee report!

@@
 
-In four dimensions, the Einstein--Hilbert action with cosmological constant is the most general theory which leads to second order field equations for the metric.
-In higher dimensions, however, particular combinations of higher order terms in the curvature can be added to the gravitational action, whose variation with respect to the metric yields also second order field equations.
-The most general theory in arbitrary dimensions, which preserves this feature of the four-dimensional Einstein--Hilbert, is named the Lanczos--Lovelock action~\cite{Lanczos:1938sf,*Lovelock:1971yv}.
-Such a theory, has no ghosts~\cite{Zumino:1985dp} and it has the same degrees of freedom as the Einstein--Hilbert Lagrangian in arbitrary dimensions~\cite{Henneaux:1990au}.
+In four dimensions, the Einstein--Hilbert action with cosmological
+constant is the most general theory which leads to second order field
+equations for the metric. In higher dimensions, however, particular
+combinations of higher order terms in the curvature can be added to
+the gravitational action, whose variation with respect to the metric
+yields also second order field equations. The most general theory in
+arbitrary dimensions, which preserves this feature of the
+four-dimensional Einstein--Hilbert, is named the Lanczos--Lovelock
+action~\cite{Lanczos:1938sf,*Lovelock:1971yv}. Such a theory, has no
+ghosts~\cite{Zumino:1985dp} and it has the same degrees of freedom as
+the Einstein--Hilbert Lagrangian in arbitrary
+dimensions~\cite{Henneaux:1990au}. It is worth mentioning that in the
+Palatini approach, where the metric and the connection are considered
+as independent fields, there are families of Lagrangians which yield
+to second order field equations, and do not possess
+ghosts~\cite{Olmo:2011uz}.
 
@@
 \end{equation}
-where $\tilde{R}_{\alpha\beta\mu\nu}$ is the Riemannian curvature of a manifold with metric $g_{\mu\nu}$ and $g$ its determinant.
-$\tilde{R}_{\mu\nu}$ and $\tilde{R}$ are the Ricci tensor and Ricci scalar, respectively.
-In four dimensions, the Gauss--Bonnet term adds no dynamics to the metric, since it represents a topological invariant proportional to the Euler characteristic class, which can be written locally as a boundary term.
-Nevertheless, it has been shown that this term can be relevant for conserved charges considerations in spacetimes with local AdS asymptotic~\cite{Aros:1999id}.
-Moreover, as it was reported in Ref.~\cite{Miskovic:2009bm}, the inclusion of topological invariants of the Euler class is equivalent to the program of Holographic Renormalization in the context of AdS/CFT.
-In dimensions higher than four it contributes to the field equations and it was also identified as the low-energy correction for a spin two field in string theory~\cite{Zwiebach:1985uq}.
+where $\tilde{R}_{\alpha\beta\mu\nu}$ is the Riemannian curvature of a
+manifold with metric $g_{\mu\nu}$ and $g$ its
+determinant. $\tilde{R}_{\mu\nu}$ and $\tilde{R}$ are the Ricci tensor
+and Ricci scalar, respectively. In four dimensions, the Gauss--Bonnet
+term adds no dynamics to the metric, since it represents a topological
+invariant proportional to the Euler characteristic class, which can be
+written locally as a boundary term. Nevertheless, it has been shown
+that this term can be relevant for conserved charges considerations in
+spacetimes with local AdS asymptotic~\cite{Aros:1999id}. Moreover, as
+it was reported in Ref.~\cite{Miskovic:2009bm}, the inclusion of
+topological invariants of the Euler class is equivalent to the program
+of Holographic Renormalization in the context of AdS/CFT. In
+dimensions higher than four it contributes to the field equations and
+it was also identified as the low-energy correction for a spin two
+field in string theory~\cite{Zwiebach:1985uq}.
 
-The Riemannian five-dimensional Lanczos--Lovelock theory has been widely studied in the literature.
-For instance, exact wormhole solutions have been found in vacuum~\cite{Dotti:2006cp,*Dotti:2007az} and coupled with matter fields~\cite{Mehdizadeh:2015jra}.
-In higher order Lanczos--Lovelock models, this class of exact solutions have been reported in~\cite{Mehdizadeh:2015dta} and in the compactified theory with torsion~\cite{Canfora:2008ka}.
-Additionally, the compactification of higher Lanczos--Lovelock terms has been considered in Refs.~\cite{MuellerHoissen:1985mm,*MuellerHoissen:1989yv} and their cosmology in Refs.~\cite{MuellerHoissen:1985ij,Deruelle:1986iv,Deruelle:1989fj}.
-It is worth mentioning that, classically, wormholes must be supported by a kind of energy compatible with the cosmological hypothesis.
-Their existence in these extended models implies the presence of degrees of freedom that could provide an explanation for the exotic matter/energy abundance in the universe.
+The Riemannian five-dimensional Lanczos--Lovelock theory has been
+widely studied in the literature. For instance, exact wormhole
+solutions which violate no energy conditions have been found in
+vacuum~\cite{Dotti:2006cp,*Dotti:2007az}, and coupled with matter
+fields satisfying the weak energy
+conditions~\cite{Mehdizadeh:2015jra}. In higher order
+Lanczos--Lovelock models, this class of exact solutions have been
+reported in~\cite{Mehdizadeh:2015dta} and in the compactified theory
+with torsion~\cite{Canfora:2008ka}. Additionally, the compactification
+of higher Lanczos--Lovelock terms has been considered in
+Refs.~\cite{MuellerHoissen:1985mm,*MuellerHoissen:1989yv} and their
+cosmology in
+Refs.~\cite{MuellerHoissen:1985ij,Deruelle:1986iv,Deruelle:1989fj}. It
+is worth mentioning that classically, in GR, wormholes must be supported by
+a kind of energy compatible with the cosmological hypothesis. Their
+existence in these extended models implies the presence of degrees of
+freedom that could provide an explanation for the exotic matter/energy
+abundance in the universe.
 
-On the other hand, it is well-known that GR assume the torsion-free condition {\it a priori}.
-However, as E.~Cartan first considered, it is possible to take metricity and parallelism as truly independent concepts~\cite{Cartan1922,*Cartan1924,*Cartan1925}.
-Such kind of geometries are known as Riemann--Cartan and offer the natural framework for Poincar\'e gauge theories, where the torsion appears as the field strength of translations, sourced by the spin-current~\cite{Kibble:1961ba,Sciama:1962,Hehl:1976kj}.
+On the other hand, it is well-known that GR assume the torsion-free
+condition {\it a priori}. However, as E.~Cartan first considered, it
+is possible to take metricity and parallelism as truly independent
+concepts~\cite{Cartan1922,*Cartan1924,*Cartan1925}. Such kind of
+geometries are known as Riemann--Cartan and offer the natural
+framework for Poincar\'e gauge theories, where the torsion appears as
+the field strength of translations, sourced by the
+spin-current~\cite{Kibble:1961ba,Sciama:1962,Hehl:1976kj}.
 %In such a framework, the vielbein is interpreted as the gauge connection of translations and the torsion as its field strength.
@@
 %Due to this, several authors have relaxed the torsion-free condition since, as E. Cartan argued, there is no a fundamental reason to assume a torsion-free condition {\it a priori}. 
-Moreover, the vacuum predictions of its simplest formulation---the Einstein--Cartan theory---holds the experimental tests of GR.
-The cosmological consequences of non-Riemannian geometries has been widely studied in the literature (for a review, see Ref.~\cite{Puetzfeld:2004yg}).
+Moreover, the vacuum predictions of its simplest formulation---the
+Einstein--Cartan theory---holds the experimental tests of GR. In the
+same spirit of the Gauss--Bonnet extension of GR, the quadratic
+corrections in curvature and torsion to Einstein--Cartan have been
+considered, see for example Refs.~\cite{Baekler:2011jt,Fabbri:2012qr}.
+The cosmological consequences of non-Riemannian geometries has been
+widely studied in the literature (for a review, see
+Ref.~\cite{Puetzfeld:2004yg}).
 %For instance, the pioneer works of Kopczy\'nski showed how the introduction of a spinning dust, acting as a source of the curvature and torsion, avoids the appearance of cosmological singularities~\cite{Kopczynski:1972,Kopczynski:1973}.
@@
 
-In the framework of the Kaluza--Klein theories in higher dimensional Riemann--Cartan geometries, phenomenology of the extra dimensional torsion was found in Ref.~\cite{Kalinowski:1980da}, metric dependent torsion in extra dimensions was studied in~\cite{Shankar:2012vd}, along with its consequences in cosmology~\cite{Chen:2009ep}.
-Compatification of higher dimensional Brans--Dicke models with torsion was considered in Ref.~\cite{German:1993bq}.
-Torsion-free black-hole solutions were found in~\cite{Aros:2007nn}, for first order compactified gravity.
+In the framework of the Kaluza--Klein theories in higher dimensional
+Riemann--Cartan geometries, phenomenology of the extra dimensional
+torsion was found in Ref.~\cite{Kalinowski:1980da}, metric dependent
+torsion in extra dimensions was studied in~\cite{Shankar:2012vd},
+along with its consequences in
+cosmology~\cite{Chen:2009ep}. Compatification of higher dimensional
+Brans--Dicke models with torsion was considered in
+Ref.~\cite{German:1993bq}. Torsion-free black-hole solutions were
+found in~\cite{Aros:2007nn}, for first order compactified gravity.
 
-The extension of the Lanczos--Lovelock theories with nonvanishing torsion is known as Lovelock--Cartan~\cite{Mardones:1990qc}.
-In that framework, we present a new class of cosmological solutions in five dimensions, where the compact dimension is $S^1$.
-The theory admits a nonvanishing torsion in vacuum due to the presence of the Gauss--Bonnet term, in contrast to the five-dimensional Einstein--Cartan theory.
-The new degrees of freedom coming from the higher dimensional geometry, generate an induced energy-momentum form in the reduced theory.
-In some cases, the solutions avoid the appearance of initial singularities and drive the accelerated expansion of the universe, while the radius of the compact manifold goes to zero.
+The extension of the Lanczos--Lovelock theories with nonvanishing
+torsion is known as Lovelock--Cartan~\cite{Mardones:1990qc}. In that
+framework, we present a new class of cosmological solutions in five
+dimensions, where the compact dimension is $S^1$. The theory admits a
+nonvanishing torsion in vacuum due to the presence of the
+Gauss--Bonnet term, in contrast to the five-dimensional
+Einstein--Cartan theory. The new degrees of freedom coming from the
+higher dimensional geometry, generate an induced energy-momentum form
+in the reduced theory. In some cases, the solutions avoid the
+appearance of initial singularities and drive the accelerated
+expansion of the universe, while the radius of the compact manifold
+goes to zero.
 
@@
 \begin{acknowledgments}
-  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt, S.~Kovalenko, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, R.~Olea, R.~Gannouji, A.~Toloza, A.~Cisterna and J.~Zanelli for valuable comments on this work.
-\mbox{O.~C.-F.} has been partially supported by the PAI project No.~79140040 (CONICYT--Chile) and F.~R.
-by the Fondecyt project No.~1160423.
-The Centro Cient\'ifico Tecnol\'ogico de Valpara\'iso (CCTVal) is funded by the Chilean Government through the Centers of Excellence Basal Financing Program of CONICYT.
+  The authors would like to thank A.~R.~Zerwekh, I.~Schmidt,
+  S.~Kovalenko, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, R.~Olea,
+  R.~Gannouji, A.~Toloza, A.~Cisterna and J.~Zanelli for valuable
+  comments on this work. \mbox{O.~C.-F.} has been partially supported
+  by the PAI project No.~79140040 (CONICYT--Chile) and F.~R. by the
+  Fondecyt project No.~1160423. The Centro Cient\'ifico Tecnol\'ogico
+  de Valpara\'iso (CCTVal) is funded by the Chilean Government through
+  the Centers of Excellence Basal Financing Program FS0821 of CONICYT.
 \end{acknowledgments}
@@
 %%%%%%%%% BIBLIOGRAPHY %%%%%%%%%
-%%\bibliographystyle{apsrev4-1}
-\bibliography{References.bib}
+%% \bibliographystyle{apsrev4-1}
+\bibliography{References}
 

### 2016-11-14 KKosmology.tex
Second correction after the referee report.

@@
 where $R=\text{d}\omega+\omega\wedge\omega$ is the curvature of the reduced spacetime $M_{N-1}$.
-Similarly from the definition of torsion in Eq.~\eqref{tordef}, we find its distinctive parts to be~\footnote{Henceforth we will refer to as distinctive parts regarding the $M_{N-1}\times S^1$ decomposition.}
+Similarly from the definition of torsion in Eq.~\eqref{tordef}, we find its distinctive parts to be\footnote{Henceforth we will refer to as distinctive parts regarding the $M_{N-1}\times S^1$ decomposition.}
 \begin{align}\label{T1}
@@
 \end{equation}
-This is taken as a proof of the tensorial nature of these new fields under the four-dimensional Lorentz transformations.~\footnote{The tensorial nature of these fields can be also derived from the decomposition of the transformation rule of $\hat{\omega}^{AB}$ under the Lorentz group.} The second Bianchi identity gives its equivalent for $M_{N-1}$, this is
+This is taken as a proof of the tensorial nature of these new fields under the four-dimensional Lorentz transformations.\footnote{The tensorial nature of these fields can be also derived from the decomposition of the transformation rule of $\hat{\omega}^{AB}$ under the Lorentz group.} The second Bianchi identity gives its equivalent for $M_{N-1}$, this is
 \begin{equation*}
@@
   \label{delta}
-  \Delta\equiv\alpha_1^2-4\alpha_0\alpha_2\neq 0.
+  \Delta\equiv\alpha_1^2-4\alpha_0\alpha_2 > 0.
 \end{equation}
-This condition place us outside the Chern--Simons point.
-Otherwise, the field equations~\eqref{delta e} and~\eqref{delta w} are invariant under a larger gauge group ($AdS_5$), while the action~\eqref{action5EGB} becomes the Chern--Simons form for that group~\cite{Zanelli:2016cs,*Troncoso:1999pk}.
-Exact solutions with torsion in Chern--Simons gravity are given in~\cite{Aros:2007nn,Banados:2003cz}.
-Its holographic properties in the Riemannian case where studied in Ref.~\cite{Banados:2005rz} and in \cite{Banados:2006fe} when torsion is considered.
+This condition place us outside the Chern--Simons point ($\Delta =
+0$). Otherwise, the field equations~\eqref{delta e} and~\eqref{delta
+  w} are invariant under a larger gauge group ($AdS_5$), while the
+action~\eqref{action5EGB} becomes the Chern--Simons form for that
+group~\cite{Zanelli:2016cs,*Troncoso:1999pk}. Exact solutions with
+torsion in Chern--Simons gravity are given
+in~\cite{Aros:2007nn,Banados:2003cz}. Holographic properties in the
+Riemannian case where studied in Ref.~\cite{Banados:2005rz} and in
+\cite{Banados:2006fe} when torsion is considered. Additionally, the
+case $\Delta < 0$ has been consider in Ref.~\cite{Canfora:2013xsa}, as
+well as its cosmology in Ref.~\cite{Canfora:2014iga}.
 
@@
 %None of the other cases, $u_- = 0$ and $u_\pm < 0$, present bouncing behavior.
-The allowed parameter space for $u_+$ and $u_-$ to be negative is shown in Fig.~\ref{cond_u_neg}, while the restriction $u_- = 0$ sets $\alpha_2 = \tfrac{1}{12 \alpha_0}$.~\footnote{Notice that in our choice of parameters $u_+$ cannot vanish.}
+The allowed parameter space for $u_+$ and $u_-$ to be negative is shown in Fig.~\ref{cond_u_neg}, while the restriction $u_- = 0$ sets $\alpha_2 = \tfrac{1}{12 \alpha_0}$.\footnote{Notice that in our choice of parameters $u_+$ cannot vanish.}
 
@@
 Thus, if a reality condition is imposed over $f$ for certain values of the coupling constants, it would provide a distinction criteria among open, flat and closed universes.
-In the light of this and given the experimental evidence supporting the flatness of the universe (see Ref.~\cite{Spergel:2006hy}) is that we restrict the discussion to the case where $k=0$.~\footnote{The cases where $k = \pm1$ can be studied analogously.} A reality condition over $f$ in Eq.~\eqref{f2sol} is satisfied within the region
+In the light of this and given the experimental evidence supporting the flatness of the universe (see Ref.~\cite{Spergel:2006hy}) is that we restrict the discussion to the case where $k=0$.\footnote{The cases where $k = \pm1$ can be studied analogously.} A reality condition over $f$ in Eq.~\eqref{f2sol} is satisfied within the region
 \begin{align*}
@@
   The authors would like to thank A.~R.~Zerwekh, I.~Schmidt,
-  S.~Kovalenko, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, R.~Olea,
-  R.~Gannouji, A.~Toloza, A.~Cisterna and J.~Zanelli for valuable
-  comments on this work. \mbox{O.~C.-F.} has been partially supported
-  by the PAI project No.~79140040 (CONICYT--Chile) and F.~R. by the
-  Fondecyt project No.~1160423. The Centro Cient\'ifico Tecnol\'ogico
-  de Valpara\'iso (CCTVal) is funded by the Chilean Government through
+  S.~Kovalenko, N.~A.~Neill, C.~O.~Dib, O.~Miskovic, R.~Gannouji,
+  A.~Toloza, A.~Cisterna and J.~Zanelli for valuable comments on this
+  work. \mbox{O.~C.-F.} has been partially supported by the PAI
+  project No.~79140040 (CONICYT--Chile) and F.~R. by the Fondecyt
+  project No.~1160423. The Centro Cient\'ifico Tecnol\'ogico de
+  Valpara\'iso (CCTVal) is funded by the Chilean Government through
   the Centers of Excellence Basal Financing Program FS0821 of CONICYT.
@@
 
-In Riemann--Cartan geometry, both the vielbein $e^A$ and the spin connection $\omega^{AB}$ are independent features of the manifold.~\footnote{Here, we will drop the hats since we will not deal with the KK decomposition.} The spin connection can be decomposed in a Riemannian part, which is metric dependent, $\tilde{\omega}^{AB}$ satisfying $\tilde{\mbox{D}}e^A=0$, and a contorsion piece $\kappa^{AB}=-\kappa^{BA}$, such that
+In Riemann--Cartan geometry, both the vielbein $e^A$ and the spin connection $\omega^{AB}$ are independent features of the manifold.\footnote{Here, we will drop the hats since we will not deal with the KK decomposition.} The spin connection can be decomposed in a Riemannian part, which is metric dependent, $\tilde{\omega}^{AB}$ satisfying $\tilde{\mbox{D}}e^A=0$, and a contorsion piece $\kappa^{AB}=-\kappa^{BA}$, such that
 \begin{equation}\label{spin separation}

### 2016-11-15 KKosmology.tex
Third (possible last) modification after referee report

@@
-\documentclass[aps,prd,12pt,superscriptaddress,showpacs,showkeys,longbibliography,reprint,nofootinbib]{revtex4-1}
+\documentclass[aps,prd,12pt,superscriptaddress,showpacs,showkeys,reprint,nofootinbib]{revtex4-1}
 \usepackage{calrsfs} 
@@
 \usepackage{breqn}
+\allowdisplaybreaks
 
@@
 
-In this work we have presented the dimensional reduction of the five-dimensional Lovelock--Cartan theory introduced in Ref.~\cite{Mardones:1990qc}, under the assumption that the compact dimension has $S^1$ topology.
-An interesting feature of these theories is that, unlike the Einstein--Cartan theory, torsion is allowed to propagate.
+In this work we have presented the dimensional reduction of the
+five-dimensional Lovelock--Cartan theory introduced in
+Ref.~\cite{Mardones:1990qc}, under the assumption that the compact
+dimension has $S^1$ topology. An interesting feature of these theories
+is that, unlike the Einstein--Cartan theory, torsion is allowed to
+propagate.
 
-In the generic reduction, the effective theory has a spin-2 particle, a spin-1 $U(1)$ gauge boson, and a spin-0 scalar particle coming from the decomposition of the metric, while the decomposition of the spin connection introduces three extra fields: two spin-1 particles, $\alpha^{ab}$ and $\gamma^a$, and the 1-form $\beta^a$ yields a new spin-2, a spin-1 and a spin-0 particles.
-These new fields do not transform under the gauge group (see Sec.~\ref{sec:bianchi}).
+Although there are several generalizations of gravitational models
+which include higher order terms in curvature and torsion---in either
+four or higher
+dimensions~\cite{Puetzfeld:2004yg,Baekler:2011jt,Fabbri:2012qr}---, we
+highlight that in our model the presence of higher order terms in
+curvature and torsion are kept minimal, in the sense that we introduce
+a sole extra coupling constant in our model (that accompanying the
+Gauss--Bonnet term in five dimensions).
+%% In the effective theory, we consider only the matter obtained from the
+%% compactification process.
+Moreover, the fact that our model is based in the Lovelock--Cartan
+gravity assures that the field equations are first order differential
+equations (since the Lagrangian can be written without the use of the
+Hodge dual), which is not the case in most gravitational models
+with higher order curvature and torsion.
 
-We used the most general ansatz compatible with the cosmological symmetries (see Appendix~\ref{homotropic}) to find solutions of the Friedmann--Robertson--Walker class in the four-dimensional theory.
-The field equations for the cosmological ansatz ensure that the solutions with nontrivial torsion have vanishing $\alpha^{ab}$ and $\beta^a$.
-The solutions are parametrized by $u_\pm$, related to the fundamental couplings of the theory through the Eq.~\eqref{u}, and exhibit three different sectors depending on whether this parameter is positive, negative or zero.
-The behavior of the universes described by our solutions is four-fold: (i)~eternal expansion, (ii)~eternal contraction, (iii)~initial expansion with asymptotic contraction, and (iv)~bouncing.
-%We show the typical evolution of the scale factor for each case in Figs.~\ref{a_u_pos}, ~\ref{a_u_0} and~\ref{a_u_neg}.
-In some cases, the solutions do not start from an initial singularity.
-The size of the extra dimension is driven by the scalar field $\phi(t)$ and its behavior at late times is compatible with a dynamical mechanism for compactification, at certain regions of the parameter space (see Figs.~\ref{phi_u_0} and~\ref{phi_u_neg}).
+In the generic reduction, the effective theory has a spin-2 particle,
+a spin-1 $U(1)$ gauge boson, and a spin-0 scalar particle coming from
+the decomposition of the metric, while the decomposition of the spin
+connection introduces three extra fields: two spin-1 particles,
+$\alpha^{ab}$ and $\gamma^a$, and the 1-form $\beta^a$ yields a new
+spin-2, a spin-1 and a spin-0 particles. These new fields do not
+transform under the gauge group (see Sec.~\ref{sec:bianchi}).
+
+We used the most general ansatz compatible with the cosmological
+symmetries (see Appendix~\ref{homotropic}) to find solutions of the
+Friedmann--Robertson--Walker class in the four-dimensional
+theory. Despite we restrict ourselves to the class of theories with
+$\Delta > 0$, in Ref.~\cite{Canfora:2014iga} the case of negative
+$\Delta$ is considered (without torsion). The field equations for the
+cosmological ansatz ensure that the solutions with nontrivial torsion
+have vanishing $\alpha^{ab}$ and $\beta^a$. The solutions are
+parametrized by $u_\pm$, related to the fundamental couplings of the
+theory through the Eq.~\eqref{u}, and exhibit three different sectors
+depending on whether this parameter is positive, negative or zero. The
+behavior of the universes described by our solutions is four-fold:
+(i)~eternal expansion, (ii)~eternal contraction, (iii)~initial
+expansion with asymptotic contraction, and (iv)~bouncing.
+%We show the typical evolution of the scale factor for each case in
+%Figs.~\ref{a_u_pos}, ~\ref{a_u_0} and~\ref{a_u_neg}.
+
+Among the different cases, there are some solutions which do not start
+from an initial singularity, however this is not the generic
+behavior. Moreover, all of our cosmological solutions are free of
+the future singularities categorized in Ref.~\cite{Shojai:2015dpa}. In
+addition, the size of the extra dimension is driven by the scalar
+field $\phi(t)$ and its behavior at late times is compatible with a
+dynamical mechanism for compactification, at certain regions of the
+parameter space (see Figs.~\ref{phi_u_0} and~\ref{phi_u_neg}).
 

### 2016-11-15 response.tex
Before the corrections by Simon

% Created 2016-11-15 Tue 14:14
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\date{}
\title{Response to the Referee\\\medskip
\large Article DX11551}
\hypersetup{
 pdfauthor={Oscar Castillo-Felisola},
 pdftitle={Response to the Referee},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 24.4.1 (Org mode 9.0)}, 
 pdflang={English}}
\begin{document}

\maketitle
Dear Referee,

First of all we want to thank your review, constructive comments and
suggestions.

Below we present a list of changes to the manuscript that will clarify
your concerns about our work.

\begin{itemize}
\item Differences four-dimensional approach vs higher dimensional: We have
added extra paragraphs to our discussion section pointing to the
following differences
\begin{itemize}
\item Gravitational models including higher order curvature/torsion are in
general not minimal, whereas our model is (in the sense that
torsion enters only through (non-Riemannian) curvature, and only an
extra coupling constant is introduced in our analysis).
\item The initial model assures that the field equation are first order.
\item The different behavior of the cosmological model are driven by the
relation between the coupling constant solely.
\item All the cosmological scenarios presented are free of \emph{finite time}
future singularities.
\item Some scenarios present dynamical compactification.
\item The scale factor does not depend on the \(k\) parameter which module
the metric structure.
\item The equations of our model is possesses only \emph{CP} even terms.
\end{itemize}

\item We thank the comment about the generalized theories with no ghosts
and second order field equations. On that respect, we have added a
comment in the manuscript, with its respective reference.

\item We have clarified that some of the wormhole solutions, mentioned in
page 2, violate no energy conditions.

\item Your question about future singularities was addressed in the
previous point regarding the difference between four- and five-
dimensional models.
\end{itemize}

\begin{center}
Best regards,

Simon del Pino

On behalf of the collaborators
\end{center}
\end{document}
